### 2016-01-05 lafaySsf2016.tex
initial commit

\documentclass{aes2e}

% Metadata Information
\jyear{2016}
\jmonth{January}
\jvol{1}
\jnum{1}

\usepackage{color}
\newcommand{\gregoire}[1]{\textcolor{red}{Gregoire : #1}}
\newcommand{\mathias}[1]{\textcolor{green}{Mathias : #1}}
\newcommand{\ml}[2][]{\textcolor{blue}{#1 #2}}

%\renewcommand{\baselinestretch}{1.5}

\begin{document}
%
% --- Author Metadata here ---
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{A novel interface for audio based sound data mining}
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%

%Author Info.
\authorgroup{
\author{Gr\'egoire Lafay$^1$}, \author{Nicolas Misdarris$^2$},
\author{Mathieu Lagrange$^1$}, \author{Mathias Rossignol$^2$}
\email{(gregoire.lafay@irccyn.ec-nantes.fr)}
\affil{IRCCyN, Ecole Centrale de Nantes $^1$ \quad\quad\quad\quad IRCAM $^2$}
}

%\author{
%\alignauthor
% Gregoire Lafay\\
%       \affaddr{IRCCyN, Ecole Centrale de Nantes}\\
%       \affaddr{1 rue de la Noe}\\
%       \affaddr{Nantes, France} 
%       %\email{gregoire.lafay@irccyn.ec-nantes.fr} 
%\alignauthor 
% Nicolas Misdarris\\
%       \affaddr{IRCAM}\\
%       \affaddr{1 Place Igor\-Stravinsky}\\
%       \affaddr{Paris, France} \\
%       \email{}
%\and
%\alignauthor 
%Mathieu Lagrange\\
%       \affaddr{IRCCyN, Ecole Centrale de Nantes}\\
%       \affaddr{1 rue de la Noe}\\
%       \affaddr{Nantes, France} \\
%\alignauthor 
% Mathias Rossignol\\
%       \affaddr{IRCAM}\\
%       \affaddr{1 Place Igor\-Stravinsky}\\
%       \affaddr{Paris, France} \\
%       \email{}
%}

% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{26 December 2014 }
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
In this paper, the design of a web interface  for audio-based sound data mining is studied. The interface allows the user to explore a sound dataset without any written textual hint. Dataset sounds are grouped into semantic classes which are themselves clustered to build a semantic hierarchical structure. Each class is represented by a circle distributed on a two dimensional space according to its semantic level. Several means of displaying sounds following this template are presented and evaluated with a crowdsourcing experiment.
\end{abstract}

%% A category with the (minimum) three required fields
%\category{H.2.8}{Database Management}{Database Applications}[Data mining]
%%%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%\terms{Design, Experimentation, Human Factors}

\keywords{audio-based sound data mining, listening oriented user interface, crowdsourcing experiment}

\section{Introduction}

With the growing capability of recording and storage, the problem of indexing large databases of audio has recently been the object of much attention \cite{Wold1996}. Most of that effort is dedicated to the automatic inference of indexing metadata from the actual audio recording \cite{Zhang1999, Tzanetakis1999}; in contrast, the ability to browse such databases in an effective manner has been less considered. The temporal aspect of sounds has been studied in \cite{Kobayashi1997} and the use of multidimensional projection of audio features in \cite{Cano2002}.

Typically, a sound data mining interface is based on keyword-driven queries. The user enters a word which characterizes the desired sound, and the interface presents him with sounds related to this word. The effectiveness of this principle is primarily based on the typological structure and nomenclature of the database. However, some issues arises from this paradigm:

\begin{enumerate}
\item Sounds, as many others things, can be described in many ways. Sound may be designated by their
sources (a car door), as well as by the action of those sources (the slamming of a car door) or their environments (slamming a car door in a garage) \cite{houix2012lexical, niessen2010categories, brown2011towards}. Designing an effective keyword-based search system requires an accurate description of each sound, which has to be adaptable to the sound representation of each user.
\item Pre-defined verbal descriptions of the sounds made available to the users may potentially bias their selections.
\item Localization of the query interface is made difficult as the translation of some words referring to qualitative aspects of the sound such as its ambience is notoriously ambiguous and subject to cultural specificities.
\item Unless considerable time and resources are invested into developing a multilingual interface, any system based on verbal descriptions can only be used with reduced performance by non-native speakers of the chosen language.
\end{enumerate}

To avoid those issues, we propose in this paper several means of displaying sounds without relying on any textual representation. Their effectiveness is studied with a search-based task whose aim is to listen to a target sound, and browse the database using the evaluated display to find this target sound as fast as possible. The proposed displays are first described along with the dataset used for evaluation. We then explain the chosen validation protocol, before finally presenting and discussing performance results.

\section{Interface framework}

\subsection{Dataset structure}

The interface framework requires a pre-organization of the sound dataset it has to display. This organization is based on semantic considerations. A sound is characterized by a tag describing the source of the sound (\textit{man-yelling}, \textit{car-passing}). Sounds are then grouped into classes according to their tags (\textit{car} $>$ \textit{car-passing}; car $>$ \textit{car-starting}). Those classes are in turn  packed into classes until high level classes describing broad concepts are reached (\textit{traffic} $>$ \textit{car} $>$ \textit{car-passing}). The sound dataset is organized into a hierarchical structure of semantic classes as described in Figure~\ref{figdataset}. The different levels of this hierarchy are called semantic levels. Strictly speaking, the sound samples of the dataset are the leaf semantic levels. All the other classes are represented by a \textit{prototype sound} that best characterizes the sounds belonging to the class.

That description implies that there are as many leaf classes as sounds in the dataset, which would be unrealistic for large datasets. We therefore propose, in that case, to adapt the organization by considering the leaf classes as \textit{collections} of semantically similar sound samples. Thus, two sounds of \textit{male-yelling} would be grouped into a  single leaf class with the tag \textit{male-yelling}. The leaf class would then also have a \textit{prototype sound} being the most representative item of the different \textit{male-yelling} sounds belonging to the leaf class.

In order for the semantic hierarchical structure to be perceptually valid, the \textit{tags} describing the classes  were chosen from the names of sound categories found by studies addressing environmental auditory scenes perception \cite{niessen2010categories, brown2011towards, dubois2006cognitive}. In cognitive psychology, sound categories may be regarded as intermediaries between collective sound representations and individual sensory experiences \cite{dubois2006cognitive}. It is our belief that using such category names to build the hierarchical structure makes the latter perceptually motivated, and thus meaningful for the users.

%We choose to represent each sound leaf class of our sound data set with a circle. Circles are dispatched in a 2D space and packed together according to the hierarchical structure of the sound set (see figure \ref{SSF}). The class is identified by listening to an acoustic prototype chosen by the authors for its representativeness. Clicking on a circle fires this audio prototype.

\subsection{Displays}

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.24]{gfx/dataset.pdf} 
\end{center}
\caption{\label{figdataset} Semantic hierarchical structure of the dataset of urban environmental sounds}
\end{figure}

\begin{figure}[t!]
\begin{center}
\includegraphics[scale=0.30]{gfx/XP2clean.pdf} 
\end{center}
\caption{\label{figXP2} Full Display (FD) with a non visible hierarchical organization of semantic classes }
\end{figure}

\begin{figure*}[t]
\begin{center}
\includegraphics[scale=0.4]{gfx/XP1clean2.pdf} 
\end{center}
\caption{\label{figXP1} Progressive Display (PD) with a  visible hierarchical organization of semantic classes: (a) initial folded version; (b) partly folded version; (c) unfolded version}
\end{figure*}

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.24]{gfx/SSF.pdf} 
\end{center}
\caption{\label{figSSF} Spatial configuration of the Progressive Display (PD) based on the semantic hierarchical structure of the dataset}
\end{figure}

In this section, two listening oriented displays, called respectively Progressive Display (PD) and Full Display (FD) are presented. Both interfaces 1) allow users to explore a sound dataset without any written textual help and 2) base their display upon the hierarchical structure of the dataset. The two interfaces have been designed in order to see whether a progressive top-down display of the hierarchical structure helps the user explore the dataset.

As shown on Figures~\ref{figXP2} and \ref{figXP1}, both interfaces are based on the same principle of distributing in a 2D space the hierarchical structure of the  sound elements of the dataset. Each sound class is represented by a circle. Circles are packed together according to the hierarchical semantic organization of the dataset, as shown on Figure \ref{figSSF}. Thus, subclasses belonging to the same class are close to each others. Circle packing functions of the D3.js (Data-Driven Documents) javascript library \cite{2011-d3} are used to distribute the sound classes in the space.

The way in which a user visualizes the hierarchical organization varies with the interface:

\begin{itemize}
\item \textit{PD}: users have access to the intermediate semantic levels of the hierarchy. Upon first using PD, they observe circles representing the top classes of the semantic hierarchical structure of the dataset. When users click on a circle, they hear the sound prototype of the class and the subclasses are progressively revealed, represented by white circles contained in the circle that has just been clicked. The same action repeats itself until the user reaches the leaf classes of the hierarchy. The leaf classes are represented with small gray circles, indicating that there is no subclass to discover. Thus the PD has a constrained exploration system. When a user click on a circle, sub-circles are  automatically revealed to him in a gradual way. Each time a sub-circle is automatically revealed, its sound prototype is played. Users may stop the discovery process by clicking on an other circle. 
\item \textit{FD}: users can directly visualize the whole hierarchy, down to the leaf classes. Those leaf classes are distributed  in the same manner as PD. In that sense, the spatial configuration of the unfolded version of PD, which may be obtained after discovering all the classes and subclasses, is similar to that of  FD, as shown on Figures~\ref{figXP2}, \ref{figXP1} and \ref{figSSF}.
\end{itemize}

\section{Validation test}
\subsection{Objective}
During this test, three interfaces are compared:
\begin{itemize}
\item PD, which provides a visible hierarchical organization of semantic classes;
\item FD, which provides a non-visible hierarchical organization of semantic classes;
\item an Acoustic Display (AD) providing a 2D representation based on acoustic descriptors. In this case, the spatial configuration is computed by 1) describing the sounds with mel-frequency cepstrum coefficients (MFCCs) and 2) using  a non metric multidimensional scaling with Kruskal's normalized stress to compute sound positions in a 2D space. An example of AD can be seen on~Figure \ref{figXP3}.
\end{itemize}

By comparing PD and FD, the effect of a visible hierarchy on the user is investigated. The goal is to check if forcing the user to browse the high levels of the hierarchy helps him to understand and learn the spatial configuration and the organisation of the sound classes. By confronting PD and FD with AD, the relative efficiencies of semantic based and acoustic based spatial configurations are compared.

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.18]{gfx/XP3.pdf} 
\end{center}
\caption{\label{figXP3} The Acoustical Display (AD) computed using  a non metric multidimensional scaling on  MFCCs based acoustic descriptors}
\end{figure}

\subsection{Experimental protocol}
We choose to test and compare the three displays through a crowdsourcing experiment. Here is the link to access the experiment web page \footnote{http://217.70.189.118/soundthings/speedSoundFinding/}.

Subjects are asked to successively retrieve 13 target sounds in a dataset of 149 urban environmental sound events. The target sounds are distributed such as there are at least two target sounds in each top-level class of the semantic hierarchical structure of the dataset. To minimize order effects, target sounds are randomly presented to each subject.

To listen to a target sound, the subject has to click a \textit{Play target sound} button. Subject may replay the target sound  as many times as they like. A timer is started when the subject clicks on a circle for the first time.

When the target sound is found, subject puts an end to its search by clicking on the \textit{Click if found} button. This action 1) stops the timer and 2) loads a new target sound. If the subject does not find the correct target sound, an error message appears, and the experiment continues.

During the experiment, two indications are communicated to the subject:
\begin{itemize}
\item \textit{The research state}: "pause" if the subject is currently not looking for a target sound (at the beginning of the experiment, or between two target sounds);  "in progress"  if the subject is currently looking for a target sound.
\item \textit{Remaining target sounds}: the number of target sounds which remain to find.
\end{itemize}

The experiment ends when all the target sounds have been found.

It is most important to note that PD do not pack up at each target sound search. When a circle is revealed, it remains visible during the whole experiment.

\subsection{Data Collection}
Three sets of data are collected during the experiment:
\begin{itemize}
\item the total duration of the entire experiment. It includes breaks between two target sound searches and it is called the \textit{absolute duration}.
\item the duration of each search. The sum of the 13 duration searches, which is the \textit{absolute duration} minus the break times between two target sound searches, is called \textit{the relative duration}.
\item the name of each sound which has been heard.
\item the time at which each sound has been heard.  
\end{itemize}

\subsection{Apparatus}
A crowd sourcing approach has been adopted. The experiment was designed to be supported by the  \textit{chrome-browser} web navigator. The link to the experiment has been sent to the subjects via three  mailing list being \textit{music-ir, auditory} and \textit{uuu-IRCAM} (internal IRCAM mailing list). Subjects were allowed to perform the experiment once, and on one interface only. Data were automatically collected server-side at the end of the experiment. Subjects were asked to use headphones. All the presented sounds were normalized to the same root mean square (RMS) level.
\subsection{Participants}
60 subject have completed the experiment, 20 for each interface.
\section{Results}
\subsection{Outlier detection}
Outlier detection is an important step of any crowdsourcing experiment as experimenters do not control the experimental environment in which the subjects perform the experiment \cite{komarov2013crowdsourcing}\cite{buchholz2011crowdsourcing}. A widely used method to detect outlier in  human-computer interaction studies is to consider as outlier an observation  which deviates of at least $\pm2$ standard deviation ($STD$) from the average \cite{komarov2013crowdsourcing}. As this method is not robust to the presence of isolated extreme observations (as it is often the case for crowdsourcing experiment), a method using the inter-quartile range ($IQR$) proposed by \cite{komarov2013crowdsourcing} is used in this paper. With this approach, an observation is considered to be an outlier if it is more than $3*IQR$ higher than the third quartile or more than $3*IQR$ lesser than the first quartile. For normalized distribute data, the $IQR$ method remove less than 0.00023\% of the data whereas the  $STD$ method remove 4.6\% of the data \cite{komarov2013crowdsourcing}. This methods is applied to the following list of parameters:

 \begin{table}[t]
\caption{\label{tab1} Standard deviations of the number of heard sounds per subject, with and without outliers.}
\begin{center}
\begin{tabular}{cccc}
Interface &  PD &  FD& AD \\
\hline
with outliers & 141 & 155 & 315 \\
without outliers & 139 & 140 & 146 \\  
\hline
\end{tabular}
\end{center}
\end{table} 
\begin{table}[t]
\caption{\label{tab2} Standard deviations of the \textit{relative duration} per subject, with and without outliers.}
\begin{center}
\begin{tabular}{cccc}
Interface &  PD &  FD &  AD \\
\hline
with outliers & 353 & 277 & 520 \\
without outliers & 363 & 249 & 273 \\  
\hline
\end{tabular}
\end{center}
\end{table}

\begin{table}[t]
\caption{\label{tab3} Standard deviations of the \textit{absolute duration} per subject, with and without outliers.}
\begin{center}
\begin{tabular}{cccc}
Interface &  PD &  FD &  AD \\
\hline
with outliers & 1401 & 339 & 4034 \\
without outliers & 408 & 327 & 273 \\  
\hline
\end{tabular}
\end{center}
\end{table}

\begin{itemize}
\item durations of  each target sound search
\item average duration of target sound searches
\item maximum duration of target sound searches
\item \textit{relative duration}
\item \textit{absolute duration}
\item number of heard sounds for each target sound search
\item average number of heard sounds
\item maximum number of heard sounds
\item total number of heard sounds
\end{itemize}

Using the $IQR$ method, 4 subjects are detected as outliers and removed from the analysis. 1 subject used PD, 1 subject FD, and 2 subjects AD. 2 subjects are detected by observing the \textit{absolute duration} (4 and 12 hours), 1 subject by observing the total number of heard sounds (1800 heard sounds, roughly 12 times the total size of the corpus) and 1 subject by observing the number of heard sound for the first target sound search (321 heard sounds, including 21 times the target sound).

The tables \ref{tab1}, \ref{tab2} and \ref{tab3} measure the effect of the presence of the  outliers on the standard deviations of three observed data being the total number of heard sounds, the \textit{relative duration} and the \textit{absolute duration}. The removal of the outliers have important effects on the data distributions, specially for AD.

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.36]{gfx/RDbis.pdf} 
\end{center}
\caption{\label{fig1} Boxplot representing the distributions of the relative durations for the PD, FD and AD}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.36]{gfx/NBRbis.pdf} 
\end{center}
\caption{\label{fig2} Boxplot representing the distributions of the numbers of heard sounds for the PD, FD and AD}
\end{figure}

\subsection{Interface efficiencies}

To characterize the displays efficiencies, three set of collected data are assessed: 
\begin{itemize}
\item the \textit{relative duration} 
\item the number of heard sounds 
\item the number of heard sounds without duplication. By "without duplication" we mean that, if a same sound prototype is heard 10 times during the 13 searches of the  experiment, it counts only for one. 
\end{itemize}

The two first data help us qualify the notion of efficiency by considering the time and the number of clicks needed to achieve the task (ie. reach the target). The goal for those values is to be as low as possible. The third data allows us to measure the selectivity of the interfaces. A low number of heard sounds without duplication indicates that subjects understood the spatial organisation of the dataset, and use this knowledge to improve their searches. In contrary, a high number of heard sounds without duplication suggest that the subject did not understood the way circles are organized in space, and tends to play all the sounds at each search. The maximum number of heard sounds without duplication is the corpus size: 149 sounds.

Concerning the \textit{relative durations}, distributions of the data are displayed on Figure~\ref{fig1} for the three interfaces. FD seems to  perform better  than the other interfaces, whereas PD and AD seem to have similar results. To refine the analysis, a two sided Wilcoxon rank sum test is considered. It is a non parametric statistical test which tests the null hypothesis that two set of observed data originate from distributions having equal median \cite{gibbons2011nonparametric}. As expected, FD is significantly better than the other interfaces (FD-PD: $p=0.0142$; FD-AD: $p=0.028$) and there is no statistical differences between PD and AD (PD-AD: $p=1$).

Distributions of  the numbers of heard sounds are displayed on Figure~\ref{fig2} for the three interfaces. Results are similar of those observed for the \textit{relative durations}. FD significantly outperforms the other interfaces (FD-PD: $p=0.0115$; FD-AD: $p=0.018$), whereas PD and AD show similar outcomes (PD-AD: $p=0.3699$).

Lastly, Figure~\ref{fig3} displays the distributions of the number of heard sounds without duplication. This time the results of AD are significantly lower than those of both PD and FD (AD-PD: $p=8.4910.10^{-4}$; AD-FD: $p=0.027$).  For AD, 75\% of the subjects heard more than 138 sounds, and 25\% heard 148 sounds, that is almost the entire database. Considering PD, 75\% of subjects heard  less than 133 sounds, against 144 for FD. There is no statistical differences between the PD and FD (PD-FD: $p=0.8607$), indicating that those two interfaces perform equally.

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.36]{gfx/SDbis.pdf} 
\end{center}
\caption{\label{fig3} Boxplot representing the distributions of the numbers of heard sounds without duplication for the PD, FD and AD}
\end{figure}

According to those results, a hierarchical organization of the dataset based on semantic values (PD and FD) allows users to retrieve the 13 target sounds 1) quicker, and 2) by listening to a smaller amount of sounds than an organization based on acoustic descriptors (AD). But those two effects are significantly compromised when users have to parse the entire hierarchy to reach the first target sound, as it the case for PD. It seems that imposing a graphical representation of the hierarchy disturbs or confuses the user instead of allowing him to learn the spatial organization of the classes. 
%The semantic-based spatial organization allow users to significantly improve the precision of their searches, by clicking on less different circles, irrespective of whether or not the hierarchy is imposed.

\newpage

\subsection{Learning phenomenon}

We now study if and how users  progressively acquire knowledge about the spatial organization of the classes. To do that, variations of the data over the searches are assessed. Three sets of collected data are used:

\begin{itemize}
\item the duration of each target sound search 
\item the number of heard sounds for each target sound search 
\item the number of heard sounds for each target sound search without duplication. 
\end{itemize}

Figure \ref{fig1PR} (bottom) displays the evolution of the medians of the durations of each target sound search observed over the subjects for PD, FD and AD. It is interesting to note that both for PD and FD, the maximum value is observed for the first search, whereas it is observed for the fourth search for AD. Moreover if the curve profiles of PD and FD  seem to progressively decrease and are very similar, the one of AD is much more irregular. If we compare PD and FD, we note  that the durations are systematically shorter for FD, except for the search 12. Furthermore, for FD, a threshold of 25 seconds is reached from the search four,  whereas it is of 50 seconds for PD.

\begin{figure}[t!]
\begin{tabular}{c}
\includegraphics[scale=0.29]{gfx/NOHS.pdf} \\
\includegraphics[scale=0.29]{gfx/NOHS.pdf} \\ % should be WD
\includegraphics[scale=0.29]{gfx/duration.pdf} 
\end{tabular}
\caption{\label{fig1PR} Medians of (top) the numbers of heard sounds at each target sound search, (middle) the numbers of heard sounds without duplication (WD) at each target sound search, (bottom) the relative durations at each target sound search \gregoire{peux-tu ajouter WD.pdf ? la deuxieme figure}}
\end{figure}
%\begin{figure}[t]
%\begin{center}
%\includegraphics[scale=0.38]{gfx/nbrPR.pdf} 
%\end{center}
%\caption{\label{fig2PR} Medians of the numbers of heard sounds at each target sound search (abscissa) for the PD (top), FD (middle) and AD (bottom)}
%\end{figure}
%\begin{figure}[t]
%\begin{center}
%\includegraphics[scale=0.38]{gfx/sdPR.pdf} 
%\end{center}
%\caption{\label{fig3PR} Medians of the numbers of heard sounds without duplication at each target sound search (abscissa) for the PD (top),  FD (middle) and AD (bottom)}
%\end{figure}
Figure \ref{fig1PR} (top) displays the evolution of the medians of the numbers of heard sounds for each target sound search, observed over the subjects for the three interfaces. If the curve profiles of PD and AD seem to be similar to those respectively observed for the durations, here the maximum value for FD  is reached for the third search. Again, values of FD are mostly below those of PD, except for the search index 3,10 and 12. For both PD and FD the curves oscillate from the search four, but those oscillations occur in a range of $9-17$ for FD and  $9-30$ for PD. Similar results are found for the evolution of the medians of the numbers of heard sounds without duplication, shown on Figure \ref{fig1PR} (middle).

Those results tend to indicate that PD and FD facilitate the learning of the spatial configuration, as the search durations and the numbers of heard sounds at each search decrease over time. Although curves for PD and FD have similar profiles, FD seem to better perform as users of FD where able to find the target sounds faster by clicking on fewer circles.

\section{Conclusion}

In this paper, two displays allowing users to explore a sound dataset without written textual help are presented. The interfaces distribute sounds represented by circles on a 2D space. The spatial organisation is driven by semantic features. The two graphical displays are assessed and compared to a third listening based interface in which spatial configuration depends upon acoustic features. The tests consist in data retrieval tasks. 
The Full-Display (FD), that allows users to directly visualize the leaf classes of the semantic hierarchical structure, proves to be the most effective interface for the task.

Two main conclusions may be derived from this experiment. First, a spatial configuration based on semantic features is more effective to retrieve target sounds than a spatial configuration based on acoustic features. Second, an imposed visualisation of the semantic hierarchical structure of the dataset does not help user to understand and learn the spatial configuration of the semantic class, but instead disturbs the navigation.

%It has to be noted that the test datasets are relatively small in terms of leaf sound classes. It may be interesting to test the FD on a larger dataset. 
%
%To conclude, if this study goal was to investigate the viability of an audio based data mining interface, other studies addressing comparison between FD and keyword based interfaces remain to be conducted.

\section{Acknowledgements}

Research project partly funded by ANR-11-JS03-005-01.

\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case

%In particular, much consideration has been given in order to avoid any semantic bias. Indeed any selection process based on keyword searching would have influenced the subject choices, and makes impossible or inadvisable to run any kind of identification tasks. 
%
%In a previous study \cite{ssf} we found  that the perceptively inspired structured of the sound data set (on which depends the spatial configuration of the circles) offers a easy-to-navigate format that allows subject to quickly understand the spatial localization of the different sound element classes and thus  enable them to quickly find a sound after a short amount of time. 
%
%%The interface selection has been tested in a previous \textit{crowd-sourcing} experiment on 60 subject. The goal was to find 10 target sound. Results showed that this interface allows subject to learn the sound spatial organisation, and to quickly retrieve a sound after some trials. 
\end{document}

### 2016-02-03 lafaySsf2016.tex
find better title
abstract, intro, and a good section of previous work done

@@
 
-\title{A novel interface for audio based sound data mining}
+\title{Semantic browsing of sound databases \\ without keywords}
 %\subtitle{[Extended Abstract]
@@
 \email{(gregoire.lafay@irccyn.ec-nantes.fr)}
-\affil{IRCCyN, Ecole Centrale de Nantes $^1$ \quad\quad\quad\quad IRCAM $^2$}
+\affil{IRCCyN, Ecole Centrale de Nantes $^1$ \quad\quad\quad\quad IRCAM, STMS, UPMC $^2$}
 }
@@
 \maketitle
+
 \begin{abstract}
-In this paper, the design of a web interface  for audio-based sound data mining is studied.
-The interface allows the user to explore a sound dataset without any written textual hint.
-Dataset sounds are grouped into semantic classes which are themselves clustered to build a semantic hierarchical structure.
-Each class is represented by a circle distributed on a two dimensional space according to its semantic level.
-Several means of displaying sounds following this template are presented and evaluated with a crowdsourcing experiment.
+In this paper, we study the relevance of a semantic organization of sounds for easing the browsing of a sound database.
+For such task, the semantic organization is traditionally hold thanks to a keyword selection process.
+However, various issues of written language like word polysemy, ambiguities translation issues may bias the browsing process.
+
+We consider here two display of sounds that are organized by considering underlying semantic information that draws a hierarchy.
+For the sake of comparison, we also consider a display whose organization of sound in the plane is based on acoustic cues.
+Those three displays are evaluated in terms of search speed in a crowd sourcing experiment.
+Results
 \end{abstract}
@@
 
-\keywords{audio-based sound data mining, listening oriented user interface, crowdsourcing experiment}
+\keywords{Audio content management and display, Semantic sound data mining}
 
@@
 With the growing capability of recording and storage, the problem of indexing large databases of audio has recently been the object of much attention \cite{Wold1996}.
-Most of that effort is dedicated to the automatic inference of indexing metadata from the actual audio recording \cite{Zhang1999, Tzanetakis1999}; in contrast, the ability to browse such databases in an effective manner has been less considered.
-The temporal aspect of sounds has been studied in \cite{Kobayashi1997} and the use of multidimensional projection of audio features in \cite{Cano2002}.
+Most of that effort is dedicated to the automatic inference of indexing metadata from the actual audio recording \cite{Zhang1999, tzanetakis2002musical}; in contrast, the ability to browse such databases in an effective manner has been less considered.
 
-Typically, a sound data mining interface is based on keyword-driven queries.
-The user enters a word which characterizes the desired sound, and the interface presents him with sounds related to this word.
+Most media assets management are based on keyword-driven queries.
+The user enters a word which characterizes the desired item, and the interface presents him with items related to this word.
 The effectiveness of this principle is primarily based on the typological structure and nomenclature of the database.
-However, some issues arises from this paradigm:
+However, for databases and more specifically for databases of sounds, issues arise:
 
@@
 Designing an effective keyword-based search system requires an accurate description of each sound, which has to be adaptable to the sound representation of each user.
-\item Pre-defined verbal descriptions of the sounds made available to the users may potentially bias their selections.
-\item Localization of the query interface is made difficult as the translation of some words referring to qualitative aspects of the sound such as its ambience is notoriously ambiguous and subject to cultural specificities.
+\item Pre-defined verbal descriptions of the sounds made available to the users may potentially bias their browsing and final selection.
+\item Localization of the query interface is made difficult as the translation of some words referring to qualitative aspects of the sound such as its ambiance is notoriously ambiguous and subject to cultural specificities.
 \item Unless considerable time and resources are invested into developing a multilingual interface, any system based on verbal descriptions can only be used with reduced performance by non-native speakers of the chosen language.
@@
 
-To avoid those issues, we propose in this paper several means of displaying sounds without relying on any textual representation.
-Their effectiveness is studied with a search-based task whose aim is to listen to a target sound, and browse the database using the evaluated display to find this target sound as fast as possible.
-The proposed displays are first described along with the dataset used for evaluation.
-We then explain the chosen validation protocol, before finally presenting and discussing performance results.
+To circumvent those issues, not relying on keywords is desirable. That said, semantics, which is traditionally conveyed through written text, may be important for easing browsing.
 
-\section{Interface framework}
+To study this issue, we consider in this paper several means of displaying sounds without relying on any textual representation.
+One, considered as a reference baseline, do not rely on any semantic information as the sounds are organized according to their acoustical properties, \textit{i.e.} time averaged spectral features.
+The second one
 
-\subsection{Dataset structure}
+Their effectiveness is studied with a search-based task whose aim is find a target sound by browsing the database using the display under evaluation.
+
+The paper is organized as follows: in Section \ref{previous} previous work on the topic of sound database browsing is reviewed.
+Then, the corpus used in this study is described in Section \ref{dataset}.
+the three displays under evaluation are next described in Section \ref{display}.
+The crowd sourcing test used to compare those displays is presented in Section \ref{test}.
+the outcomes of this experiment are discussed in Section \ref{results}.
+
+\section{Previous work} \label{previous}
+
+Most of the research effort in Sound Design and Music Information Retrieval (MIR) communities is focused on acoustical based indexing and browsing \cite{tzanetakis2003musescape, streich2008music}.
+Typically, the items (sound effects or pieces of music) are modeled by processing the digital audio waveform using some signal processing pipeline in order to get a compact description for each  items
+\cite{coleman2007mused}.
+Then, statistical projections or embedding techniques, like Principal components analysis (PCA),  Multi Dimensional Scaling (MDS) \cite{schwarz2009sound,Cano2002}, Self Organizing Maps (SOM) \cite{pampalk2004exploring} and the like are used to project the items in a two or three dimensional space while preserving as much as possible distances among items.
+One of the advantage of such an approach is its ability to scale to very large databases \cite{schwarz2009scalability} as it does not need any kind of manual annotations and allows to search by similarity efficiently according to acoustical properties.
+
+Though, such acoustical models are inherently subject to observation noise and biases.
+Selecting the most relevant features to achieve the correct projection of the data may only be performed by an expert user.
+If done \textit{a priori} by an expert or by some above cited dimensionality reduction technique, the induced bias can strongly limit the user in its ability to .
+In that respect, semantic tags, if available for the data at hand have the advantage of implicitly structuring the similarity space, thus eventually easing the browsing process even if the actual tags are not -- as in this study -- exposed to the user.
+
+\section{Dataset} \label{dataset}
 
@@
 
-\subsection{Displays}
+\section{Displays} \label{display}
 
@@
 
-\section{Validation test}
+\section{Experiment} \label{test}
+
 \subsection{Objective}
@@
 60 subject have completed the experiment, 20 for each interface.
-\section{Results}
+
+\section{Results} \label{results}
 \subsection{Outlier detection}

### 2016-02-11 lafaySsf2016.tex
gl version

@@
 \item the duration of each search.
-The sum of the 13 duration searches, which is the \textit{absolute duration} minus the break times between two target sound searches, is called \textit{the relative duration}.
+The sum of the 13 search durations, which is the \textit{absolute duration} minus the break times between two target sound searches, is called the \textit{total search duration}.
 \item the name of each sound which has been heard.
@@
 All the presented sounds were normalized to the same root mean square (RMS) level.
+
 \subsection{Participants}
-60 subject have completed the experiment, 20 for each interface.
+60 subjects have completed the experiment, 20 for each interface.
+
+\subsection{Statistical analysis}
+We use one-way ANOVA and two samples \textit{t} tests at the 5\% significance level to test for statistical significance.
+All statistical analysis are performed using native functions of the computing environment MATLAB.
 
@@
 \subsection{Outlier detection}
-Outlier detection is an important step of any crowdsourcing experiment as experimenters do not control the experimental environment in which the subjects perform the experiment \cite{komarov2013crowdsourcing}\cite{buchholz2011crowdsourcing}.
-A widely used method to detect outlier in  human-computer interaction studies is to consider as outlier an observation  which deviates of at least $\pm2$ standard deviation ($STD$) from the average \cite{komarov2013crowdsourcing}.
-As this method is not robust to the presence of isolated extreme observations (as it is often the case for crowdsourcing experiment), a method using the inter-quartile range ($IQR$) proposed by \cite{komarov2013crowdsourcing} is used in this paper.
-With this approach, an observation is considered to be an outlier if it is more than $3*IQR$ higher than the third quartile or more than $3*IQR$ lesser than the first quartile.
-For normalized distribute data, the $IQR$ method remove less than 0.00023\% of the data whereas the  $STD$ method remove 4.6\% of the data \cite{komarov2013crowdsourcing}.
-This methods is applied to the following list of parameters:
 
 \begin{table}[t]
-\caption{\label{tab1} Standard deviations of the number of heard sounds per subject, with and without outliers.}
+\caption{\label{tab1} Criterion used to detect outliers}
 \begin{center}
-\begin{tabular}{cccc}
-Interface &  PD &  FD& AD \\
+\begin{tabular}{ccccc}
+criterion      &  1  &  2  &  3  &  4  \\
 \hline
-with outliers & 141 & 155 & 315 \\
-without outliers & 139 & 140 & 146 \\  
-\hline
-\end{tabular}
-\end{center}
-\end{table} 
-\begin{table}[t]
-\caption{\label{tab2} Standard deviations of the \textit{relative duration} per subject, with and without outliers.}
-\begin{center}
-\begin{tabular}{cccc}
-Interface &  PD &  FD &  AD \\
-\hline
-with outliers & 353 & 277 & 520 \\
-without outliers & 363 & 249 & 273 \\  
+outlier 1 (PD) & --- & --- & --- &  x  \\
+outlier 2 (PD) &  x  & --- & --- & --- \\  
+outlier 3 (AD) & --- & --- & --- &  x \\
+outlier 4 (AD) &  x  &  x  &  x  & --- \\
+outlier 5 (AD) &  x  & --- & --- & --- \\
 \hline
@@
 
-\begin{table}[t]
-\caption{\label{tab3} Standard deviations of the \textit{absolute duration} per subject, with and without outliers.}
-\begin{center}
-\begin{tabular}{cccc}
-Interface &  PD &  FD &  AD \\
-\hline
-with outliers & 1401 & 339 & 4034 \\
-without outliers & 408 & 327 & 273 \\  
-\hline
-\end{tabular}
-\end{center}
-\end{table}
+Outlier detection is an important step of any crowdsourcing experiment as experimenters do not control the experimental environment in which the subjects perform the experiment \cite{komarov2013crowdsourcing}\cite{buchholz2011crowdsourcing}.
+A widely used method to detect outlier in  human-computer interaction studies is to consider as outlier an observation  which deviates of at least $\pm2$ standard deviation ($STD$) from the average \cite{komarov2013crowdsourcing}.
+As this method is not robust to the presence of isolated extreme observations (as it is often the case for crowdsourcing experiment), we follow the  method proposed by proposed by \cite{komarov2013crowdsourcing} and used the inter-quartile range ($IQR$).
+With this approach, an observation is considered to be an outlier if its value is more than $3*IQR$ higher than the third quartile or more than $3*IQR$ lesser than the first quartile.
+This methods is applied to the four following criterion :
 
-\begin{itemize}
-\item durations of  each target sound search
-\item average duration of target sound searches
-\item maximum duration of target sound searches
-\item \textit{relative duration}
-\item \textit{absolute duration}
-\item number of heard sounds for each target sound search
-\item average number of heard sounds
-\item maximum number of heard sounds
-\item total number of heard sounds
-\end{itemize}
+\begin{enumerate}
+\item the \textit{absolute duration}: to detect subjects who took abnormally long breaks between searches.
+\item the \textit{total search duration}: to detect subjects who took an abnormally long time to complete the experiment.
+\item number of heard sounds: to detect subjects who had to hear an abnormally high number of sounds to complete the experiment.
+\item the maximum number of times a target sound had to be heard before being found: to detect subjects who had difficulty to recognize the target sounds.
+\end{enumerate}
 
-Using the $IQR$ method, 4 subjects are detected as outliers and removed from the analysis.
-1 subject used PD, 1 subject FD, and 2 subjects AD.
-2 subjects are detected by observing the \textit{absolute duration} (4 and 12 hours), 1 subject by observing the total number of heard sounds (1800 heard sounds, roughly 12 times the total size of the corpus) and 1 subject by observing the number of heard sound for the first target sound search (321 heard sounds, including 21 times the target sound).
-
-The tables \ref{tab1}, \ref{tab2} and \ref{tab3} measure the effect of the presence of the  outliers on the standard deviations of three observed data being the total number of heard sounds, the \textit{relative duration} and the \textit{absolute duration}.
-The removal of the outliers have important effects on the data distributions, specially for AD.
+Using the $IQR$ method, 5 subjects are detected as outliers and removed from the analysis.
+2 subjects used PD and 3 subjects AD.
+The tables~\ref{tab1} show what criterion are use to detect outliers.
+3 subjects are detected by observing the \textit{absolute duration}.
+They spent respectively 50 minutes, 5 hours and 17.5 hours to complete the experiment.
+1 subject is detected by observing the total search duration (46 minutes), and an other by observing the total number of heard sounds (1800 heard sounds, roughly 12 times the total size of the corpus).
+And lastly 2 subjects are detected by observing the maximum number of times they had to hear a target sounds before finding it (4 and 8 times).
 
@@
 \begin{center}
-\includegraphics[scale=0.36]{gfx/RDbis.pdf} 
+\includegraphics[scale=0.4]{gfx/analyse1.eps} 
 \end{center}
-\caption{\label{fig1} Boxplot representing the distributions of the relative durations for the PD, FD and AD}
+\caption{\label{fig1} Average and standard deviation for (a) the total search durations, (b) the total search durations and (c) the total search durations}
 \end{figure}
@@
 \begin{center}
-\includegraphics[scale=0.36]{gfx/NBRbis.pdf} 
+\includegraphics[scale=0.4]{gfx/analyse2.eps} 
 \end{center}
-\caption{\label{fig2} Boxplot representing the distributions of the numbers of heard sounds for the PD, FD and AD}
+\caption{\label{fig2} Medians of (top) the numbers of heard sounds at each target sound search, (middle) the numbers of heard sounds without duplication (WD) at each target sound search, (bottom) the relative durations at each target sound search }
 \end{figure}
@@
 
-To characterize the displays efficiencies, three set of collected data are assessed: 
+To characterize the interface efficiencies, three set of collected data are assessed:
+
 \begin{itemize}
-\item the \textit{relative duration} 
+\item the \textit{total search duration}
 \item the number of heard sounds 
@@
 
-Concerning the \textit{relative durations}, distributions of the data are displayed on Figure~\ref{fig1} for the three interfaces.
-FD seems to  perform better  than the other interfaces, whereas PD and AD seem to have similar results.
-To refine the analysis, a two sided Wilcoxon rank sum test is considered.
-It is a non parametric statistical test which tests the null hypothesis that two set of observed data originate from distributions having equal median \cite{gibbons2011nonparametric}.
-As expected, FD is significantly better than the other interfaces (FD-PD: $p=0.0142$; FD-AD: $p=0.028$) and there is no statistical differences between PD and AD (PD-AD: $p=1$).
+The type of interface have a significant effect on the the \textit{total search durations} ($F[2,52]=3.63; p=0.03$), the number of heard sounds ($F[2,52]=5.65; p<0.01$) and the number of heard sounds without duplication ($F[2,52]=8.65; p<0.01$).
 
-Distributions of  the numbers of heard sounds are displayed on Figure~\ref{fig2} for the three interfaces.
-Results are similar of those observed for the \textit{relative durations}.
-FD significantly outperforms the other interfaces (FD-PD: $p=0.0115$; FD-AD: $p=0.018$), whereas PD and AD show similar outcomes (PD-AD: $p=0.3699$).
+Figure~\ref{fig1}~(a) shows the average and standard deviations of the \textit{total search durations} for the three interfaces.
+FD seems to  perform better  than the other interfaces (FD-PD: $p=0.04$; FD-AD: $p=0.02$), whereas PD and AD seem to have similar results (PD-AD: $p=0.88$).
 
-Lastly, Figure~\ref{fig3} displays the distributions of the number of heard sounds without duplication.
-This time the results of AD are significantly lower than those of both PD and FD (AD-PD: $p=8.4910.10^{-4}$; AD-FD: $p=0.027$).
+We found similar results for the numbers of heard sounds (Figure~\ref{fig1}~(b)).
+FD significantly outperforms the other interfaces (FD-PD: $p=0.02$; FD-AD: $p<0.01$), whereas PD and AD show similar outcomes (PD-AD: $p=0.42$).
+
+Figure~\ref{fig1}~(c) shows the results for the number of heard sounds without duplication.
+This time the results of AD are significantly lower than those of both PD and FD (AD-PD: $p<0.01$; AD-FD: $p<0.01$).
 For AD, 75\% of the subjects heard more than 138 sounds, and 25\% heard 148 sounds, that is almost the entire database.
 Considering PD, 75\% of subjects heard  less than 133 sounds, against 144 for FD.
-There is no statistical differences between the PD and FD (PD-FD: $p=0.8607$), indicating that those two interfaces perform equally.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[scale=0.36]{gfx/SDbis.pdf} 
-\end{center}
-\caption{\label{fig3} Boxplot representing the distributions of the numbers of heard sounds without duplication for the PD, FD and AD}
-\end{figure}
+This time PD and FP seem to perform equally (PD-FD: $p=0.58$)
 
@@
 It seems that imposing a graphical representation of the hierarchy disturbs or confuses the user instead of allowing him to learn the spatial organization of the classes.
-%The semantic-based spatial organization allow users to significantly improve the precision of their searches, by clicking on less different circles, irrespective of whether or not the hierarchy is imposed.
-
-\newpage
 
@@
 
-Figure \ref{fig1PR} (bottom) displays the evolution of the medians of the durations of each target sound search observed over the subjects for PD, FD and AD.
+Figure~\ref{fig2}~(a) shows the evolution of the average durations of each target sound search observed for PD, FD and AD.
 It is interesting to note that both for PD and FD, the maximum value is observed for the first search, whereas it is observed for the fourth search for AD.
 Moreover if the curve profiles of PD and FD  seem to progressively decrease and are very similar, the one of AD is much more irregular.
-If we compare PD and FD, we note  that the durations are systematically shorter for FD, except for the search 12.
-Furthermore, for FD, a threshold of 25 seconds is reached from the search four,  whereas it is of 50 seconds for PD.
+Similar results are found observing the evolution of the average numbers of heard sounds with and without duplication (Figures~\ref{fig2}~(b) and (c))
 
-\begin{figure}[t!]
-\begin{tabular}{c}
-\includegraphics[scale=0.29]{gfx/NOHS.pdf} \\
-\includegraphics[scale=0.29]{gfx/NOHS.pdf} \\ % should be WD
-\includegraphics[scale=0.29]{gfx/duration.pdf} 
-\end{tabular}
-\caption{\label{fig1PR} Medians of (top) the numbers of heard sounds at each target sound search, (middle) the numbers of heard sounds without duplication (WD) at each target sound search, (bottom) the relative durations at each target sound search \gregoire{peux-tu ajouter WD.pdf ?
-la deuxieme figure}}
-\end{figure}
-%\begin{figure}[t]
-%\begin{center}
-%\includegraphics[scale=0.38]{gfx/nbrPR.pdf} 
-%\end{center}
-%\caption{\label{fig2PR} Medians of the numbers of heard sounds at each target sound search (abscissa) for the PD (top), FD (middle) and AD (bottom)}
-%\end{figure}
-%\begin{figure}[t]
-%\begin{center}
-%\includegraphics[scale=0.38]{gfx/sdPR.pdf} 
-%\end{center}
-%\caption{\label{fig3PR} Medians of the numbers of heard sounds without duplication at each target sound search (abscissa) for the PD (top),  FD (middle) and AD (bottom)}
-%\end{figure}
-Figure \ref{fig1PR} (top) displays the evolution of the medians of the numbers of heard sounds for each target sound search, observed over the subjects for the three interfaces.
-If the curve profiles of PD and AD seem to be similar to those respectively observed for the durations, here the maximum value for FD  is reached for the third search.
-Again, values of FD are mostly below those of PD, except for the search index 3,10 and 12.
-For both PD and FD the curves oscillate from the search four, but those oscillations occur in a range of $9-17$ for FD and  $9-30$ for PD.
-Similar results are found for the evolution of the medians of the numbers of heard sounds without duplication, shown on Figure \ref{fig1PR} (middle).
-
-Those results tend to indicate that PD and FD facilitate the learning of the spatial configuration, as the search durations and the numbers of heard sounds at each search decrease over time.
+These results tend to indicate that PD and FD facilitate the learning of the spatial configuration, as the search durations and the numbers of heard sounds at each search decrease over time.
 Although curves for PD and FD have similar profiles, FD seem to better perform as users of FD where able to find the target sounds faster by clicking on fewer circles.

### 2016-02-11 lafaySsf2016.tex
submit to the king of English living in a deep forest ;)

@@
 
-\usepackage{color}
+\usepackage{color,url}
 \newcommand{\gregoire}[1]{\textcolor{red}{Gregoire : #1}}
@@
 \begin{abstract}
-In this paper, we study the relevance of a semantic organization of sounds for easing the browsing of a sound database.
+~In this paper, we study the relevance of a semantic organization of sounds for easing the browsing of a sound database.
 For such task, the semantic organization is traditionally hold thanks to a keyword selection process.
@@
 
-The interface framework requires a pre-organization of the sound dataset it has to display.
-This organization is based on semantic considerations.
-A sound is characterized by a tag describing the source of the sound (\textit{man-yelling}, \textit{car-passing}).
-Sounds are then grouped into classes according to their tags (\textit{car} $>$ \textit{car-passing}; car $>$ \textit{car-starting}).
-Those classes are in turn  packed into classes until high level classes describing broad concepts are reached (\textit{traffic} $>$ \textit{car} $>$ \textit{car-passing}).
-The sound dataset is organized into a hierarchical structure of semantic classes as described in Figure~\ref{figdataset}.
-The different levels of this hierarchy are called semantic levels.
+\begin{figure}[t]
+\begin{center}
+\includegraphics[scale=0.24]{gfx/dataset.pdf} 
+\end{center}
+\caption{\label{figdataset} Semantic hierarchical structure of the dataset of urban environmental sounds}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[scale=0.24]{gfx/SSF.pdf} 
+\end{center}
+\caption{\label{figSSF} Spatial configuration of the Progressive Display (PD) based on the semantic hierarchical structure of the dataset}
+\end{figure}
+
+The dataset considered in this study is composed of 149 urban environmental sound events.
+It has a semantic organization that is as follows: a sound is characterized by a tag describing the physical source of the sound (\textit{man-yelling}, \textit{car-passing}).
+Sounds are then hierarchically grouped into classes according to their tags (\textit{car} $>$ \textit{car-passing}; car $>$ \textit{car-starting}).
+Those classes are in turn  packed into classes until high level classes describing broader concepts are reached (\textit{traffic} $>$ \textit{car} $>$ \textit{car-passing}).
+The sound dataset is thus organized into a hierarchical structure of semantic classes as described in Figure~\ref{figdataset}.
 Strictly speaking, the sound samples of the dataset are the leaf semantic levels.
@@
 
-That description implies that there are as many leaf classes as sounds in the dataset, which would be unrealistic for large datasets.
-We therefore propose, in that case, to adapt the organization by considering the leaf classes as \textit{collections} of semantically similar sound samples.
-Thus, two sounds of \textit{male-yelling} would be grouped into a  single leaf class with the tag \textit{male-yelling}.
-The leaf class would then also have a \textit{prototype sound} being the most representative item of the different \textit{male-yelling} sounds belonging to the leaf class.
-
-In order for the semantic hierarchical structure to be perceptually valid, the \textit{tags} describing the classes  were chosen from the names of sound categories found by studies addressing environmental auditory scenes perception \cite{niessen2010categories, brown2011towards, dubois2006cognitive}.
+In order for the semantic hierarchical structure to be perceptually valid, the \textit{tags} describing the classes  are chosen from sound categories presented in studies addressing environmental auditory scenes perception \cite{niessen2010categories, brown2011towards, dubois2006cognitive}.
 In cognitive psychology, sound categories may be regarded as intermediaries between collective sound representations and individual sensory experiences \cite{dubois2006cognitive}.
-It is our belief that using such category names to build the hierarchical structure makes the latter perceptually motivated, and thus meaningful for the users.
+It is our belief that using such category names to build the hierarchical structure makes the latter perceptually motivated, and can thus help the users for efficiently browsing the dataset.
 
@@
 \begin{center}
-\includegraphics[scale=0.24]{gfx/dataset.pdf} 
+\includegraphics[scale=0.18]{gfx/XP3.pdf} 
 \end{center}
-\caption{\label{figdataset} Semantic hierarchical structure of the dataset of urban environmental sounds}
+\caption{\label{figXP3} Acoustical Display (AD) \gregoire{fill the dots ;)}}
 \end{figure}
@@
 \end{center}
-\caption{\label{figXP2} Full Display (FD) with a non visible hierarchical organization of semantic classes }
+\caption{\label{figXP2} Full Semantic Display (FSD): implicit hierarchical organization of semantic classes }
 \end{figure}
@@
 \end{center}
-\caption{\label{figXP1} Progressive Display (PD) with a  visible hierarchical organization of semantic classes: (a) initial folded version; (b) partly folded version; (c) unfolded version}
+\caption{\label{figXP1} Progressive Semantic Display (PSD):  explicit hierarchical organization of semantic classes. (a) initial folded version; (b) partly folded version; (c) unfolded version}
 \end{figure*}
 
-\begin{figure}[t]
-\begin{center}
-\includegraphics[scale=0.24]{gfx/SSF.pdf} 
-\end{center}
-\caption{\label{figSSF} Spatial configuration of the Progressive Display (PD) based on the semantic hierarchical structure of the dataset}
-\end{figure}
+The aim of the displays presented in this section is to allow the user to efficiently browse the above presented dataset of sounds explore without any written textual help.
+For each display, a sound is graphically represented by a filled circle whose, once clicked, plays the actual sound.
+The spatial organization is specific to each display.
 
-In this section, two listening oriented displays, called respectively Progressive Display (PD) and Full Display (FD) are presented.
-Both interfaces 1) allow users to explore a sound dataset without any written textual help and 2) base their display upon the hierarchical structure of the dataset.
-The two interfaces have been designed in order to see whether a progressive top-down display of the hierarchical structure helps the user explore the dataset.
+As a reference, an Acoustic Display (AD) provides a spatial representation based on acoustic descriptors.
+For describing each sound, mel-frequency cepstrum coefficients (MFCCs) are computed with standard parameter settings (13 lowest quefrency coefficients are kept out of 40).
+The Euclidean distance between time averaged features is then computed between each sounds.
+A non metric multidimensional scaling with Kruskal's normalized stress is then considered to project the data into two dimensions \ref{figXP3} according to this distance.
 
-As shown on Figures~\ref{figXP2} and \ref{figXP1}, both interfaces are based on the same principle of distributing in a 2D space the hierarchical structure of the  sound elements of the dataset.
-Each sound class is represented by a circle.
-Circles are packed together according to the hierarchical semantic organization of the dataset, as shown on Figure \ref{figSSF}.
+Two semantically oriented displays are then proposed, called respectively Semantic Display (FSD) and Progressive Semantic Display (PSD), respecively shown on Figures~\ref{figXP2} and \ref{figXP1}.
+Both display  consider the hierarchical structure of the dataset to organize sounds.
+Each sound class is represented by a filled circle.
+Circles are then packed together according to the hierarchical semantic organization of the dataset, as shown on Figure \ref{figSSF}.
 Thus, subclasses belonging to the same class are close to each others.
 Circle packing functions of the D3.js (Data-Driven Documents) javascript library \cite{2011-d3} are used to distribute the sound classes in the space.
-
-The way in which a user visualizes the hierarchical organization varies with the interface:
-
+Depending on the display, the user can whether acces directly to each leaf class or has to go through intermediate levels of the hierarchy.
+More precisely: \\
 \begin{itemize}
-\item \textit{PD}: users have access to the intermediate semantic levels of the hierarchy.
+\item \textit{FSD}: users can directly visualize the whole hierarchy, down to the leaf classes.
+Those leaf classes are distributed  in the same manner as PD.
+In that sense, the spatial configuration of the unfolded version of PD, which may be obtained after discovering all the classes and subclasses, is similar to that of  FD, as shown on Figures~\ref{figXP2}, \ref{figXP1} and \ref{figSSF}.
+\item \textit{PSD}: users have access to the intermediate semantic levels of the hierarchy.
 Upon first using PD, they observe circles representing the top classes of the semantic hierarchical structure of the dataset.
@@
 Users may stop the discovery process by clicking on an other circle.
-\item \textit{FD}: users can directly visualize the whole hierarchy, down to the leaf classes.
-Those leaf classes are distributed  in the same manner as PD.
-In that sense, the spatial configuration of the unfolded version of PD, which may be obtained after discovering all the classes and subclasses, is similar to that of  FD, as shown on Figures~\ref{figXP2}, \ref{figXP1} and \ref{figSSF}.
 \end{itemize}
+
+The two interfaces have been designed in order to see whether a progressive top-down display of the hierarchical structure helps the user explore the dataset.
 
@@
 \subsection{Objective}
-During this test, three interfaces are compared:
-\begin{itemize}
-\item PD, which provides a visible hierarchical organization of semantic classes;
-\item FD, which provides a non-visible hierarchical organization of semantic classes;
-\item an Acoustic Display (AD) providing a 2D representation based on acoustic descriptors.
-In this case, the spatial configuration is computed by 1) describing the sounds with mel-frequency cepstrum coefficients (MFCCs) and 2) using  a non metric multidimensional scaling with Kruskal's normalized stress to compute sound positions in a 2D space.
-An example of AD can be seen on~Figure \ref{figXP3}.
-\end{itemize}
 
-By comparing PD and FD, the effect of a visible hierarchy on the user is investigated.
-The goal is to check if forcing the user to browse the high levels of the hierarchy helps him to understand and learn the spatial configuration and the organisation of the sound classes.
-By confronting PD and FD with AD, the relative efficiencies of semantic based and acoustic based spatial configurations are compared.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[scale=0.18]{gfx/XP3.pdf} 
-\end{center}
-\caption{\label{figXP3} The Acoustical Display (AD) computed using  a non metric multidimensional scaling on  MFCCs based acoustic descriptors}
-\end{figure}
+During this experiment, the three displays are compared.
+By comparing AD and xSD (the two semantic displays), the goal is to evaluate the  relative efficiencies of semantic based and acoustic based spatial configurations.
+By comparing FSD and PSD, we study the impact of an enforced hierarchical exploration of the corpus by the user.
+The goal is to check if constraining the user to browse the high levels of the hierarchy helps him to grasp and memorize the spatial configuration and the organization of the sound classes.
 
 \subsection{Experimental protocol}
-We choose to test and compare the three displays through a crowdsourcing experiment. Here is the link to access the experiment web page \footnote{http://217.70.189.118/soundthings/speedSoundFinding/}.
 
-Subjects are asked to successively retrieve 13 target sounds in a dataset of 149 urban environmental sound events.
-The target sounds are distributed such as there are at least two target sounds in each top-level class of the semantic hierarchical structure of the dataset.
-To minimize order effects, target sounds are randomly presented to each subject.
+We evaluate and compare those three displays with a crowdsourcing web-based experiment \footnote{The test is available at \url{http://217.70.189.118/soundthings/speedSoundFinding}} for which we adopt a between subject approach, \textit{i.e.} a subject can only test one display.
+In this experiment, subjects are asked to successively retrieve 13 target sounds among the 149 of the whole dataset.
+The target sounds are selected such as there are at least two target sounds in each top-level class of the semantic hierarchical structure of the dataset.
+To reduce order effects, target sounds are randomly presented to each subject.
 
-To listen to a target sound, the subject has to click a \textit{Play target sound} button.
-Subject may replay the target sound  as many times as they like.
-A timer is started when the subject clicks on a circle for the first time.
-
+First, the subject click a \textit{Play target sound} button to listen to the target sound.
+Then, the subject may replay the target sound  as many times as he likes.
+A timer starts when the subject clicks on a circle for the first time.
 When the target sound is found, subject puts an end to its search by clicking on the \textit{Click if found} button.
 This action 1) stops the timer and 2) loads a new target sound.
-If the subject does not find the correct target sound, an error message appears, and the experiment continues.
+If the subject does not find the correct target sound, an error message appears, and the experiment goes on with the same target sound.
 
-During the experiment, two indications are communicated to the subject:
+During the experiment, two visual cues are shown to the subject:
 \begin{itemize}
-\item \textit{The research state}: "pause" if the subject is currently not looking for a target sound (at the beginning of the experiment, or between two target sounds);  "in progress"  if the subject is currently looking for a target sound.
+\item \textit{The search state}: which can be "pause" if the subject is currently not looking for a target sound (at the beginning of the experiment, or between two target sounds) or  "in progress"  if the subject is currently looking for a target sound.
 \item \textit{Remaining target sounds}: the number of target sounds which remain to find.
@@
 The experiment ends when all the target sounds have been found.
-
-It is most important to note that PD do not pack up at each target sound search. When a circle is revealed, it remains visible during the whole experiment.
+One shall notice that the PSD display is not reintialized at each change of the target sound.
+Thus, when a circle is unfolded, it remains visible during the whole experiment.
 
 \subsection{Data Collection}
-Three sets of data are collected during the experiment:
+Four types of data are collected during the experiment:
 \begin{itemize}
@@
 \subsection{Apparatus}
-A crowd sourcing approach has been adopted.
-The experiment was designed to be supported by the  \textit{chrome-browser} web navigator.
-The link to the experiment has been sent to the subjects via three  mailing list being \textit{music-ir, auditory} and \textit{uuu-IRCAM} (internal IRCAM mailing list).
-Subjects were allowed to perform the experiment once, and on one interface only.
+
+A crowdsourcing approach is adopted.
+The experiment is designed to be performed using the  \textit{chrome} web browser.
+The link to the experiment has been sent to three  mailing lists, namely \textit{Music-Ir, Auditory} and an internal IRCAM mailing list.
+Subjects are only allowed to perform the experiment once, and on one interface only.
 Data were automatically collected server-side at the end of the experiment.
 Subjects were asked to use headphones.
-All the presented sounds were normalized to the same root mean square (RMS) level.
+All sounds of the dataset are normalized to the same root mean square (RMS) level.
 
 \subsection{Participants}
+
 60 subjects have completed the experiment, 20 for each interface.
@@
 \subsection{Statistical analysis}
+
 We use one-way ANOVA and two samples \textit{t} tests at the 5\% significance level to test for statistical significance.
@@
 
-Outlier detection is an important step of any crowdsourcing experiment as experimenters do not control the experimental environment in which the subjects perform the experiment \cite{komarov2013crowdsourcing}\cite{buchholz2011crowdsourcing}.
-A widely used method to detect outlier in  human-computer interaction studies is to consider as outlier an observation  which deviates of at least $\pm2$ standard deviation ($STD$) from the average \cite{komarov2013crowdsourcing}.
-As this method is not robust to the presence of isolated extreme observations (as it is often the case for crowdsourcing experiment), we follow the  method proposed by proposed by \cite{komarov2013crowdsourcing} and used the inter-quartile range ($IQR$).
+Outlier detection is an important step of any crowdsourcing experiment as experimenters do not control the environment in which the subjects perform the experiment \cite{komarov2013crowdsourcing,buchholz2011crowdsourcing}.
+A commonly used method to detect outliers in  human-computer interaction studies is to consider as outlier an observation  which deviates of at least $\pm2$ standard deviation ($STD$) from the average \cite{komarov2013crowdsourcing}.
+though, as this latter method is not robust to the presence of isolated extreme observations (as it is often the case for crowdsourcing experiment), we follow the  method proposed by proposed by \cite{komarov2013crowdsourcing} and used the inter-quartile range ($IQR$).
 With this approach, an observation is considered to be an outlier if its value is more than $3*IQR$ higher than the third quartile or more than $3*IQR$ lesser than the first quartile.
-This methods is applied to the four following criterion :
+This methods is applied in this study to the four following criteria :
 
@@
 \item the \textit{absolute duration}: to detect subjects who took abnormally long breaks between searches.
-\item the \textit{total search duration}: to detect subjects who took an abnormally long time to complete the experiment.
+\item the \textit{total search duration}: to detect subjects who spent an abnormally long time to complete the experiment.
 \item number of heard sounds: to detect subjects who had to hear an abnormally high number of sounds to complete the experiment.
@@
 2 subjects used PD and 3 subjects AD.
-The tables~\ref{tab1} show what criterion are use to detect outliers.
+The tables~\ref{tab1} show the criteria used to detect the 5 outliers.
 3 subjects are detected by observing the \textit{absolute duration}.
@@
 \begin{center}
-\includegraphics[scale=0.4]{gfx/analyse1.eps} 
+\includegraphics[scale=0.4]{gfx/analyse1} 
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.4]{gfx/analyse2.eps} 
+\includegraphics[scale=0.4]{gfx/analyse2} 
 \end{center}
-\caption{\label{fig2} Medians of (top) the numbers of heard sounds at each target sound search, (middle) the numbers of heard sounds without duplication (WD) at each target sound search, (bottom) the relative durations at each target sound search }
+\caption{\label{fig2} Medians of (top) the numbers of heard sounds at each target sound search, (middle) the numbers of heard sounds without duplication (WD) at each target sound search, (bottom) the relative durations at each target sound search \gregoire{polynomial beautiful regression} }
 \end{figure}
 
-\subsection{Interface efficiencies}
+\subsection{Efficiency}
 
-To characterize the interface efficiencies, three set of collected data are assessed:
+To characterize the relative efficiency of the three displays, three metrics are considered:
 
@@
 \item the \textit{total search duration}
-\item the number of heard sounds 
-\item the number of heard sounds without duplication.
-By "without duplication" we mean that, if a same sound prototype is heard 10 times during the 13 searches of the  experiment, it counts only for one.
+\item the number of sounds heard \mathias{on met heard avant ou apres sounds ?} 
+\item the number of unique sounds heard. By "unique" we mean that, if a same sound is heard 10 times during the 13 searches of the  experiment, it counts only for one. 
 \end{itemize}
 
-The two first data help us qualify the notion of efficiency by considering the time and the number of clicks needed to achieve the task (ie.
-reach the target).
+The first two metrics quantify the notion of efficiency by considering the time and the number of clicks needed to achieve the task, \textit{ie.} reach the target.
 The goal for those values is to be as low as possible.
 The third data allows us to measure the selectivity of the interfaces.
-A low number of heard sounds without duplication indicates that subjects understood the spatial organisation of the dataset, and use this knowledge to improve their searches.
+A low number of unique heard sounds indicates that subjects understood the spatial organisation of the dataset, and use this knowledge to improve their searches.
 In contrary, a high number of heard sounds without duplication suggest that the subject did not understood the way circles are organized in space, and tends to play all the sounds at each search.
-The maximum number of heard sounds without duplication is the corpus size: 149 sounds.
+The maximum number of unique heard sounds is the corpus size: 149 sounds.
 
-The type of interface have a significant effect on the the \textit{total search durations} ($F[2,52]=3.63; p=0.03$), the number of heard sounds ($F[2,52]=5.65; p<0.01$) and the number of heard sounds without duplication ($F[2,52]=8.65; p<0.01$).
+The averages and standard deviations for the three metrics are presented on Figure \ref{fig1}.
+The type of display have a significant effect on the the \textit{total search duration} ($F[2,52]=3.63; p=0.03$), the number of heard sounds ($F[2,52]=5.65; p<0.01$) and the number of unique heard sounds ($F[2,52]=8.65; p<0.01$).
+Post hoc analysis on the total search time (Figure~\ref{fig1}~(a)) shows that FSD  perform better  than the other interfaces (FSD-PSD: $p=0.04$; FSD-AD: $p=0.02$), whereas PSD and AD seem to have similar results (PSD-AD: $p=0.88$).
 
-Figure~\ref{fig1}~(a) shows the average and standard deviations of the \textit{total search durations} for the three interfaces.
-FD seems to  perform better  than the other interfaces (FD-PD: $p=0.04$; FD-AD: $p=0.02$), whereas PD and AD seem to have similar results (PD-AD: $p=0.88$).
+Similar results are found for the numbers of heard sounds Figure~\ref{fig1}~(b).
+FSD significantly outperforms the other interfaces (FSD-PSD: $p=0.02$; FSD-ASD: $p<0.01$), whereas PSD and AD show similar outcomes (PSD-AD: $p=0.42$).
 
-We found similar results for the numbers of heard sounds (Figure~\ref{fig1}~(b)).
-FD significantly outperforms the other interfaces (FD-PD: $p=0.02$; FD-AD: $p<0.01$), whereas PD and AD show similar outcomes (PD-AD: $p=0.42$).
-
-Figure~\ref{fig1}~(c) shows the results for the number of heard sounds without duplication.
+Figure~\ref{fig1}~(c) shows the results for the number of unique heard sounds.
 This time the results of AD are significantly lower than those of both PD and FD (AD-PD: $p<0.01$; AD-FD: $p<0.01$).
@@
 
-According to those results, a hierarchical organization of the dataset based on semantic values (PD and FD) allows users to retrieve the 13 target sounds 1) quicker, and 2) by listening to a smaller amount of sounds than an organization based on acoustic descriptors (AD).
-But those two effects are significantly compromised when users have to parse the entire hierarchy to reach the first target sound, as it the case for PD.
-It seems that imposing a graphical representation of the hierarchy disturbs or confuses the user instead of allowing him to learn the spatial organization of the classes.
+According to those results, a hierarchical organization of the dataset based on semantic values (PSD and FSD) allows the users to retrieve the 13 target sounds 1) more efficiently, and 2) by listening to a smaller amount of sounds than an organization based on acoustic descriptors (AD).
+However, those two effects are significantly compromised when users have to parse the entire hierarchy to reach the first target sound, as in the PSD display.
+It seems that enforcing an explicit exploration of the hierarchy disturbs or confuses the user instead of allowing him to learn the semantically motivated spatial organization of the classes.
 
@@
 We now study if and how users  progressively acquire knowledge about the spatial organization of the classes.
-To do that, variations of the data over the searches are assessed.
+To do that, the evolution of the above described metrics with respect to the search indexes is considered.
 Three sets of collected data are used:
@@
 \item the number of heard sounds for each target sound search 
-\item the number of heard sounds for each target sound search without duplication. 
+\item the number of unique heard sounds for each target sound search. 
 \end{itemize}
 
-Figure~\ref{fig2}~(a) shows the evolution of the average durations of each target sound search observed for PD, FD and AD.
-It is interesting to note that both for PD and FD, the maximum value is observed for the first search, whereas it is observed for the fourth search for AD.
-Moreover if the curve profiles of PD and FD  seem to progressively decrease and are very similar, the one of AD is much more irregular.
-Similar results are found observing the evolution of the average numbers of heard sounds with and without duplication (Figures~\ref{fig2}~(b) and (c))
-
-These results tend to indicate that PD and FD facilitate the learning of the spatial configuration, as the search durations and the numbers of heard sounds at each search decrease over time.
-Although curves for PD and FD have similar profiles, FD seem to better perform as users of FD where able to find the target sounds faster by clicking on fewer circles.
+Figure~\ref{fig2}~(a) shows the evolution of the average durations of each target sound search observed for AD, FSD and PSD.
+As shown by a linear regression of the data, there is an overall increase of efficiency with respect to the index of the search.
+For the duration and the number of heard sounds, the AD and PSD perform equivalently, whereas the FSD exhibit better performance.
+In the case of the number of unique heard sounds, the two semantic displays are equivalent.
+This latter result suggest that for both semantic displays, users have parsed the same range of the dataset.
+Although users of FSD have done so by listening to fewer sounds and more quickly.
 
@@
 Second, an imposed visualisation of the semantic hierarchical structure of the dataset does not help user to understand and learn the spatial configuration of the semantic class, but instead disturbs the navigation.
+
+In the dataset considered in this study, there are as many leaf classes as sounds in the dataset, which would be unrealistic for large datasets.
+Though the organization can easily be adapted by considering the leaf classes as \textit{collections} of semantically similar sound samples.
+Thus, two sounds of \textit{male-yelling} would be grouped into a  single leaf class with the tag \textit{male-yelling}.
+The leaf class would then also have a \textit{prototype sound} being the most representative item of the different \textit{male-yelling} sounds belonging to the leaf class.
 

### 2016-02-11 lafaySsf2016.tex
newversion

@@
 \end{center}
-\caption{\label{figSSF} Spatial configuration of the Progressive Display (PD) based on the semantic hierarchical structure of the dataset}
+\caption{\label{figSSF} Spatial configuration of the Progressive Semantic Display (PSD) based on the semantic hierarchical structure of the dataset}
 \end{figure}
@@
 \end{center}
-\caption{\label{figXP3} Acoustical Display (AD) \gregoire{fill the dots ;)}}
+\caption{\label{figXP3} Acoustical Display (AD)}
 \end{figure}
@@
 \begin{center}
-\includegraphics[scale=0.4]{gfx/analyse1} 
+\includegraphics[scale=0.4]{gfx/analyse1.eps} 
 \end{center}
-\caption{\label{fig1} Average and standard deviation for (a) the total search durations, (b) the total search durations and (c) the total search durations}
+\caption{\label{fig1} Average and standard deviation for (a) the total search durations, (b) the number of heard sounds and (c) the number of unique heard sounds}
 \end{figure}
@@
 \begin{center}
-\includegraphics[scale=0.4]{gfx/analyse2} 
+\includegraphics[scale=0.4]{gfx/analyse2.eps} 
 \end{center}
-\caption{\label{fig2} Medians of (top) the numbers of heard sounds at each target sound search, (middle) the numbers of heard sounds without duplication (WD) at each target sound search, (bottom) the relative durations at each target sound search \gregoire{polynomial beautiful regression} }
+\caption{\label{fig2} evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) average the numbers of unique heard sounds at each target sound search.
+Lines are linear regression.}
 \end{figure}
@@
 This time the results of AD are significantly lower than those of both PD and FD (AD-PD: $p<0.01$; AD-FD: $p<0.01$).
-For AD, 75\% of the subjects heard more than 138 sounds, and 25\% heard 148 sounds, that is almost the entire database.
-Considering PD, 75\% of subjects heard  less than 133 sounds, against 144 for FD.
-This time PD and FP seem to perform equally (PD-FD: $p=0.58$)
+For AD, 75\% of the subjects heard more than 140 sounds, and 25\% heard at least 148 sounds, that is almost the entire database.
+Considering PSD, 75\% of subjects heard  less than 134 sounds, against 143 for FSD.
+This time PSD and FSD seem to perform equally (PD-FD: $p=0.58$)
 

### 2016-02-11 lafaySsf2016.tex
remove useless fig and add new XP3.fig
add new xp link

@@
 \begin{center}
-\includegraphics[scale=0.18]{gfx/XP3.pdf} 
+\includegraphics[scale=0.18]{gfx/xp3.png} 
 \end{center}
@@
 
-We evaluate and compare those three displays with a crowdsourcing web-based experiment \footnote{The test is available at \url{http://217.70.189.118/soundthings/speedSoundFinding}} for which we adopt a between subject approach, \textit{i.e.} a subject can only test one display.
+We evaluate and compare those three displays with a crowdsourcing web-based experiment \footnote{The test is available at \url{http://soundthings.org/research/speedSoundFinding}} for which we adopt a between subject approach, \textit{i.e.} a subject can only test one display.
 In this experiment, subjects are asked to successively retrieve 13 target sounds among the 149 of the whole dataset.

### 2016-02-12 lafaySsf2016.tex
new figure
second round of typo fixing

@@
 ~In this paper, we study the relevance of a semantic organization of sounds for easing the browsing of a sound database.
-For such task, the semantic organization is traditionally hold thanks to a keyword selection process.
-However, various issues of written language like word polysemy, ambiguities translation issues may bias the browsing process.
+For such task, the semantic organization is traditionally held thanks to a keyword selection process.
+However, various issues with written language like word polysemy, ambiguities translation issues may bias the browsing process.
 
-We consider here two display of sounds that are organized by considering underlying semantic information that draws a hierarchy.
+We consider here displays of sounds that are organized by considering the underlying semantic information that draws a hierarchy.
 For the sake of comparison, we also consider a display whose organization of sound in the plane is based on acoustic cues.
 Those three displays are evaluated in terms of search speed in a crowd sourcing experiment.
-Results
+Results demonstrate the usefulness of the use of implicit semantic organization to display sounds both in terms of search speed and efficiency of learning.
 \end{abstract}
@@
 
-With the growing capability of recording and storage, the problem of indexing large databases of audio has recently been the object of much attention \cite{Wold1996}.
+With the growing capability of recording and storage devices, the problem of indexing large databases of audio has recently been the object of much attention \cite{Wold1996}.
 Most of that effort is dedicated to the automatic inference of indexing metadata from the actual audio recording \cite{Zhang1999, tzanetakis2002musical}; in contrast, the ability to browse such databases in an effective manner has been less considered.
@@
 Most media assets management are based on keyword-driven queries.
-The user enters a word which characterizes the desired item, and the interface presents him with items related to this word.
+The user enters a word which best characterizes the desired item, and the interface presents him with items related to this word.
 The effectiveness of this principle is primarily based on the typological structure and nomenclature of the database.
-However, for databases and more specifically for databases of sounds, issues arise:
+However, for databases and more specifically for databases of sounds, several issues arise:
 
@@
 sources (a car door), as well as by the action of those sources (the slamming of a car door) or their environments (slamming a car door in a garage) \cite{houix2012lexical, niessen2010categories, brown2011towards}.
-Designing an effective keyword-based search system requires an accurate description of each sound, which has to be adaptable to the sound representation of each user.
+Designing an effective keyword-based search system requires an accurate description of each sound, which has to be tunable to the sound representation of each user.
 \item Pre-defined verbal descriptions of the sounds made available to the users may potentially bias their browsing and final selection.
@@
 \end{enumerate}
-
 To circumvent those issues, not relying on keywords is desirable. That said, semantics, which is traditionally conveyed through written text, may be important for easing browsing.
 
-To study this issue, we consider in this paper several means of displaying sounds without relying on any textual representation.
-One, considered as a reference baseline, do not rely on any semantic information as the sounds are organized according to their acoustical properties, \textit{i.e.} time averaged spectral features.
-The second one
+Thus, we consider in this paper several means of displaying sounds without relying on any textual representation.
+The first one, considered as a reference baseline, do not rely on any semantic information as the sounds are organized according to their acoustical properties, \textit{i.e.} time averaged spectral features.
+The second and third displays have their spatial organization based on the on a predefined hierarchical semantic organization.
+And those last two displays differs by huw the semantic is used.
 
-Their effectiveness is studied with a search-based task whose aim is find a target sound by browsing the database using the display under evaluation.
+Their effectiveness is studied with a search-based task whose aim is to find a target sound by browsing the database using the display under evaluation.
 
@@
 The crowd sourcing test used to compare those displays is presented in Section \ref{test}.
-the outcomes of this experiment are discussed in Section \ref{results}.
+The outcomes of this experiment are discussed in Section \ref{results}.
 
@@
 
-Most of the research effort in Sound Design and Music Information Retrieval (MIR) communities is focused on acoustical based indexing and browsing \cite{tzanetakis2003musescape, streich2008music}.
+Most of the research effort in sound design and Music Information Retrieval (MIR) communities is focused on acoustical based indexing and browsing \cite{tzanetakis2003musescape, streich2008music}.
 Typically, the items (sound effects or pieces of music) are modeled by processing the digital audio waveform using some signal processing pipeline in order to get a compact description for each  items
 \cite{coleman2007mused}.
-Then, statistical projections or embedding techniques, like Principal components analysis (PCA),  Multi Dimensional Scaling (MDS) \cite{schwarz2009sound,Cano2002}, Self Organizing Maps (SOM) \cite{pampalk2004exploring} and the like are used to project the items in a two or three dimensional space while preserving as much as possible distances among items.
-One of the advantage of such an approach is its ability to scale to very large databases \cite{schwarz2009scalability} as it does not need any kind of manual annotations and allows to search by similarity efficiently according to acoustical properties.
+Then, statistical projections or embedding techniques, like Principal Components Analysis (PCA),  Multi Dimensional Scaling (MDS) \cite{schwarz2009sound,Cano2002}, Self Organizing Maps (SOM) \cite{pampalk2004exploring} and the like are used to project the items in a two or three dimensional space while preserving as much as possible distances among items.
+One of the advantage of such an approach is its ability to scale to very large databases \cite{schwarz2009scalability} as it does not need any kind of manual annotations and allows the user to search by similarity efficiently according to acoustical properties.
 
@@
 Selecting the most relevant features to achieve the correct projection of the data may only be performed by an expert user.
-If done \textit{a priori} by an expert or by some above cited dimensionality reduction technique, the induced bias can strongly limit the user in its ability to .
-In that respect, semantic tags, if available for the data at hand have the advantage of implicitly structuring the similarity space, thus eventually easing the browsing process even if the actual tags are not -- as in this study -- exposed to the user.
+If done \textit{a priori} by an expert or by some above cited dimensionality reduction technique, the induced bias can strongly limit the user in its ability to access the data searched for.
+In that respect, semantic tags, if available for the data at hand have the advantage of implicitly structuring the similarity space, thus eventually easing the browsing process even if the actual tags are not -- as in this study -- explicitly exposed to the user.
 
@@
 \end{center}
-\caption{\label{figSSF} Spatial configuration of the Progressive Semantic Display (PSD) based on the semantic hierarchical structure of the dataset}
+\caption{\label{figSSF} Semantic displays are  based on the semantic hierarchical structure of the dataset}
 \end{figure}
@@
 
-In order for the semantic hierarchical structure to be perceptually valid, the \textit{tags} describing the classes  are chosen from sound categories presented in studies addressing environmental auditory scenes perception \cite{niessen2010categories, brown2011towards, dubois2006cognitive}.
+In order for the semantic hierarchical structure to be perceptually valid, the \textit{tags} that describe the classes  are chosen from sound categories presented in studies addressing environmental auditory scenes perception \cite{niessen2010categories, brown2011towards, dubois2006cognitive}.
 In cognitive psychology, sound categories may be regarded as intermediaries between collective sound representations and individual sensory experiences \cite{dubois2006cognitive}.
@@
 
+The aim of the displays described in this section is to allow the user to efficiently browse the above presented dataset of sounds explore without any written textual help.
+For each display, a sound is graphically represented by a filled circle whose, once clicked, plays the actual sound.
+The spatial organization is specific to each display.
+
+As a reference, an Acoustic Display (AD) provides a spatial representation based on acoustic descriptors.
+For describing each sound, Mel-Frequency Cepstrum Coefficients (MFCCs) are computed with standard parameter settings (13 lowest quefrency coefficients are kept out of 40).
+The Euclidean distance between time averaged features is then computed between each sounds.
+A non metric multidimensional scaling with Kruskal's normalized stress is then considered to project the data into two dimensions according to this distance, see Figure \ref{figXP3}.
+
 \begin{figure}[t]
@@
 \end{figure}
+
+Two semantically oriented displays are then proposed, called respectively Semantic Display (FSD) and Progressive Semantic Display (PSD), respecively shown on Figures~\ref{figXP2} and \ref{figXP1}.
+Both display  consider the hierarchical structure of the dataset to organize sounds.
+Each sound class is represented by a filled circle.
+Circles are then packed together according to the hierarchical semantic organization of the dataset, as shown on Figure \ref{figSSF}.
+Thus, subclasses belonging to the same class are close to each others.
+Circle packing functions of the D3.js (Data-Driven Documents) javascript library \cite{2011-d3} are used to distribute the sound classes in the space.
+Depending on the display, the user can whether acces directly to each leaf class or has to go through intermediate levels of the hierarchy.
+
+More precisely with the FSD display, users can directly visualize the whole hierarchy, down to the leaf classes.
+On contrary, with the PSD display, users have access to the intermediate semantic levels of the hierarchy.
+Upon first using PD, they observe circles representing the top classes of the semantic hierarchical structure of the dataset.
+When users click on a circle, they hear the sound prototype of the class and the subclasses are progressively revealed, represented by white circles contained in the circle that has just been clicked.
+The same action repeats itself until the user reaches the leaf classes of the hierarchy.
+The leaf classes are represented with small gray circles, indicating that there is no subclass to discover.
+Thus the PD has a constrained exploration system.
+When a user click on a circle, sub-circles are  automatically revealed to him in a gradual way.
+Each time a sub-circle is automatically revealed, its sound prototype is played.
+Users may stop the discovery process by clicking on an other circle.
+In this display, the leaf classes are distributed  in the same manner as FSD.
+Thus, the spatial configuration of the unfolded version of PSD, which may be obtained after discovering all the classes and subclasses, is equivalent to the one of FSD, as shown on Figures~\ref{figXP2}, \ref{figXP1} and \ref{figSSF}.
+
+The latter interface is introduced in order to study if a progressive top-down display of the hierarchical structure helps the user explore and understand the dataset hierarchy.
 
@@
 
-The aim of the displays presented in this section is to allow the user to efficiently browse the above presented dataset of sounds explore without any written textual help.
-For each display, a sound is graphically represented by a filled circle whose, once clicked, plays the actual sound.
-The spatial organization is specific to each display.
-
-As a reference, an Acoustic Display (AD) provides a spatial representation based on acoustic descriptors.
-For describing each sound, mel-frequency cepstrum coefficients (MFCCs) are computed with standard parameter settings (13 lowest quefrency coefficients are kept out of 40).
-The Euclidean distance between time averaged features is then computed between each sounds.
-A non metric multidimensional scaling with Kruskal's normalized stress is then considered to project the data into two dimensions \ref{figXP3} according to this distance.
-
-Two semantically oriented displays are then proposed, called respectively Semantic Display (FSD) and Progressive Semantic Display (PSD), respecively shown on Figures~\ref{figXP2} and \ref{figXP1}.
-Both display  consider the hierarchical structure of the dataset to organize sounds.
-Each sound class is represented by a filled circle.
-Circles are then packed together according to the hierarchical semantic organization of the dataset, as shown on Figure \ref{figSSF}.
-Thus, subclasses belonging to the same class are close to each others.
-Circle packing functions of the D3.js (Data-Driven Documents) javascript library \cite{2011-d3} are used to distribute the sound classes in the space.
-Depending on the display, the user can whether acces directly to each leaf class or has to go through intermediate levels of the hierarchy.
-More precisely: \\
-\begin{itemize}
-\item \textit{FSD}: users can directly visualize the whole hierarchy, down to the leaf classes.
-Those leaf classes are distributed  in the same manner as PD.
-In that sense, the spatial configuration of the unfolded version of PD, which may be obtained after discovering all the classes and subclasses, is similar to that of  FD, as shown on Figures~\ref{figXP2}, \ref{figXP1} and \ref{figSSF}.
-\item \textit{PSD}: users have access to the intermediate semantic levels of the hierarchy.
-Upon first using PD, they observe circles representing the top classes of the semantic hierarchical structure of the dataset.
-When users click on a circle, they hear the sound prototype of the class and the subclasses are progressively revealed, represented by white circles contained in the circle that has just been clicked.
-The same action repeats itself until the user reaches the leaf classes of the hierarchy.
-The leaf classes are represented with small gray circles, indicating that there is no subclass to discover.
-Thus the PD has a constrained exploration system.
-When a user click on a circle, sub-circles are  automatically revealed to him in a gradual way.
-Each time a sub-circle is automatically revealed, its sound prototype is played.
-Users may stop the discovery process by clicking on an other circle.
-\end{itemize}
-
-The two interfaces have been designed in order to see whether a progressive top-down display of the hierarchical structure helps the user explore the dataset.
-
 \section{Experiment} \label{test}
@@
 By comparing FSD and PSD, we study the impact of an enforced hierarchical exploration of the corpus by the user.
-The goal is to check if constraining the user to browse the high levels of the hierarchy helps him to grasp and memorize the spatial configuration and the organization of the sound classes.
+The goal is to check if constraining the user to first browse the high levels of the hierarchy helps him to grasp and memorize the spatial configuration and the organization of the sound classes.
 
@@
 
-We evaluate and compare those three displays with a crowdsourcing web-based experiment \footnote{The test is available at \url{http://soundthings.org/research/speedSoundFinding}} for which we adopt a between subject approach, \textit{i.e.} a subject can only test one display.
+We evaluate and compare those three displays with a crowdsourcing web-based experiment \footnote{The test is available at \url{http://soundthings.org/research/speedSoundFinding/index.html}, each of the 3 displays is also available by appending a digit from 1 to 3 to \textit{index} in the above cited url.} for which we adopt a between subject approach, \textit{i.e.} a subject can only test one display.
 In this experiment, subjects are asked to successively retrieve 13 target sounds among the 149 of the whole dataset.
@@
 A timer starts when the subject clicks on a circle for the first time.
-When the target sound is found, subject puts an end to its search by clicking on the \textit{Click if found} button.
+When the target sound is found, subject puts an end to the search by clicking on the \textit{Click if found} button.
 This action 1) stops the timer and 2) loads a new target sound.
@@
 Outlier detection is an important step of any crowdsourcing experiment as experimenters do not control the environment in which the subjects perform the experiment \cite{komarov2013crowdsourcing,buchholz2011crowdsourcing}.
-A commonly used method to detect outliers in  human-computer interaction studies is to consider as outlier an observation  which deviates of at least $\pm2$ standard deviation ($STD$) from the average \cite{komarov2013crowdsourcing}.
-though, as this latter method is not robust to the presence of isolated extreme observations (as it is often the case for crowdsourcing experiment), we follow the  method proposed by proposed by \cite{komarov2013crowdsourcing} and used the inter-quartile range ($IQR$).
+A commonly used method to detect outliers in  human-computer interaction studies is to consider as outlier an observation  which deviates of at least $\pm2$ standard deviation from the average \cite{komarov2013crowdsourcing}.
+though, as this latter method is not robust to the presence of isolated extreme observations (as it is often the case for crowdsourcing experiment), we follow the  method proposed by \cite{komarov2013crowdsourcing} and used the Inter-Quartile Range ($IQR$).
 With this approach, an observation is considered to be an outlier if its value is more than $3*IQR$ higher than the third quartile or more than $3*IQR$ lesser than the first quartile.
@@
 The tables~\ref{tab1} show the criteria used to detect the 5 outliers.
-3 subjects are detected by observing the \textit{absolute duration}.
+3 subjects are detected by considering the \textit{absolute duration}.
 They spent respectively 50 minutes, 5 hours and 17.5 hours to complete the experiment.
-1 subject is detected by observing the total search duration (46 minutes), and an other by observing the total number of heard sounds (1800 heard sounds, roughly 12 times the total size of the corpus).
+1 subject is detected by considering the total search duration (46 minutes), and an other by observing the total number of heard sounds (1800 heard sounds, roughly 12 times the total size of the corpus).
 And lastly 2 subjects are detected by observing the maximum number of times they had to hear a target sounds before finding it (4 and 8 times).
@@
 
-To characterize the relative efficiency of the three displays, three metrics are considered:
+To study the relative efficiency of the three displays, three metrics are considered:
 
@@
 A low number of unique heard sounds indicates that subjects understood the spatial organisation of the dataset, and use this knowledge to improve their searches.
-In contrary, a high number of heard sounds without duplication suggest that the subject did not understood the way circles are organized in space, and tends to play all the sounds at each search.
+In contrary, a high number of heard sounds without duplication suggest that the subject did not understood the way sounds are organized in space, and tends to play all the sounds at each search.
 The maximum number of unique heard sounds is the corpus size: 149 sounds.
@@
 
-Similar results are found for the numbers of heard sounds Figure~\ref{fig1}~(b).
+Similar results are found for the numbers of heard sounds (Figure~\ref{fig1}~(b)).
 FSD significantly outperforms the other interfaces (FSD-PSD: $p=0.02$; FSD-ASD: $p<0.01$), whereas PSD and AD show similar outcomes (PSD-AD: $p=0.42$).
@@
 Considering PSD, 75\% of subjects heard  less than 134 sounds, against 143 for FSD.
-This time PSD and FSD seem to perform equally (PD-FD: $p=0.58$)
+This time PSD and FSD perform equivalently (PD-FD: $p=0.58$).
 
@@
 In the case of the number of unique heard sounds, the two semantic displays are equivalent.
-This latter result suggest that for both semantic displays, users have parsed the same range of the dataset.
-Although users of FSD have done so by listening to fewer sounds and more quickly.
+This latter result suggest that for both semantic displays, users have parsed the same range of the dataset, although users of FSD have done so by listening to fewer sounds and more quickly.
 
@@
 
-In this paper, two displays allowing users to explore a sound dataset without written textual help are presented.
+In this paper, two displays allowing users to explore a semantically organized sound dataset without written textual help are presented.
 The interfaces distribute sounds represented by circles on a 2D space.
 The spatial organisation is driven by semantic features.
-The two graphical displays are assessed and compared to a third listening based interface in which spatial configuration depends upon acoustic features.
+Those two semantic displays are assessed and compared to a third display in which spatial configuration depends upon acoustic features.
 The tests consist in data retrieval tasks.

### 2016-02-13 lafaySsf2016.tex
Relu et corrigé, un peu reformulé. Restent deux questions en attente (cf. issues).

@@
 \begin{abstract}
-~In this paper, we study the relevance of a semantic organization of sounds for easing the browsing of a sound database.
-For such task, the semantic organization is traditionally held thanks to a keyword selection process.
-However, various issues with written language like word polysemy, ambiguities translation issues may bias the browsing process.
+~In this paper, we study the relevance of a semantic organization of sounds to ease the browsing of a sound database.
+For such a task, semantic access to data is traditionally implemented by a keyword selection process.
+However, various limitations of written language, such as word polysemy, ambiguities, or translation issues, may bias the browsing process.
 
-We consider here displays of sounds that are organized by considering the underlying semantic information that draws a hierarchy.
-For the sake of comparison, we also consider a display whose organization of sound in the plane is based on acoustic cues.
-Those three displays are evaluated in terms of search speed in a crowd sourcing experiment.
-Results demonstrate the usefulness of the use of implicit semantic organization to display sounds both in terms of search speed and efficiency of learning.
+We present and study the efficiency of two sound presentation strategies that organize sounds spatially so as to reflect an underlying semantic hierarchy.
+For the sake of comparison, we also consider a display whose spatial organization is based on acoustic cues.
+Those three displays are evaluated in terms of search speed in a crowdsourcing experiment.
+Results demonstrate the usefulness of using an implicit semantic organization to display sounds, both in terms of search speed and of efficiency of learning.
 \end{abstract}
@@
 \item Pre-defined verbal descriptions of the sounds made available to the users may potentially bias their browsing and final selection.
-\item Localization of the query interface is made difficult as the translation of some words referring to qualitative aspects of the sound such as its ambiance is notoriously ambiguous and subject to cultural specificities.
+\item Localization of the query interface is made difficult as the translation of some words referring to qualitative aspects of the sound, such as its ambience, is notoriously ambiguous and subject to cultural specificities.
 \item Unless considerable time and resources are invested into developing a multilingual interface, any system based on verbal descriptions can only be used with reduced performance by non-native speakers of the chosen language.
 \end{enumerate}
-To circumvent those issues, not relying on keywords is desirable. That said, semantics, which is traditionally conveyed through written text, may be important for easing browsing.
+To circumvent those issues, not relying on keywords is desirable.
+However, conveying semantics remains a necessity for ease of browsing, which raises the question of what alternate means, other than written language, can be used to that end.
 
-Thus, we consider in this paper several means of displaying sounds without relying on any textual representation.
-The first one, considered as a reference baseline, do not rely on any semantic information as the sounds are organized according to their acoustical properties, \textit{i.e.} time averaged spectral features.
-The second and third displays have their spatial organization based on the on a predefined hierarchical semantic organization.
-And those last two displays differs by huw the semantic is used.
+We thus consider in this paper three means of displaying sounds without relying on any textual representation, all based on a spatial organization of sounds.
+The first one, considered as a reference baseline, does not rely on any semantic information and positions sounds according to their acoustical properties (time averaged spectral features).
+The second and third displays exhibit a spatial organization based  on a predefined hierarchical semantic organization; those last two displays differs by how semantics are used.
 
-Their effectiveness is studied with a search-based task whose aim is to find a target sound by browsing the database using the display under evaluation.
+Their effectiveness is studied through a search-based task whose aim is to find a given target sound by browsing the database using the display under evaluation.
 
-The paper is organized as follows: in Section \ref{previous} previous work on the topic of sound database browsing is reviewed.
-Then, the corpus used in this study is described in Section \ref{dataset}.
-the three displays under evaluation are next described in Section \ref{display}.
-The crowd sourcing test used to compare those displays is presented in Section \ref{test}.
-The outcomes of this experiment are discussed in Section \ref{results}.
+The paper is organized as follows: in Section~\ref{previous}, previous work on the topic of sound database browsing is reviewed.
+Then, the corpus used in this study is described in Section~\ref{dataset}.
+The three displays under evaluation are next described in Section~\ref{display}, and the crowdsourcing test used to compare them is presented in Section~\ref{test}.
+Finally, the outcomes of this experiment are discussed in Section~\ref{results}.
 
@@
 
-Most of the research effort in sound design and Music Information Retrieval (MIR) communities is focused on acoustical based indexing and browsing \cite{tzanetakis2003musescape, streich2008music}.
-Typically, the items (sound effects or pieces of music) are modeled by processing the digital audio waveform using some signal processing pipeline in order to get a compact description for each  items
+Most of the research effort in the sound design and Music Information Retrieval (MIR) communities is focused on acoustical based indexing and browsing \cite{tzanetakis2003musescape, streich2008music}.
+Typically, the items (sound effects or pieces of music) are modeled by processing the digital audio waveform through some signal processing pipeline in order to get a compact description of each item
 \cite{coleman2007mused}.
-Then, statistical projections or embedding techniques, like Principal Components Analysis (PCA),  Multi Dimensional Scaling (MDS) \cite{schwarz2009sound,Cano2002}, Self Organizing Maps (SOM) \cite{pampalk2004exploring} and the like are used to project the items in a two or three dimensional space while preserving as much as possible distances among items.
-One of the advantage of such an approach is its ability to scale to very large databases \cite{schwarz2009scalability} as it does not need any kind of manual annotations and allows the user to search by similarity efficiently according to acoustical properties.
+Then, statistical projections or embedding techniques, like Principal Components Analysis (PCA),  Multi Dimensional Scaling (MDS) \cite{schwarz2009sound,Cano2002}, Self Organizing Maps (SOM) \cite{pampalk2004exploring} and the like are used to project the items in a two or three dimensional space while preserving as much as possible the distances among them.
+One of the advantages of such an approach is its ability to scale to very large databases \cite{schwarz2009scalability} as it does not need any kind of manual annotation and allows the user to search by similarity efficiently according to acoustical properties.
 
-Though, such acoustical models are inherently subject to observation noise and biases.
-Selecting the most relevant features to achieve the correct projection of the data may only be performed by an expert user.
-If done \textit{a priori} by an expert or by some above cited dimensionality reduction technique, the induced bias can strongly limit the user in its ability to access the data searched for.
-In that respect, semantic tags, if available for the data at hand have the advantage of implicitly structuring the similarity space, thus eventually easing the browsing process even if the actual tags are not -- as in this study -- explicitly exposed to the user.
+However, acoustical models are inherently subject to observation noise and biases.
+Selecting the most relevant features to achieve the correct projection of the data may only be performed by an expert user, taking into account the specifities of the data.
+If done \textit{a priori} by an expert or by some above cited dimensionality reduction technique, the induced bias can strongly limit the users' ability to access what they search for.
+In that respect, semantic tags, if available, have the advantage of implicitly structuring the similarity space, thus possibly easing the browsing process even if the actual tags are not -- as in this study -- explicitly exposed to the user.
 
@@
 \end{center}
-\caption{\label{figdataset} Semantic hierarchical structure of the dataset of urban environmental sounds}
+\caption{\label{figdataset} Semantic hierarchical structure of the dataset of urban environmental sounds.}
 \end{figure}
@@
 \end{center}
-\caption{\label{figSSF} Semantic displays are  based on the semantic hierarchical structure of the dataset}
+\caption{\label{figSSF} Semantic displays are  based on the semantic hierarchical structure of the dataset.}
 \end{figure}
@@
 The dataset considered in this study is composed of 149 urban environmental sound events.
-It has a semantic organization that is as follows: a sound is characterized by a tag describing the physical source of the sound (\textit{man-yelling}, \textit{car-passing}).
-Sounds are then hierarchically grouped into classes according to their tags (\textit{car} $>$ \textit{car-passing}; car $>$ \textit{car-starting}).
+Its semantic organization is as follows: a sound is characterized by a tag describing the physical source of the sound (\textit{man-yelling}, \textit{car-passing}).
+Sounds are then hierarchically grouped into classes according to their tags (\textit{car} $>$ \textit{car-passing}; \textit{car} $>$ \textit{car-starting}).
 Those classes are in turn packed into classes until high level classes describing broader concepts are reached (\textit{traffic} $>$ \textit{car} $>$ \textit{car-passing}).
@@
 In cognitive psychology, sound categories may be regarded as intermediaries between collective sound representations and individual sensory experiences \cite{dubois2006cognitive}.
-It is our belief that using such category names to build the hierarchical structure makes the latter perceptually motivated, and can thus help the users for efficiently browsing the dataset.
+It is our belief that using such category names to build the hierarchical structure makes it perceptually motivated, and can thus help the users to efficiently browse the dataset.
 
@@
 
-The aim of the displays described in this section is to allow the user to efficiently browse the above presented dataset of sounds explore without any written textual help.
-For each display, a sound is graphically represented by a filled circle whose, once clicked, plays the actual sound.
-The spatial organization is specific to each display.
+The aim of the displays described in this section is to allow the user to efficiently browse the above presented dataset of sounds without any written textual help.
+In all displays, each sound is graphically represented by a filled circle which, once clicked, plays the actual sound.
+The spatial organization of circles is specific to each display.
 
 As a reference, an Acoustic Display (AD) provides a spatial representation based on acoustic descriptors.
-For describing each sound, Mel-Frequency Cepstrum Coefficients (MFCCs) are computed with standard parameter settings (13 lowest quefrency coefficients are kept out of 40).
-The Euclidean distance between time averaged features is then computed between each sounds.
-A non metric multidimensional scaling with Kruskal's normalized stress is then considered to project the data into two dimensions according to this distance, see Figure \ref{figXP3}.
+Each sound is described by time-averaged Mel-Frequency Cepstrum Coefficients (MFCCs) computed with standard parameter settings (13 lowest quefrency coefficients are kept out of 40).
+The Euclidean distance between time averaged features is then computed for each sound pair.
+A non metric multidimensional scaling with Kruskal's normalized stress is then employed to project the data into two dimensions according to this distance, see Figure \ref{figXP3}.
 
@@
 \end{center}
-\caption{\label{figXP3} Acoustical Display (AD)}
+\caption{\label{figXP3} Acoustical Display (AD).}
 \end{figure}
 
-Two semantically oriented displays are then proposed, called respectively Semantic Display (FSD) and Progressive Semantic Display (PSD), respecively shown on Figures~\ref{figXP2} and \ref{figXP1}.
+Two semantically oriented displays are then proposed, called Full Semantic Display (FSD) and Progressive Semantic Display (PSD), shown on Figures~\ref{figXP2} and~\ref{figXP1}, respectively.
 Both display  consider the hierarchical structure of the dataset to organize sounds.
 Each sound class is represented by a filled circle.
-Circles are then packed together according to the hierarchical semantic organization of the dataset, as shown on Figure \ref{figSSF}.
+Circles are then packed together according to the hierarchical semantic organization of the dataset, as shown on Figure~\ref{figSSF}.
 Thus, subclasses belonging to the same class are close to each others.
 Circle packing functions of the D3.js (Data-Driven Documents) javascript library \cite{2011-d3} are used to distribute the sound classes in the space.
-Depending on the display, the user can whether acces directly to each leaf class or has to go through intermediate levels of the hierarchy.
+Depending on the display, the user can either access each leaf class directly, or has to click through intermediate levels of the hierarchy.
 
-More precisely with the FSD display, users can directly visualize the whole hierarchy, down to the leaf classes.
-On contrary, with the PSD display, users have access to the intermediate semantic levels of the hierarchy.
-Upon first using PD, they observe circles representing the top classes of the semantic hierarchical structure of the dataset.
+More precisely, with the FSD display, users can directly visualize the whole hierarchy, down to the leaf classes, while with the PSD display, users only have access, at first, to the intermediate semantic levels of the hierarchy.
+Upon first using PSD, they observe circles representing the top classes of the semantic hierarchical structure of the dataset.
 When users click on a circle, they hear the sound prototype of the class and the subclasses are progressively revealed, represented by white circles contained in the circle that has just been clicked.
 The same action repeats itself until the user reaches the leaf classes of the hierarchy.
-The leaf classes are represented with small gray circles, indicating that there is no subclass to discover.
-Thus the PD has a constrained exploration system.
-When a user click on a circle, sub-circles are  automatically revealed to him in a gradual way.
+The leaf classes are represented with small gray circles, indicating that there is no subclass to uncover.
+Thus the PSD has a constrained exploration system.
+When a user clicks on a circle, sub-circles are  automatically revealed to him in a gradual way.
 Each time a sub-circle is automatically revealed, its sound prototype is played.
@@
 \end{center}
-\caption{\label{figXP2} Full Semantic Display (FSD): implicit hierarchical organization of semantic classes }
+\caption{\label{figXP2} Full Semantic Display (FSD): implicit hierarchical organization of semantic classes.}
 \end{figure}
@@
 \end{center}
-\caption{\label{figXP1} Progressive Semantic Display (PSD):  explicit hierarchical organization of semantic classes. (a) initial folded version; (b) partly folded version; (c) unfolded version}
+\caption{\label{figXP1} Progressive Semantic Display (PSD):  explicit hierarchical organization of semantic classes. (a) initial folded version; (b) partly folded version; (c) unfolded version.}
 \end{figure*}
@@
 In this experiment, subjects are asked to successively retrieve 13 target sounds among the 149 of the whole dataset.
-The target sounds are selected such as there are at least two target sounds in each top-level class of the semantic hierarchical structure of the dataset.
-To reduce order effects, target sounds are randomly presented to each subject.
+The target sounds are selected such that there are at least two target sounds in each top-level class of the semantic hierarchical structure of the dataset.
+To reduce order effects, target sounds are presented to each subject in random order.
 
-First, the subject click a \textit{Play target sound} button to listen to the target sound.
-Then, the subject may replay the target sound  as many times as he likes.
+First, the subject clicks a \textit{Play target sound} button to listen to the target sound.
+They may then replay the target sound as many times as they like, even while browsing.
 A timer starts when the subject clicks on a circle for the first time.
-When the target sound is found, subject puts an end to the search by clicking on the \textit{Click if found} button.
+When the target sound is found, the subject puts an end to the search by clicking on the \textit{Click if found} button.
 This action 1) stops the timer and 2) loads a new target sound.
-If the subject does not find the correct target sound, an error message appears, and the experiment goes on with the same target sound.
+If the subject designates and incorrect sound, an error message appears, and the experiment goes on with the same target sound.
 
@@
 \begin{itemize}
-\item \textit{The search state}: which can be "pause" if the subject is currently not looking for a target sound (at the beginning of the experiment, or between two target sounds) or  "in progress"  if the subject is currently looking for a target sound.
-\item \textit{Remaining target sounds}: the number of target sounds which remain to find.
+\item \textit{The search state}: which can be "pause" if the subject is not currently looking for a target sound (at the beginning of the experiment, or between two target sounds) or  "in progress"  if they are.
+\item \textit{Remaining target sounds}: the number of target sounds that remain to be found.
 \end{itemize}
@@
 A crowdsourcing approach is adopted.
-The experiment is designed to be performed using the  \textit{chrome} web browser.
+The experiment is designed to be performed using the \textit{Chrome} web browser.
 The link to the experiment has been sent to three  mailing lists, namely \textit{Music-Ir, Auditory} and an internal IRCAM mailing list.
 Subjects are only allowed to perform the experiment once, and on one interface only.
-Data were automatically collected server-side at the end of the experiment.
+Data were automatically collected server-side at the end of the experiments.
 Subjects were asked to use headphones.
@@
 We use one-way ANOVA and two samples \textit{t} tests at the 5\% significance level to test for statistical significance.
-All statistical analysis are performed using native functions of the computing environment MATLAB.
+All statistical analyses are performed using native functions of the MATLAB computing environment.
 
 \section{Results} \label{results}
+
 \subsection{Outlier detection}
@@
 \begin{table}[t]
-\caption{\label{tab1} Criterion used to detect outliers}
+\caption{\label{tab1} Criterion used to detect outliers.}
 \begin{center}
@@
 A commonly used method to detect outliers in  human-computer interaction studies is to consider as outlier an observation  which deviates of at least $\pm2$ standard deviation from the average \cite{komarov2013crowdsourcing}.
-though, as this latter method is not robust to the presence of isolated extreme observations (as it is often the case for crowdsourcing experiment), we follow the  method proposed by \cite{komarov2013crowdsourcing} and used the Inter-Quartile Range ($IQR$).
-With this approach, an observation is considered to be an outlier if its value is more than $3*IQR$ higher than the third quartile or more than $3*IQR$ lesser than the first quartile.
+However, as this latter method is not robust to the presence of isolated extreme observations (as it is often the case for crowdsourcing experiment), we follow the  method proposed by \cite{komarov2013crowdsourcing} and use the Inter-Quartile Range ($IQR$).
+With this approach, an observation is considered to be an outlier if its value is more than $3*IQR$ higher than the third quartile or more than $3*IQR$ less than the first quartile.
 This methods is applied in this study to the four following criteria :
@@
 Using the $IQR$ method, 5 subjects are detected as outliers and removed from the analysis.
-2 subjects used PD and 3 subjects AD.
-The tables~\ref{tab1} show the criteria used to detect the 5 outliers.
+2 of those subjects used PD and 3 subjects AD.
+Table~\ref{tab1} shows the criteria used to detect the 5 outliers.
 3 subjects are detected by considering the \textit{absolute duration}.
@@
 1 subject is detected by considering the total search duration (46 minutes), and an other by observing the total number of heard sounds (1800 heard sounds, roughly 12 times the total size of the corpus).
-And lastly 2 subjects are detected by observing the maximum number of times they had to hear a target sounds before finding it (4 and 8 times).
+And lastly 2 subjects are detected by observing the maximum number of times they had to hear a target sound before finding it (4 and 8 times).
 
@@
 \end{center}
-\caption{\label{fig1} Average and standard deviation for (a) the total search durations, (b) the number of heard sounds and (c) the number of unique heard sounds}
+\caption{\label{fig1} Average and standard deviation for (a) the total search durations, (b) the number of heard sounds and (c) the number of unique heard sounds.}
 \end{figure}
@@
 \end{center}
-\caption{\label{fig2} evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) average the numbers of unique heard sounds at each target sound search.
-Lines are linear regression.}
+\caption{\label{fig2} evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search.
+Lines are linear regression fits.}
 \end{figure}
@@
 \item the \textit{total search duration}
-\item the number of sounds heard \mathias{on met heard avant ou apres sounds ?} 
+\item the number of heard sounds
 \item the number of unique sounds heard. By "unique" we mean that, if a same sound is heard 10 times during the 13 searches of the  experiment, it counts only for one. 
@@
 The third data allows us to measure the selectivity of the interfaces.
-A low number of unique heard sounds indicates that subjects understood the spatial organisation of the dataset, and use this knowledge to improve their searches.
-In contrary, a high number of heard sounds without duplication suggest that the subject did not understood the way sounds are organized in space, and tends to play all the sounds at each search.
+A low number of unique heard sounds indicates that subjects understood the spatial organisation of the dataset, and used this knowledge to improve their searches.
+On the contrary, a high number of heard sounds without duplication suggests that the subject did not understand the way sounds are organized in space, and tended to play all the sounds at each search.
 The maximum number of unique heard sounds is the corpus size: 149 sounds.
 
-The averages and standard deviations for the three metrics are presented on Figure \ref{fig1}.
-The type of display have a significant effect on the the \textit{total search duration} ($F[2,52]=3.63; p=0.03$), the number of heard sounds ($F[2,52]=5.65; p<0.01$) and the number of unique heard sounds ($F[2,52]=8.65; p<0.01$).
-Post hoc analysis on the total search time (Figure~\ref{fig1}~(a)) shows that FSD  perform better  than the other interfaces (FSD-PSD: $p=0.04$; FSD-AD: $p=0.02$), whereas PSD and AD seem to have similar results (PSD-AD: $p=0.88$).
+The averages and standard deviations for the three metrics are presented on Figure~\ref{fig1}.
+The type of display has a significant effect on the the \textit{total search duration} ($F[2,52]=3.63; p=0.03$), the number of heard sounds ($F[2,52]=5.65; p<0.01$) and the number of unique heard sounds ($F[2,52]=8.65; p<0.01$).
+Post hoc analysis on the total search time (Figure~\ref{fig1}~(a)) shows that FSD performs better than the other interfaces (FSD-PSD: $p=0.04$; FSD-AD: $p=0.02$), whereas PSD and AD seem to have similar results (PSD-AD: $p=0.88$).
 
@@
 For AD, 75\% of the subjects heard more than 140 sounds, and 25\% heard at least 148 sounds, that is almost the entire database.
-Considering PSD, 75\% of subjects heard  less than 134 sounds, against 143 for FSD.
+Considering PSD, 75\% of subjects heard less than 134 sounds, versus 143 for FSD.
 This time PSD and FSD perform equivalently (PD-FD: $p=0.58$).
 
-According to those results, a hierarchical organization of the dataset based on semantic values (PSD and FSD) allows the users to retrieve the 13 target sounds 1) more efficiently, and 2) by listening to a smaller amount of sounds than an organization based on acoustic descriptors (AD).
+According to those results, a spatial display following a hierarchical organization of the dataset based on semantic values (PSD and FSD) allows the users to retrieve the 13 target sounds (1) more efficiently and (2) by listening to a smaller amount of sounds, than an organization based on acoustic descriptors (AD).
 However, those two effects are significantly compromised when users have to parse the entire hierarchy to reach the first target sound, as in the PSD display.
@@
 \begin{itemize}
-\item the duration of each target sound search 
-\item the number of heard sounds for each target sound search 
+\item the duration of each target sound search,
+\item the number of heard sounds for each target sound search,
 \item the number of unique heard sounds for each target sound search. 
@@
 
-Figure~\ref{fig2}~(a) shows the evolution of the average durations of each target sound search observed for AD, FSD and PSD.
+Figure~\ref{fig2}~(a) shows the evolution of the average duration of each target sound search observed for AD, FSD and PSD.
 As shown by a linear regression of the data, there is an overall increase of efficiency with respect to the index of the search.
@@
 In this paper, two displays allowing users to explore a semantically organized sound dataset without written textual help are presented.
-The interfaces distribute sounds represented by circles on a 2D space.
-The spatial organisation is driven by semantic features.
-Those two semantic displays are assessed and compared to a third display in which spatial configuration depends upon acoustic features.
+The interfaces distribute sounds represented by circles on a 2D space, following an underlying hierarchical semantic organization.
+Those two semantic displays are assessed and compared to a third display in which the spatial configuration depends upon acoustic features.
 The tests consist in data retrieval tasks.
-The Full-Display (FD), that allows users to directly visualize the leaf classes of the semantic hierarchical structure, proves to be the most effective interface for the task.
+The Full Semantic Display (FD), that allows users to directly visualize the leaf classes of the semantic hierarchical structure, proves to be the most effective interface for the task.
 
@@
 First, a spatial configuration based on semantic features is more effective to retrieve target sounds than a spatial configuration based on acoustic features.
-Second, an imposed visualisation of the semantic hierarchical structure of the dataset does not help user to understand and learn the spatial configuration of the semantic class, but instead disturbs the navigation.
+Second, imposing the exploration of the semantic hierarchical structure by constraining its visibility does not help the user understand and learn the spatial configuration of the semantic class, but instead disturbs the navigation.
 
-In the dataset considered in this study, there are as many leaf classes as sounds in the dataset, which would be unrealistic for large datasets.
-Though the organization can easily be adapted by considering the leaf classes as \textit{collections} of semantically similar sound samples.
+In the dataset considered in this study, there are as many leaf classes as sounds in the dataset, which would be unrealistic for large datasets, but the organization can easily be adapted by considering the leaf classes as \textit{collections} of semantically similar sound samples.
 Thus, two sounds of \textit{male-yelling} would be grouped into a single leaf class with the tag \textit{male-yelling}.

### 2016-02-14 lafaySsf2016.tex
submission v1

@@
 %\texttt{www.acm.org/eaddress.htm}}}
-%
+%strain
 % You need the command \numberofauthors to handle the 'placement
@@
 The Euclidean distance between time averaged features is then computed for each sound pair.
-A non metric multidimensional scaling with Kruskal's normalized stress is then employed to project the data into two dimensions according to this distance, see Figure \ref{figXP3}.
+A non metric multidimensional scaling with Kruskal's normalized stress \cite{kruskal1964multidimensional} is then employed to project the data into two dimensions according to this distance, see Figure \ref{figXP3}.
 
@@
 Thus the PSD has a constrained exploration system.
-When a user clicks on a circle, sub-circles are  automatically revealed to him in a gradual way.
 Each time a sub-circle is automatically revealed, its sound prototype is played.

### 2016-05-09 responseToReviewers.tex
adding response

\documentclass[10pt]{article}

\title{Response to Reviewers: \\ AES-D-16-00016
Semantic browsing of sound databases without keywords}

\usepackage{color}
\newcommand{\gl}[1]{\textcolor{red}{Gregoire : #1}}
\newcommand{\nm}[1]{\textcolor{magenta}{Nicolas : #1}}
\newcommand{\ml}[1]{\textcolor{blue}{ Mathieu : #1}}

\begin{document}

\maketitle

We would like to thank the editor and the reviewers for their comments and suggestions. Following these comments, we made several changes to the article, which are summarized here. The next sections list our answers to each of the reviewer’s comments, with references to the revised manuscript where appropriate.

\ml{TODO: put analysis data online, and website}

\section{Responses to Reviewer 1}

\begin{enumerate}

\item \emph{}

\end{enumerate}

\section{Responses to Reviewer 2}

\begin{enumerate}

\item \emph{}

\end{enumerate}

\section{Full text}

Reviewer 1: This paper proposes a new method for navigating environmental sound libraries using a constrained and unconstrained display that visualises semantic organisation of sounds. The authors build and evaluate two interfaces for navigation and compare them to a baseline method in which sounds are arranged in a 2D space using on MFCC based distances mapped to two dimensions using multi dimensional scaling. The semantic grouping in contrast are based on a hierarchical structure of conceptual/cognitive relatedness that is manually annotated. The assessment looks at search efficiency in absolute terms using information collected during user studies, including search time and the number of sounds a user had to listen to before reaching a target sound. The evaluation also looks at how efficiency changes over time as users become more familiar with the interface.

The idea is relatively novel and clearly presented, the collected data is valuable and subjected to appropriate statistical validation. There are clear conclusions drawn from this. For these reasons I recommend the paper to be published after minor editorial changes and clarifications.

The only question regarding the technical content that wasn't entirely clear to me is whether the sounds heard counter was reset for each search? That is, wehther the total and unique heard sounds are related to the whole experiment or an average value per target? How many target sounds were assessed by each subject? Was this controlled by the study or was the study simply time limited? Perhaps in the post-hoc analyses it would be useful to know if FSD produces significant results in pair-wise tests with Bonferroni correction (since this is, albeit simple, still a multiple comparison test).

Typos:
On page 5, Section 5.2  "...the three metrics are presented on Figure 6. " => should be in Figure 6.

Reviewer 2: Any additional comments you write here will be available to the author. You may also upload an annotated manuscript and/or detailed comments file.

The authors describe some user experiments investigating the use of semantic organization for visual browsing of sound databases without keywords. 
The main idea is to display the hierarchical organization of the data to facilitate browsing. Overall the paper is well written and easy to understand. The 
experimental methodology is good and the results support the conclusions with some caveats that I mention below. The bibliography is good but 
could be improved with some more recent references in music browsing. I think it will make a good contribution to the special issue especially 
if one or more of the following suggestions are addressed by the authors - they are not essential but I think they would make the paper 
much stronger and help it have more impact. Currently it reads like a strong conference paper but a weak journal article. 
My suggestions are:

1) The title is misleading i.e sound databases is too general and could encompass music, speech such oral histories, sound effects, bio-acoustic signals etc. 
The assumption that the findings generalize to other types of sound is too far stretched as for example music and urban sound are extremely different 
in how they are perceived organized by listeners. I would prefer a more honest title such as:

Semantic browsing of urban envrionment sounds without keywords

If the author replicate the same experiment for example using a music collection or a collection of bird songs and the findings were consistent 
this would justify the old title (and make the paper much stronger)

2) The references are relatively old. There have been several newer efforts to develop browsing interfaces especially in the music domain. 
For example the Musicream, Musicrainbow and Musicsun interfaces. 
Knees, Peter, et al. "An innovative three-dimensional user interface for exploring music collections enriched." Proceedings of the 14th annual ACM international conference on Multimedia. ACM, 2006.
Lillie, Anita Shen. MusicBox: Navigating the space of your music. Diss. Massachusetts Institute of Technology, 2008.

3) The uses of Kruskal's normalized stress is interesting but raises the concern that the hierarchical organization improvement might be specific to this approach. It would have been 
really good to have a few other ways to map the audio features to a spatial display (the classic PCA for example or SOM) and then compare the different interfaces. I realize that 
this would require more extensive experiments but it would make the paper much stronger.

4) Finally even though I agree with some of the arguments for avoiding keywords as a way of searching they can be used to help the user understand the space. Even some thing 
as simple as showing the word associated with a category when navigating would probably help. This would be an easy configuration to test and I am surprised the authors 
have done so. I suspect that the speed and efficiency would improve.

\end{document}

### 2016-05-31 responseToReviewers.tex
preparing v2

@@
 
-\ml{TODO: put analysis data online, and website}
+\ml{TODO: put analysis data online}
 
@@
 
+\gl{do it}
+
 \begin{enumerate}
 
-\item \emph{}
+\item \emph{[Are] the sounds heard counter was reset for each search? That is, whether the total and unique heard sounds are related to the whole experiment or an average value per target?}
+
+\item \emph{How many target sounds were assessed by each subject?}
+
+\item \emph{Was this controlled by the study or was the study simply time limited?}
+
+\item \emph{Perhaps in the post-hoc analyses it would be useful to know if FSD produces significant results in pair-wise tests with Bonferroni correction (since this is, albeit simple, still a multiple comparison test).}
+
+\item \emph{On page 5, Section 5.2  "...the three metrics are presented on Figure 6. " => should be in Figure 6.}
+
+This typo is now fixed.
 
@@
 
-\item \emph{}
+\item \emph{If the author replicate the same experiment for example using a music collection or a collection of bird songs and the findings were consistent 
+this would justify the old title (and make the paper much stronger)}
+
+This is an excellent suggestion that we decided to do by considering an alternative dataset of sounds produced by musical instruments.
+As shown on Section 6 of the updated paper, the findings are consistent with the ones obtained using the urban sound dataset.
+
+We thus keep the original title.
+
+\item \emph{The references are relatively old. There have been several newer efforts to develop browsing interfaces especially in the music domain. 
+For example the Musicream, Musicrainbow and Musicsun interfaces. 
+Knees, Peter, et al. "An innovative three-dimensional user interface for exploring music collections enriched." Proceedings of the 14th annual ACM international conference on Multimedia. ACM, 2006.
+Lillie, Anita Shen. MusicBox: Navigating the space of your music. Diss. Massachusetts Institute of Technology, 2008.}
+
+Those references are added to the previous work (Section 1).
+
+\item \emph{The uses of Kruskal's normalized stress is interesting but raises the concern that the hierarchical organization improvement might be specific to this approach. It would have been 
+really good to have a few other ways to map the audio features to a spatial display (the classic PCA for example or SOM) and then compare the different interfaces.}
+
+We agree that more extensive research should be done regarding the type of features used and the type of projection used.
+We originally chose the MDS as it has better performance than the PCA and have the advantage of not relying on any parameter setting unlike the SOM approach.
+
+This latter motivation is added to the manuscript in Section 3.
+
+Comparing to an optimized SOM display would have been interested, along with the use of other features.
+That said, we decided to dedicate the complementary experiments to the use of an alternate dataset.
+
+\item \emph{some thing 
+as simple as showing the word associated with a category when navigating would probably help. This would be an easy configuration to test and I am surprised the authors [not]
+have done so. I suspect that the speed and efficiency would improve.
+}
+
+We agree that this configuration would have helped us gaining a more precise understanding of the results.
+However, we chose to focus our study on completely non verbal displays as stated in the introduction.
+
+Even if this complementary study would have increase the level of knowledge, we decided to focus the complementary experiments to the use of an alternative dataset.
 
@@
 
-\section{Full text}
-
-Reviewer 1: This paper proposes a new method for navigating environmental sound libraries using a constrained and unconstrained display that visualises semantic organisation of sounds.
-The authors build and evaluate two interfaces for navigation and compare them to a baseline method in which sounds are arranged in a 2D space using on MFCC based distances mapped to two dimensions using multi dimensional scaling.
-The semantic grouping in contrast are based on a hierarchical structure of conceptual/cognitive relatedness that is manually annotated.
-The assessment looks at search efficiency in absolute terms using information collected during user studies, including search time and the number of sounds a user had to listen to before reaching a target sound.
-The evaluation also looks at how efficiency changes over time as users become more familiar with the interface.
-
-The idea is relatively novel and clearly presented, the collected data is valuable and subjected to appropriate statistical validation.
-There are clear conclusions drawn from this.
-For these reasons I recommend the paper to be published after minor editorial changes and clarifications.
-
-The only question regarding the technical content that wasn't entirely clear to me is whether the sounds heard counter was reset for each search?
-That is, wehther the total and unique heard sounds are related to the whole experiment or an average value per target?
-How many target sounds were assessed by each subject?
-Was this controlled by the study or was the study simply time limited?
-Perhaps in the post-hoc analyses it would be useful to know if FSD produces significant results in pair-wise tests with Bonferroni correction (since this is, albeit simple, still a multiple comparison test).
-
-Typos:
-On page 5, Section 5.2  "...the three metrics are presented on Figure 6. " => should be in Figure 6.
-
-Reviewer 2: Any additional comments you write here will be available to the author. You may also upload an annotated manuscript and/or detailed comments file.
-
-The authors describe some user experiments investigating the use of semantic organization for visual browsing of sound databases without keywords. 
-The main idea is to display the hierarchical organization of the data to facilitate browsing. Overall the paper is well written and easy to understand. The 
-experimental methodology is good and the results support the conclusions with some caveats that I mention below. The bibliography is good but 
-could be improved with some more recent references in music browsing. I think it will make a good contribution to the special issue especially 
-if one or more of the following suggestions are addressed by the authors - they are not essential but I think they would make the paper 
-much stronger and help it have more impact. Currently it reads like a strong conference paper but a weak journal article. 
-My suggestions are:
-
-1) The title is misleading i.e sound databases is too general and could encompass music, speech such oral histories, sound effects, bio-acoustic signals etc. 
-The assumption that the findings generalize to other types of sound is too far stretched as for example music and urban sound are extremely different 
-in how they are perceived organized by listeners. I would prefer a more honest title such as:
-
-Semantic browsing of urban envrionment sounds without keywords
-
-If the author replicate the same experiment for example using a music collection or a collection of bird songs and the findings were consistent 
-this would justify the old title (and make the paper much stronger)
-
-2) The references are relatively old. There have been several newer efforts to develop browsing interfaces especially in the music domain. 
-For example the Musicream, Musicrainbow and Musicsun interfaces. 
-Knees, Peter, et al. "An innovative three-dimensional user interface for exploring music collections enriched." Proceedings of the 14th annual ACM international conference on Multimedia. ACM, 2006.
-Lillie, Anita Shen. MusicBox: Navigating the space of your music. Diss. Massachusetts Institute of Technology, 2008.
-
-3) The uses of Kruskal's normalized stress is interesting but raises the concern that the hierarchical organization improvement might be specific to this approach. It would have been 
-really good to have a few other ways to map the audio features to a spatial display (the classic PCA for example or SOM) and then compare the different interfaces. I realize that 
-this would require more extensive experiments but it would make the paper much stronger.
-
-4) Finally even though I agree with some of the arguments for avoiding keywords as a way of searching they can be used to help the user understand the space. Even some thing 
-as simple as showing the word associated with a category when navigating would probably help. This would be an easy configuration to test and I am surprised the authors 
-have done so. I suspect that the speed and efficiency would improve.
-
 \end{document}

### 2016-05-31 lafaySsf2016.tex
nico names
preparing v2

@@
 \usepackage{color,url}
-\newcommand{\gregoire}[1]{\textcolor{red}{Gregoire : #1}}
-\newcommand{\mathias}[1]{\textcolor{green}{Mathias : #1}}
+\newcommand{\gl}[1]{\textcolor{red}{Gregoire : #1}}
+\newcommand{\mr}[1]{\textcolor{green}{Mathias : #1}}
 \newcommand{\ml}[2][]{\textcolor{blue}{#1 #2}}
@@
 \authorgroup{
-\author{Gr\'egoire Lafay$^1$}, \author{Nicolas Misdarris$^2$},
+\author{Gr\'egoire Lafay$^1$}, \author{Nicolas Misdariis$^2$},
 \author{Mathieu Lagrange$^1$}, \author{Mathias Rossignol$^2$}
@@
 
-%% A category with the (minimum) three required fields
-%\category{H.2.8}{Database Management}{Database Applications}[Data mining]
-%%%A category including the fourth, optional field follows...
-%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
-
-%\terms{Design, Experimentation, Human Factors}
-
 \keywords{Audio content management and display, Semantic sound data mining}
@@
 \item Pre-defined verbal descriptions of the sounds made available to the users may potentially bias their browsing and final selection.
-\item Localization of the query interface is made difficult as the translation of some words referring to qualitative aspects of the sound, such as its ambience, is notoriously ambiguous and subject to cultural specificities.
+\item Localization of the query interface is made difficult as the translation of some words referring to qualitative aspects of the sound, such as its ambiance, is notoriously ambiguous and subject to cultural specificities.
 \item Unless considerable time and resources are invested into developing a multilingual interface, any system based on verbal descriptions can only be used with reduced performance by non-native speakers of the chosen language.
@@
 
-Most of the research effort in the sound design and Music Information Retrieval (MIR) communities is focused on acoustical based indexing and browsing \cite{tzanetakis2003musescape, streich2008music}.
+Most of the research effort in the sound design and Music Information Retrieval (MIR) communities is focused on acoustical based indexing and browsing \cite{tzanetakis2003musescape, streich2008music,goto2005musicream,pampalk2007musicsun}.
 Typically, the items (sound effects or pieces of music) are modeled by processing the digital audio waveform through some signal processing pipeline in order to get a compact description of each item
-\cite{coleman2007mused}.
-Then, statistical projections or embedding techniques, like Principal Components Analysis (PCA),  Multi Dimensional Scaling (MDS) \cite{schwarz2009sound,Cano2002}, Self Organizing Maps (SOM) \cite{pampalk2004exploring} and the like are used to project the items in a two or three dimensional space while preserving as much as possible the distances among them.
+\cite{coleman2007mused} with sometime additional textual features.
+Then, statistical projections or embedding techniques, like Principal Components Analysis (PCA) \cite{MusicBox},  Multi Dimensional Scaling (MDS) \cite{schwarz2009sound,Cano2002}, Self Organizing Maps (SOM) \cite{pampalk2004exploring,pampalk2006musicrainbow,knees2006innovative} and the like are used to project the items in a two or three dimensional space while preserving as much as possible the distances among them.
 One of the advantages of such an approach is its ability to scale to very large databases \cite{schwarz2009scalability} as it does not need any kind of manual annotation and allows the user to search by similarity efficiently according to acoustical properties.
@@
 
-\section{Dataset} \label{dataset}
+\section{Datasets} \label{dataset}
 
@@
 \end{center}
-\caption{\label{figSSF} Semantic displays are  based on the semantic hierarchical structure of the dataset.}
+\caption{\label{figSSF} Semantic displays are  based on the semantic hierarchical structure of the dataset. \gl{generate an alternative bw figure}}
 \end{figure}
 
-The dataset considered in this study is composed of 149 urban environmental sound events.
-Its semantic organization is as follows: a sound is characterized by a tag describing the physical source of the sound (\textit{man-yelling}, \textit{car-passing}).
+The datasets considered in this study are respectively composed of 149 urban environmental sound events and 137 musical sounds produced by musical instruments.
+
+The former has a semantic organization as follows: a sound is characterized by a tag describing the physical source of the sound (\textit{man-yelling}, \textit{car-passing}).
 Sounds are then hierarchically grouped into classes according to their tags (\textit{car} $>$ \textit{car-passing}; \textit{car} $>$ \textit{car-starting}).
@@
 It is our belief that using such category names to build the hierarchical structure makes it perceptually motivated, and can thus help the users to efficiently browse the dataset.
+
+Similarly, the latter also has a semantic organization \gl{describe music dataset}
 
@@
 The Euclidean distance between time averaged features is then computed for each sound pair.
-A non metric multidimensional scaling with Kruskal's normalized stress \cite{kruskal1964multidimensional} is then employed to project the data into two dimensions according to this distance, see Figure \ref{figXP3}.
+A non metric multidimensional scaling (MDS) with Kruskal's normalized stress \cite{kruskal1964multidimensional} is then employed to project the data into two dimensions according to this distance, see Figures \ref{figXP3urban} \ref{figXP3music}.
+The latter exhibit a clear distinction between percussive (left) and sustained sounds (right).
+
+Among the several alternative that exists, MFCCs features are considered as they are widely used in the audio processing community.
+The MDS is selected as it has lower sensitivity to outliers than the PCA and have the advantage of not relying on any parameter setting unlike the SOM approach.
+
+%\begin{figure}[t]
+%\begin{center}
+%\includegraphics[scale=0.18]{gfx/xp3.png} 
+%\end{center}
+%\caption{\label{figXP3} Acoustical Display (AD).}
+%\end{figure}
 
@@
 \begin{center}
-\includegraphics[scale=0.18]{gfx/xp3.png} 
+\includegraphics[width=0.5\textwidth]{gfx/urban_mds.png} 
 \end{center}
-\caption{\label{figXP3} Acoustical Display (AD).}
+\caption{\label{figXP3urban} Acoustical Display for the urban sound dataset (AD). Colors and legend are removed during the experiment.}
 \end{figure}
 
-Two semantically oriented displays are then proposed, called Full Semantic Display (FSD) and Progressive Semantic Display (PSD), shown on Figures~\ref{figXP2} and~\ref{figXP1}, respectively.
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=0.5\textwidth]{gfx/music_mds.png} 
+\end{center}
+\caption{\label{figXP3music} Acoustical Display for the musical sound dataset (AD). Colors and legend are removed during the experiment.}
+\end{figure}
+
+Two semantically oriented displays are then proposed, called Full Semantic Display (FSD) and Progressive Semantic Display (PSD)%, shown in Figures~\ref{figXP2} and~\ref{figXP1}, respectively. 
 Both display  consider the hierarchical structure of the dataset to organize sounds.
 Each sound class is represented by a filled circle.
-Circles are then packed together according to the hierarchical semantic organization of the dataset, as shown on Figure~\ref{figSSF}.
+Circles are then packed together according to the hierarchical semantic organization of the dataset, as shown in Figure~\ref{figSSF}.
 Thus, subclasses belonging to the same class are close to each others.
@@
 In this display, the leaf classes are distributed  in the same manner as FSD.
-Thus, the spatial configuration of the unfolded version of PSD, which may be obtained after discovering all the classes and subclasses, is equivalent to the one of FSD, as shown on Figures~\ref{figXP2}, \ref{figXP1} and \ref{figSSF}.
+Thus, the spatial configuration of the unfolded version of PSD, which may be obtained after discovering all the classes and subclasses, is equivalent to the one of FSD, as shown in Figures~\ref{figXP2}, \ref{figXP1} and \ref{figSSF}.
 
@@
 
-\begin{figure}[t!]
-\begin{center}
-\includegraphics[scale=0.30]{gfx/XP2clean.pdf} 
-\end{center}
-\caption{\label{figXP2} Full Semantic Display (FSD): implicit hierarchical organization of semantic classes.}
-\end{figure}
+%\begin{figure}[t!]
+%\begin{center}
+%\includegraphics[scale=0.30]{gfx/XP2clean.pdf} 
+%\end{center}
+%\caption{\label{figXP2} Full Semantic Display (FSD): implicit hierarchical organization of semantic classes.}
+%\end{figure}
+%
+%\begin{figure*}[t]
+%\begin{center}
+%\includegraphics[scale=0.4]{gfx/XP1clean2.pdf} 
+%\end{center}
+%\caption{\label{figXP1} Progressive Semantic Display (PSD):  explicit hierarchical organization of semantic classes. (a) initial folded version; (b) partly folded version; (c) unfolded version.}
+%\end{figure*}
 
-\begin{figure*}[t]
-\begin{center}
-\includegraphics[scale=0.4]{gfx/XP1clean2.pdf} 
-\end{center}
-\caption{\label{figXP1} Progressive Semantic Display (PSD):  explicit hierarchical organization of semantic classes. (a) initial folded version; (b) partly folded version; (c) unfolded version.}
-\end{figure*}
-
-\section{Experiment} \label{test}
+\section{Experiments} \label{test}
 
@@
 
-During this experiment, the three displays are compared.
+During this experiment, the three displays are compared using two corpora: the urban sound corpus and the musicla sound corpus.
 By comparing AD and xSD (the two semantic displays), the goal is to evaluate the  relative efficiencies of semantic based and acoustic based spatial configurations.
@@
 
-We evaluate and compare those three displays with a crowdsourcing web-based experiment \footnote{The test is available at \url{http://soundthings.org/research/speedSoundFinding/index.html}, each of the 3 displays is also available by appending a digit from 1 to 3 to \textit{index} in the above cited url.} for which we adopt a between subject approach, \textit{i.e.} a subject can only test one display.
-In this experiment, subjects are asked to successively retrieve 13 target sounds among the 149 of the whole dataset.
+We evaluate and compare those three displays with a crowdsourcing web-based experiment \footnote{The tests are available at \url{http://soundthings.org/research/speedSoundFinding/index.html} and \url{http://soundthings.org/research/speedSoundFindingMusic/index.html}, each of the 3 displays is also available by appending a digit from 1 to 3 to \textit{index} in the above cited url.} for which we adopt a between subject approach, \textit{i.e.} a subject can only test one display.
+In this experiment, subjects are asked to successively retrieve 13 target sounds among the whole dataset.
 The target sounds are selected such that there are at least two target sounds in each top-level class of the semantic hierarchical structure of the dataset.
@@
 
-60 subjects have completed the experiment, 20 for each interface.
+60 subjects have completed the experiment for the urban sound dataset, 20 for each interface. \ml{describe for music}
 
-\subsection{Statistical analysis}
-
-We use one-way ANOVA and two samples \textit{t} tests at the 5\% significance level to test for statistical significance.
-All statistical analyses are performed using native functions of the MATLAB computing environment.
-
-\section{Results} \label{results}
-
-\subsection{Outlier detection}
-
-\begin{table}[t]
-\caption{\label{tab1} Criterion used to detect outliers.}
-\begin{center}
-\begin{tabular}{ccccc}
-criterion      &  1  &  2  &  3  &  4  \\
-\hline
-outlier 1 (PD) & --- & --- & --- &  x  \\
-outlier 2 (PD) &  x  & --- & --- & --- \\  
-outlier 3 (AD) & --- & --- & --- &  x \\
-outlier 4 (AD) &  x  &  x  &  x  & --- \\
-outlier 5 (AD) &  x  & --- & --- & --- \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
+\subsection{Outliers Detection}
 
@@
 
+\subsection{Statistical analysis}
+
+We use one-way ANOVA and two samples \textit{t} tests at the 5\% significance level to test for statistical significance.
+All statistical analyses are performed using native functions of the MATLAB computing environment.
+
+\section{Results on Urban Sounds} \label{resultsu}
+
+\subsection{Outliers}
+
+%\begin{table}[t]
+%\caption{\label{tab1} Criterion used to detect outliers.}
+%\begin{center}
+%\begin{tabular}{ccccc}
+%criterion      &  1  &  2  &  3  &  4  \\
+%\hline
+%outlier 1 (PD) & --- & --- & --- &  x  \\
+%outlier 2 (PD) &  x  & --- & --- & --- \\  
+%outlier 3 (AD) & --- & --- & --- &  x \\
+%outlier 4 (AD) &  x  &  x  &  x  & --- \\
+%outlier 5 (AD) &  x  & --- & --- & --- \\
+%\hline
+%\end{tabular}
+%\end{center}
+%\end{table}
+
 Using the $IQR$ method, 5 subjects are detected as outliers and removed from the analysis.
 2 of those subjects used PD and 3 subjects AD.
-Table~\ref{tab1} shows the criteria used to detect the 5 outliers.
+%Table~\ref{tab1} shows the criteria used to detect the 5 outliers.
 3 subjects are detected by considering the \textit{absolute duration}.
@@
 \begin{center}
-\includegraphics[scale=0.4]{gfx/analyse1.eps} 
+\includegraphics[scale=0.4]{gfx/analyse_urban1.eps} 
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.4]{gfx/analyse2.eps} 
+\includegraphics[scale=0.4]{gfx/analyse_urban2.eps} 
 \end{center}
@@
 
-The averages and standard deviations for the three metrics are presented on Figure~\ref{fig1}.
+The averages and standard deviations for the three metrics are presented in Figure~\ref{fig1}.
 The type of display has a significant effect on the the \textit{total search duration} ($F[2,52]=3.63; p=0.03$), the number of heard sounds ($F[2,52]=5.65; p<0.01$) and the number of unique heard sounds ($F[2,52]=8.65; p<0.01$).
@@
 
-\subsection{Learning phenomenon}
+\subsection{Learning}
 
@@
 This latter result suggest that for both semantic displays, users have parsed the same range of the dataset, although users of FSD have done so by listening to fewer sounds and more quickly.
+
+\section{Results on Musical Sounds}
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[scale=0.4]{gfx/analyse_music1.eps} 
+\end{center}
+\caption{\label{fig1Music} Average and standard deviation for (a) the total search durations, (b) the number of heard sounds and (c) the number of unique heard sounds.}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[scale=0.4]{gfx/analyse_music2.eps} 
+\end{center}
+\caption{\label{fig2Music} evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search.
+Lines are linear regression fits.}
+\end{figure}
+
+\gl{merge fig. 5 and 7}
+% if space is need fig 1 could be merged
+
+\subsection{Outliers}
+
+\subsection{Efficiency}
+
+\ref{fig1Music}
+
+\subsection{Learning}
+
+\ref{fig2Music}
 

### 2016-06-01 lafaySsf2016.tex
add results for music datasets

@@
 \end{center}
-\caption{\label{figSSF} Semantic displays are  based on the semantic hierarchical structure of the dataset. \gl{generate an alternative bw figure}}
+\caption{\label{figSSF} Semantic displays are  based on the semantic hierarchical structure of the dataset.}
 \end{figure}
@@
 
-Similarly, the latter also has a semantic organization \gl{describe music dataset}
+\gl{Similarly, the latter also has a semantic organization: high and mid level classes categorize musical instruments according to the means by which sound is produced, while leaf level classes are related to specific instruments (\textit{string} $>$ \textit{plucked-string} $>$ \textit{viola}; \textit{free-aerophone} $>$ \textit{brassy} $>$ \textit{trumpet})}
 
@@
 In this display, the leaf classes are distributed  in the same manner as FSD.
-Thus, the spatial configuration of the unfolded version of PSD, which may be obtained after discovering all the classes and subclasses, is equivalent to the one of FSD, as shown in Figures~\ref{figXP2}, \ref{figXP1} and \ref{figSSF}.
+Thus, the spatial configuration of the unfolded version of PSD, which may be obtained after discovering all the classes and subclasses, is equivalent to the one of FSD.
 
@@
 
-60 subjects have completed the experiment for the urban sound dataset, 20 for each interface. \ml{describe for music}
+60 subjects have completed the experiment for the urban sound dataset, 20 for each interface.
+\gl{43 subjects have completed the experiment for the urban sound dataset: 15 for the PSD interface, 13 for the FSD interface and 15 for the AD interface.}
 
@@
 
-We use one-way ANOVA and two samples \textit{t} tests at the 5\% significance level to test for statistical significance.
-All statistical analyses are performed using native functions of the MATLAB computing environment.
+\gl{We use one-way ANOVA to test for statistical significance.
+Post hoc analyses are done using two samples \textit{t} tests with Bonferroni correction.
+All statistical analyses are performed at the 5\% significance level using native functions of the MATLAB computing environment.}
 
@@
 
-Using the $IQR$ method, 5 subjects are detected as outliers and removed from the analysis.
-2 of those subjects used PD and 3 subjects AD.
+Using the $IQR$ method, 5 subjects are detected as outliers for the urban experiment, and removed from the analysis.
+2 of those subjects used PSD and 3 subjects AD.
 %Table~\ref{tab1} shows the criteria used to detect the 5 outliers.
@@
 They spent respectively 50 minutes, 5 hours and 17.5 hours to complete the experiment.
-1 subject is detected by considering the total search duration (46 minutes), and an other by observing the total number of heard sounds (1800 heard sounds, roughly 12 times the total size of the corpus).
+1 subject is detected by considering the \textit{total search duration} (46 minutes), and an other by observing the \textit{total number of heard sounds} (1800 heard sounds, roughly 12 times the total size of the corpus).
 And lastly 2 subjects are detected by observing the maximum number of times they had to hear a target sound before finding it (4 and 8 times).
@@
 \end{center}
-\caption{\label{fig1} Average and standard deviation for (a) the total search durations, (b) the number of heard sounds and (c) the number of unique heard sounds.}
+\caption{\label{fig1urban} Average and standard deviation for (a) the total search durations, (b) the number of heard sounds and (c) the number of unique heard sounds.}
 \end{figure}
@@
 \end{center}
-\caption{\label{fig2} evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search.
+\caption{\label{fig2urban} evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search.
 Lines are linear regression fits.}
@@
 
-The averages and standard deviations for the three metrics are presented in Figure~\ref{fig1}.
-The type of display has a significant effect on the the \textit{total search duration} ($F[2,52]=3.63; p=0.03$), the number of heard sounds ($F[2,52]=5.65; p<0.01$) and the number of unique heard sounds ($F[2,52]=8.65; p<0.01$).
-Post hoc analysis on the total search time (Figure~\ref{fig1}~(a)) shows that FSD performs better than the other interfaces (FSD-PSD: $p=0.04$; FSD-AD: $p=0.02$), whereas PSD and AD seem to have similar results (PSD-AD: $p=0.88$).
+The averages and standard deviations for the three metrics are presented in Figure~\ref{fig1urban}.
+The type of display has a significant effect on the the \textit{total search duration} ($F[2,52]=3.63; p<0.05$), the number of heard sounds ($F[2,52]=5.65; p<0.01$) and the number of unique heard sounds ($F[2,52]=8.65; p<0.01$).
 
-Similar results are found for the numbers of heard sounds (Figure~\ref{fig1}~(b)).
-FSD significantly outperforms the other interfaces (FSD-PSD: $p=0.02$; FSD-ASD: $p<0.01$), whereas PSD and AD show similar outcomes (PSD-AD: $p=0.42$).
+Post hoc analysis on the total search time (Figure~\ref{fig1urban}.a) shows no significant differences between the three interfaces (PSD-FSD: $p=0.09$; PSD-AD: $p=1$; FSD-AD: $p=0.07$), although the FSD interface achieves the best results.
 
-Figure~\ref{fig1}~(c) shows the results for the number of unique heard sounds.
-This time the results of AD are significantly lower than those of both PD and FD (AD-PD: $p<0.01$; AD-FD: $p<0.01$).
+Concerning the numbers of heard sounds (Figure~\ref{fig1urban}.b), FSD significantly outperforms the AD interface (FSD-AD: $p<0.01$),  whereas PSD and FSD as well as PSD and AD show similar outcomes (PSD-FSD: $p=0.06$; PSD-AD: $p=1$).
+
+Figure~\ref{fig1urban}.c shows the results for the number of unique heard sounds.
+This time the results of AD are significantly lower than those of both PSD and FSD (PSD-AD: $p<0.01$; FSD-AD: $p<0.01$).
 For AD, 75\% of the subjects heard more than 140 sounds, and 25\% heard at least 148 sounds, that is almost the entire database.
 Considering PSD, 75\% of subjects heard less than 134 sounds, versus 143 for FSD.
-This time PSD and FSD perform equivalently (PD-FD: $p=0.58$).
+This time PSD and FSD perform equivalently (PD-FD: $p=1$).
 
-According to those results, a spatial display following a hierarchical organization of the dataset based on semantic values (PSD and FSD) allows the users to retrieve the 13 target sounds (1) more efficiently and (2) by listening to a smaller amount of sounds, than an organization based on acoustic descriptors (AD).
+According to those results, a spatial display following a hierarchical organization of the dataset based on semantic values (PSD and FSD) is more efficient than an organization based on acoustic descriptors (AD) as it allows the users to quickly apprehend the dataset range and thus to specify their search.
+Furthermore the FSD interface allows the users to retrieve the 13 target sounds by listening to a smaller amount of sounds, than the AD.
 However, those two effects are significantly compromised when users have to parse the entire hierarchy to reach the first target sound, as in the PSD display.
@@
 
-Figure~\ref{fig2}~(a) shows the evolution of the average duration of each target sound search observed for AD, FSD and PSD.
+Figure~\ref{fig2urban}.a shows the evolution of the average duration of each target sound search observed for AD, FSD and PSD.
 As shown by a linear regression of the data, there is an overall increase of efficiency with respect to the index of the search.
-For the duration and the number of heard sounds, the AD and PSD perform equivalently, whereas the FSD exhibit better performance.
-In the case of the number of unique heard sounds, the two semantic displays are equivalent.
+For the duration and the number of heard sounds (Figure~\ref{fig2urban}.b), the AD and PSD perform equivalently, whereas the FSD exhibit better performance.
+In the case of the number of unique heard sounds (Figure~\ref{fig2urban}.c), the two semantic displays are equivalent.
 This latter result suggest that for both semantic displays, users have parsed the same range of the dataset, although users of FSD have done so by listening to fewer sounds and more quickly.
@@
 
+\gl{2 subjects are detected as outliers for the music experiment and removed from the analysis.
+1 subject used PSD and 1 subject used FSD.
+The first outlier is detected by considering the \textit{absolute duration} (5 hours to complete the experiment) and the second by looking at the \textit{absolute duration} (1 hour and 30 minutes) as well as the \textit{total search duration} (1 hour and 15 minutes) and the total \textit{number of heard sounds} (901).}
+
 \subsection{Efficiency}
 
-\ref{fig1Music}
+\gl{The averages and standard deviations for the three metrics are presented in Figure~\ref{fig1Music}.
+The type of display has a significant effect on the the \textit{total search duration} ($F[2,38]=10.25; p<0.01$), the number of heard sounds ($F[2,38]=13.15; p<0.01$) and the number of unique heard sounds ($F[2,38]=6.39; p<0.01$).
+}
+
+\gl{Considering the \textit{total search duration} (Figure~\ref{fig1Music}.a), post hoc analyses shows that PSD and FSD perform better than the AD interfaces (PSD-AD: $p<0.05$; FSD-AD: $p<0.01$;).
+No significant difference is found between PSD and FSD (PSD-FSD: $p=0.32$).
+}
+
+\gl{Figure~\ref{fig1Music}.b shows the results for the numbers of heard sounds.
+This time AD and PSD exhibit similar results (PSD-AD: $p<0.43$), while being both outperformed by FSD (FSD-PSD: $p<0.01$; FSD-AD: $p<0.01$).
+}
+
+\gl{Considering the numbers of unique heard sounds (Figure~\ref{fig1Music}.c), FSD significantly outperforms the AD interface (FSD-AD: $p<0.01$), while both PSD \emph{vs.} FSD and PSD \emph{vs.} AD show similar outcomes (PSD-FSD: $p=0.6$; PSD-AD: $p=0.9$).}
+
+\gl{This results confirms the trends observed for the urban sounds (see Section~\ref{resultsu}):}
+
+\begin{itemize}
+\item A semantic organization of the dataset significantly reduces  the number of heard sounds needed to find a target sound.
+\item Forcing the users to parse the entire hierarchical organization of the dataset seems to diminish the interface efficiency by increasing
+\end{itemize}
+
+\gl{It has to be noted that this time a semantic organization of the dataset (PSD and FSD) significantly reduces the search durations compared to an acoustic organization (AD).}
 
@@
 
-\ref{fig2Music}
+\gl{Figure~\ref{fig2Music} shows the evolution of the metrics computed at each target sound.
+Results for the music dataset are similar to those observed for the urban dataset.
+An overall increase of efficiency with respect to the index of the search is observed for the tree interfaces.
+For the duration and the number of heard sounds (Figure~\ref{fig2Music}.a and .b) the FSD interface exhibits better performance.
+Considering the number of unique heard sounds (Figure~\ref{fig2Music}.c), the learning of the dataset range seems to be quicker for the FSD interface than for the PSD interface, thus confirming that forcing the users to parse the hierarchy disturbs their researches.}
 
@@
 The tests consist in data retrieval tasks.
-The Full Semantic Display (FD), that allows users to directly visualize the leaf classes of the semantic hierarchical structure, proves to be the most effective interface for the task.
+The Full Semantic Display (FSD), that allows users to directly visualize the leaf classes of the semantic hierarchical structure, proves to be the most effective interface for the task.
 

### 2016-06-01 lafaySsf2016.tex
update abstract intro conclusion and merge contribution of gregoire

@@
 For the sake of comparison, we also consider a display whose spatial organization is based on acoustic cues.
-Those three displays are evaluated in terms of search speed in a crowdsourcing experiment.
-Results demonstrate the usefulness of using an implicit semantic organization to display sounds, both in terms of search speed and of efficiency of learning.
+Those three displays are evaluated in terms of search speed in a crowdsourcing experiment using two different corpora: the first being composed of sounds of urban environment and the second of sounds produced by musical instruments.
+Coherent results achieved by considering the two different corpora demonstrate the usefulness of using an implicit semantic organization to display sounds, both in terms of search speed and of efficiency of learning.
 \end{abstract}
@@
 Their effectiveness is studied through a search-based task whose aim is to find a given target sound by browsing the database using the display under evaluation.
+Two corpora are considered for this study.
+The first one is composed of sounds recorded in a urban environment and the second one is composed of sounds of musical instruments.
 
 The paper is organized as follows: in Section~\ref{previous}, previous work on the topic of sound database browsing is reviewed.
-Then, the corpus used in this study is described in Section~\ref{dataset}.
-The three displays under evaluation are next described in Section~\ref{display}, and the crowdsourcing test used to compare them is presented in Section~\ref{test}.
-Finally, the outcomes of this experiment are discussed in Section~\ref{results}.
+Then, the two corpora used in this study are described in Section~\ref{dataset}.
+The three displays under evaluation are next described in Section~\ref{display}, and the crowdsourcing test used to compare them using the two different corpora is presented in Section~\ref{test}.
+The outcomes of those experiments are discussed in Sections~\ref{resultsu} and ~\ref{resultsm}.
 
@@
 
-\gl{Similarly, the latter also has a semantic organization: high and mid level classes categorize musical instruments according to the means by which sound is produced, while leaf level classes are related to specific instruments (\textit{string} $>$ \textit{plucked-string} $>$ \textit{viola}; \textit{free-aerophone} $>$ \textit{brassy} $>$ \textit{trumpet})}
+Similarly, the latter also has a semantic organization: high and mid level classes categorize musical instruments according to the means by which sound is produced, while leaf level classes are related to specific instruments (\textit{string} $>$ \textit{plucked-string} $>$ \textit{viola}; \textit{free-aerophone} $>$ \textit{brassy} $>$ \textit{trumpet}).
 
@@
 60 subjects have completed the experiment for the urban sound dataset, 20 for each interface.
-\gl{43 subjects have completed the experiment for the urban sound dataset: 15 for the PSD interface, 13 for the FSD interface and 15 for the AD interface.}
+43 subjects have completed the experiment for the urban sound dataset: 15 for the PSD interface, 13 for the FSD interface and 15 for the AD interface.
+
+\ml{to update before final submission}
 
@@
 
-\gl{We use one-way ANOVA to test for statistical significance.
+We use one-way ANOVA to test for statistical significance.
 Post hoc analyses are done using two samples \textit{t} tests with Bonferroni correction.
-All statistical analyses are performed at the 5\% significance level using native functions of the MATLAB computing environment.}
+All statistical analyses are performed at the 5\% significance level.% using native functions of the MATLAB computing environment.
 
@@
 
-\section{Results on Musical Sounds}
+\section{Results on Musical Sounds} \label{resultsm}
 
@@
 
-\gl{merge fig. 5 and 7}
+%\gl{merge fig. 5 and 7}
 % if space is need fig 1 could be merged
@@
 
-\gl{2 subjects are detected as outliers for the music experiment and removed from the analysis.
+2 subjects are detected as outliers for the music experiment and removed from the analysis.
 1 subject used PSD and 1 subject used FSD.
-The first outlier is detected by considering the \textit{absolute duration} (5 hours to complete the experiment) and the second by looking at the \textit{absolute duration} (1 hour and 30 minutes) as well as the \textit{total search duration} (1 hour and 15 minutes) and the total \textit{number of heard sounds} (901).}
+The first outlier is detected by considering the \textit{absolute duration} (5 hours to complete the experiment) and the second by looking at the \textit{absolute duration} (1 hour and 30 minutes) as well as the \textit{total search duration} (1 hour and 15 minutes) and the total \textit{number of heard sounds} (901).
 
@@
 
-\gl{The averages and standard deviations for the three metrics are presented in Figure~\ref{fig1Music}.
+The averages and standard deviations for the three metrics are presented in Figure~\ref{fig1Music}.
 The type of display has a significant effect on the the \textit{total search duration} ($F[2,38]=10.25; p<0.01$), the number of heard sounds ($F[2,38]=13.15; p<0.01$) and the number of unique heard sounds ($F[2,38]=6.39; p<0.01$).
-}
 
-\gl{Considering the \textit{total search duration} (Figure~\ref{fig1Music}.a), post hoc analyses shows that PSD and FSD perform better than the AD interfaces (PSD-AD: $p<0.05$; FSD-AD: $p<0.01$;).
+Considering the \textit{total search duration} (Figure~\ref{fig1Music}.a), post hoc analyses shows that PSD and FSD perform better than the AD interfaces (PSD-AD: $p<0.05$; FSD-AD: $p<0.01$;).
 No significant difference is found between PSD and FSD (PSD-FSD: $p=0.32$).
-}
 
-\gl{Figure~\ref{fig1Music}.b shows the results for the numbers of heard sounds.
+Figure~\ref{fig1Music}.b shows the results for the numbers of heard sounds.
 This time AD and PSD exhibit similar results (PSD-AD: $p<0.43$), while being both outperformed by FSD (FSD-PSD: $p<0.01$; FSD-AD: $p<0.01$).
-}
 
-\gl{Considering the numbers of unique heard sounds (Figure~\ref{fig1Music}.c), FSD significantly outperforms the AD interface (FSD-AD: $p<0.01$), while both PSD \emph{vs.} FSD and PSD \emph{vs.} AD show similar outcomes (PSD-FSD: $p=0.6$; PSD-AD: $p=0.9$).}
-
-\gl{This results confirms the trends observed for the urban sounds (see Section~\ref{resultsu}):}
+Considering the numbers of unique heard sounds (Figure~\ref{fig1Music}.c), FSD significantly outperforms the AD interface (FSD-AD: $p<0.01$), while both PSD \emph{vs.} FSD and PSD \emph{vs.} AD show similar outcomes (PSD-FSD: $p=0.6$; PSD-AD: $p=0.9$).
+Those results confirms the trends observed for the urban sounds (see Section~\ref{resultsu}):
 
@@
 \item A semantic organization of the dataset significantly reduces  the number of heard sounds needed to find a target sound.
-\item Forcing the users to parse the entire hierarchical organization of the dataset seems to diminish the interface efficiency by increasing
+\item Forcing the users to parse the entire hierarchical organization of the dataset seems to diminish the interface efficiency by increasing the number of sounds to be listened to.
 \end{itemize}
 
-\gl{It has to be noted that this time a semantic organization of the dataset (PSD and FSD) significantly reduces the search durations compared to an acoustic organization (AD).}
+Considering that the musical corpus have perhaps a obvious semantic organization, PSD and FSD significantly reduces the search durations compared to an acoustic organization (AD), which was not the case for the urban sound dataset.
 
@@
 
-\gl{Figure~\ref{fig2Music} shows the evolution of the metrics computed at each target sound.
+Figure~\ref{fig2Music} shows the evolution of the metrics with respect to the target sound index.
 Results for the music dataset are similar to those observed for the urban dataset.
@@
 For the duration and the number of heard sounds (Figure~\ref{fig2Music}.a and .b) the FSD interface exhibits better performance.
-Considering the number of unique heard sounds (Figure~\ref{fig2Music}.c), the learning of the dataset range seems to be quicker for the FSD interface than for the PSD interface, thus confirming that forcing the users to parse the hierarchy disturbs their researches.}
+Considering the number of unique heard sounds (Figure~\ref{fig2Music}.c), the learning of the dataset range seems to be quicker for the FSD interface than for the PSD interface.
+Again, results show that enforcing the parsing of the hierarchy is not beneficial.
 
@@
 Those two semantic displays are assessed and compared to a third display in which the spatial configuration depends upon acoustic features.
-The tests consist in data retrieval tasks.
+The tests consist in data retrieval tasks done using two different corpora.
 The Full Semantic Display (FSD), that allows users to directly visualize the leaf classes of the semantic hierarchical structure, proves to be the most effective interface for the task.
@@
 
-In the dataset considered in this study, there are as many leaf classes as sounds in the dataset, which would be unrealistic for large datasets, but the organization can easily be adapted by considering the leaf classes as \textit{collections} of semantically similar sound samples.
+In the datasets considered in this study, there are as many leaf classes as sounds in the dataset, which would be unrealistic for large datasets.
+The organization can easily be adapted by considering the leaf classes as \textit{collections} of semantically similar sound samples.
 Thus, two sounds of \textit{male-yelling} would be grouped into a single leaf class with the tag \textit{male-yelling}.

### 2016-06-01 responseToReviewers.tex
coucou respons2review
response to reviewers 1

@@
 
-\gl{do it}
-
 \begin{enumerate}
@@
 
+The total and unique heard sounds are related to the whole experiment.
+
 \item \emph{How many target sounds were assessed by each subject?}
+
+All the subjects had to assess 13 target sounds.
 
@@
 
+Heu de quoi ?
+
 \item \emph{Perhaps in the post-hoc analyses it would be useful to know if FSD produces significant results in pair-wise tests with Bonferroni correction (since this is, albeit simple, still a multiple comparison test).}
+
+Post hoc analyses are now down using the Bonferroni procedure.
+All the p-values are updated.
+Considering the urban dataset, we don't achieve statistical significance for the search durations any-more.
+But this does not change the paper conclusions.
 

### 2016-06-01 responseToReviewers.tex
polishing response

@@
 
-The total and unique heard sounds are related to the whole experiment.
+The total and unique heard sounds are related to the whole experiment. The Section 5.2 of the updated manuscript is now more explicit in this respect.
 
-\item \emph{How many target sounds were assessed by each subject?}
+\item \emph{How many target sounds were assessed by each subject? Was this controlled by the study or was the study simply time limited?}
 
-All the subjects had to assess 13 target sounds.
-
-\item \emph{Was this controlled by the study or was the study simply time limited?}
-
-Heu de quoi ?
+All the subjects had to assess 13 target sounds, thus controlled by the study, see Section 4.2 of the updated manuscript.
 

### 2016-06-07 lafaySsf2016.tex
polishing response
include nicolas review, merge Fig 5 and 7, increase font size

@@
 \usepackage{color,url}
-\newcommand{\gl}[1]{\textcolor{red}{Gregoire : #1}}
+\newcommand{\gl}[1]{\textcolor{red}{gl: #1}}
 \newcommand{\mr}[1]{\textcolor{green}{Mathias : #1}}
 \newcommand{\ml}[2][]{\textcolor{blue}{#1 #2}}
-
+\newcommand{\nm}[1]{\textcolor{magenta}{nm: #1}}
 %\renewcommand{\baselinestretch}{1.5}
@@
 \email{(gregoire.lafay@irccyn.ec-nantes.fr)}
-\affil{IRCCyN, Ecole Centrale de Nantes $^1$ \quad\quad\quad\quad IRCAM, STMS, UPMC $^2$}
+\affil{IRCCyN, Ecole Centrale de Nantes $^1$ \quad\quad\quad\quad STMS Ircam-CNRS-UPMC $^2$}
 }
-
-%\author{
-%\alignauthor
-% Gregoire Lafay\\
-%       \affaddr{IRCCyN, Ecole Centrale de Nantes}\\
-%       \affaddr{1 rue de la Noe}\\
-%       \affaddr{Nantes, France} 
-%       %\email{gregoire.lafay@irccyn.ec-nantes.fr} 
-%\alignauthor 
-% Nicolas Misdarris\\
-%       \affaddr{IRCAM}\\
-%       \affaddr{1 Place Igor\-Stravinsky}\\
-%       \affaddr{Paris, France} \\
-%       \email{}
-%\and
-%\alignauthor 
-%Mathieu Lagrange\\
-%       \affaddr{IRCCyN, Ecole Centrale de Nantes}\\
-%       \affaddr{1 rue de la Noe}\\
-%       \affaddr{Nantes, France} \\
-%\alignauthor 
-% Mathias Rossignol\\
-%       \affaddr{IRCAM}\\
-%       \affaddr{1 Place Igor\-Stravinsky}\\
-%       \affaddr{Paris, France} \\
-%       \email{}
-%}
 
@@
 For the sake of comparison, we also consider a display whose spatial organization is based on acoustic cues.
-Those three displays are evaluated in terms of search speed in a crowdsourcing experiment using two different corpora: the first being composed of sounds of urban environment and the second of sounds produced by musical instruments.
-Coherent results achieved by considering the two different corpora demonstrate the usefulness of using an implicit semantic organization to display sounds, both in terms of search speed and of efficiency of learning.
+Those three displays are evaluated in terms of search speed in a crowdsourcing experiment using two different corpora: the first being composed of sounds of urban environment \nm{je mettrais plutot: "environmental sounds from urban"} and the second of sounds produced by musical instruments.
+Coherent results achieved by considering the two different corpora demonstrate the usefulness of using an implicit semantic organization to display sounds, both in terms of search speed and of \gl{learning efficiency}.
 \end{abstract}
@@
 
-With the growing capability of recording and storage devices, the problem of indexing large databases of audio has recently been the object of much attention \cite{Wold1996}.
+With the growing capability of recording and storage devices, the problem of indexing large databases of audio \nm{"audio databases" ?} has recently been the object of much attention \cite{Wold1996}.
 Most of that effort is dedicated to the automatic inference of indexing metadata from the actual audio recording \cite{Zhang1999, tzanetakis2002musical}; in contrast, the ability to browse such databases in an effective manner has been less considered.
@@
 \begin{enumerate}
-\item Sounds, as many others things, can be described in many ways. Sound may be designated by their
+\item Sounds, as many others things, can be described in many ways.
+\nm{ceci est valable pour des sons de nature "causale", i.e.
+dont la source est identifiable; pour les sons abstraits, la semantique de description est davanage de l'ordre de l'acoustique (briallnt, aigu, sourd, rond, et.) voir p.
+ex.
+Lemaitre, G., Houix, O., Misdariis, N., Susini, P., « Listener expertise and sound identification influence the categorization of environmental sounds », Journal of Experimental Psychology: Applied, 2010, vol.
+16, n° 1, pp.
+16-32} Sound may be designated by their
 sources (a car door), as well as by the action of those sources (the slamming of a car door) or their environments (slamming a car door in a garage) \cite{houix2012lexical, niessen2010categories, brown2011towards}.
@@
 Two corpora are considered for this study.
-The first one is composed of sounds recorded in a urban environment and the second one is composed of sounds of musical instruments.
+The first one is composed of sounds recorded in a urban environment and the second one is composed of sounds of musical \nm{"produced by musical ..." ?} instruments.
 
@@
 
-Similarly, the latter also has a semantic organization: high and mid level classes categorize musical instruments according to the means by which sound is produced, while leaf level classes are related to specific instruments (\textit{string} $>$ \textit{plucked-string} $>$ \textit{viola}; \textit{free-aerophone} $>$ \textit{brassy} $>$ \textit{trumpet}).
+Similarly, the latter \nm{"the latter corpus" / "the second corpus" ?} also has a semantic organization: high and mid level classes categorize musical instruments according to the means by which sound is produced, while leaf level classes are related to specific instruments (\textit{string} $>$ \textit{plucked-string} $>$ \textit{viola}; \gl{\textit{brass} $>$ \textit{ordinario} $>$ \textit{trumpet}}).
 
@@
 
-%\begin{figure}[t]
-%\begin{center}
-%\includegraphics[scale=0.18]{gfx/xp3.png} 
-%\end{center}
-%\caption{\label{figXP3} Acoustical Display (AD).}
-%\end{figure}
-
 \begin{figure}[t]
 \begin{center}
-\includegraphics[width=0.5\textwidth]{gfx/urban_mds.png} 
+\includegraphics[width=\columnwidth]{gfx/urban_mds.png} 
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{gfx/music_mds.png} 
+\includegraphics[width=\columnwidth]{gfx/music_mds.png} 
 \end{center}
-\caption{\label{figXP3music} Acoustical Display for the musical sound dataset (AD). Colors and legend are removed during the experiment.}
+\caption{\label{figXP3music} Acoustical Display for the musical sound dataset (AD). Colors and legend are removed during the experiment. \gl{remove ?}}
 \end{figure}
 
-Two semantically oriented displays are then proposed, called Full Semantic Display (FSD) and Progressive Semantic Display (PSD)%, shown in Figures~\ref{figXP2} and~\ref{figXP1}, respectively. 
+Two semantically oriented displays are then proposed, called Full Semantic Display (FSD) and Progressive Semantic Display (PSD).
 Both display  consider the hierarchical structure of the dataset to organize sounds.
@@
 
-%\begin{figure}[t!]
-%\begin{center}
-%\includegraphics[scale=0.30]{gfx/XP2clean.pdf} 
-%\end{center}
-%\caption{\label{figXP2} Full Semantic Display (FSD): implicit hierarchical organization of semantic classes.}
-%\end{figure}
-%
-%\begin{figure*}[t]
-%\begin{center}
-%\includegraphics[scale=0.4]{gfx/XP1clean2.pdf} 
-%\end{center}
-%\caption{\label{figXP1} Progressive Semantic Display (PSD):  explicit hierarchical organization of semantic classes. (a) initial folded version; (b) partly folded version; (c) unfolded version.}
-%\end{figure*}
-
 \section{Experiments} \label{test}
@@
 
-During this experiment, the three displays are compared using two corpora: the urban sound corpus and the musicla sound corpus.
+During this experiment, the three displays are compared using two corpora: the urban sound corpus and the musical sound corpus.
 By comparing AD and xSD (the two semantic displays), the goal is to evaluate the  relative efficiencies of semantic based and acoustic based spatial configurations.
@@
 The sum of the 13 search durations, which is the \textit{absolute duration} minus the break times between two target sound searches, is called the \textit{total search duration}.
-\item the name of each sound which has been heard.
+\item the name of each sound which has been heard. \nm{... and so the number of heard sound (to make the link with sect. 4.6 below)}
 \item the time at which each sound has been heard.  
@@
 Subjects are only allowed to perform the experiment once, and on one interface only.
-Data were automatically collected server-side at the end of the experiments.
+\nm{may be worth specifying how this is done : IP id tracking ?} Data were automatically collected server-side at the end of the experiments.
 Subjects were asked to use headphones.
@@
 60 subjects have completed the experiment for the urban sound dataset, 20 for each interface.
-43 subjects have completed the experiment for the urban sound dataset: 15 for the PSD interface, 13 for the FSD interface and 15 for the AD interface.
-
-\ml{to update before final submission}
+\gl{48 subjects have completed the experiment for the urban sound dataset: 17 for the PSD interface, 16 for the FSD interface and 15 for the AD interface.}
 
@@
 Post hoc analyses are done using two samples \textit{t} tests with Bonferroni correction.
-All statistical analyses are performed at the 5\% significance level.% using native functions of the MATLAB computing environment.
+All statistical analyses are performed at the 5\% significance level.
 
@@
 
-%\begin{table}[t]
-%\caption{\label{tab1} Criterion used to detect outliers.}
-%\begin{center}
-%\begin{tabular}{ccccc}
-%criterion      &  1  &  2  &  3  &  4  \\
-%\hline
-%outlier 1 (PD) & --- & --- & --- &  x  \\
-%outlier 2 (PD) &  x  & --- & --- & --- \\  
-%outlier 3 (AD) & --- & --- & --- &  x \\
-%outlier 4 (AD) &  x  &  x  &  x  & --- \\
-%outlier 5 (AD) &  x  & --- & --- & --- \\
-%\hline
-%\end{tabular}
-%\end{center}
-%\end{table}
-
-Using the $IQR$ method, 5 subjects are detected as outliers for the urban experiment, and removed from the analysis.
+Using the $IQR$ method, 5 subjects \nm{sauf erreur, le total des sujets outliers fait 7 (3+1+1+2) alors qu'il est annonce 5 en debut de paragraphe} \gl{non il y a bien 5 outliers} are detected as outliers for the urban experiment, and removed from the analysis.
 2 of those subjects used PSD and 3 subjects AD.
-%Table~\ref{tab1} shows the criteria used to detect the 5 outliers.
 3 subjects are detected by considering the \textit{absolute duration}.
 They spent respectively 50 minutes, 5 hours and 17.5 hours to complete the experiment.
-1 subject is detected by considering the \textit{total search duration} (46 minutes), and an other by observing the \textit{total number of heard sounds} (1800 heard sounds, roughly 12 times the total size of the corpus).
+1 subject is detected by considering the \textit{total search duration} (46 minutes), and another by observing the \textit{total number of heard sounds} (1800 heard sounds, roughly 12 times the total size of the corpus).
 And lastly 2 subjects are detected by observing the maximum number of times they had to hear a target sound before finding it (4 and 8 times).
@@
 \begin{center}
-\includegraphics[scale=0.4]{gfx/analyse_urban1.eps} 
+\includegraphics[width=\columnwidth]{gfx/urbanMusic1.eps} 
 \end{center}
-\caption{\label{fig1urban} Average and standard deviation for (a) the total search durations, (b) the number of heard sounds and (c) the number of unique heard sounds.}
+\caption{\label{urbanMusic} \gl{Average and standard deviation for (a,d) the total search durations, (b,e) the number of heard sounds and (c,f) the number of unique heard sounds; considering the urban dataset (a,b,c) or the music dataset (d,e,f)}}
 \end{figure}
@@
 \begin{center}
-\includegraphics[scale=0.4]{gfx/analyse_urban2.eps} 
+\includegraphics[width=\columnwidth]{gfx/analyse_urban2.eps} 
 \end{center}
-\caption{\label{fig2urban} evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search.
-Lines are linear regression fits.}
+\caption{\label{fig2urban} \gl{evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the urban dataset.
+Lines are linear regression fits.}}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=\columnwidth]{gfx/analyse_music2.eps} 
+\end{center}
+\caption{\label{fig2Music} \gl{evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the music dataset.
+Lines are linear regression fits.}}
 \end{figure}
@@
 \item the \textit{total search duration}
-\item the number of heard sounds
-\item the number of unique sounds heard. By "unique" we mean that, if a same sound is heard 10 times during the 13 searches of the  experiment, it counts only for one. 
+\item the number of heard sounds during the whole experiment
+\item the number of unique sounds heard during the whole experiment. By "unique" we mean that, if a same sound is heard 10 times during the 13 searches of the  experiment, it counts only for one. 
 \end{itemize}
@@
 
-The averages and standard deviations for the three metrics are presented in Figure~\ref{fig1urban}.
-The type of display has a significant effect on the the \textit{total search duration} ($F[2,52]=3.63; p<0.05$), the number of heard sounds ($F[2,52]=5.65; p<0.01$) and the number of unique heard sounds ($F[2,52]=8.65; p<0.01$).
+The averages and standard deviations for the three metrics are presented in Figure~\ref{urbanMusic}.
+The type of display has a significant effect on the the \textit{total search duration} ($F[2,52]=4; p<0.05$), the number of heard sounds ($F[2,52]=6; p<0.01$) and the number of unique heard sounds ($F[2,52]=9; p<0.01$).
 
-Post hoc analysis on the total search time (Figure~\ref{fig1urban}.a) shows no significant differences between the three interfaces (PSD-FSD: $p=0.09$; PSD-AD: $p=1$; FSD-AD: $p=0.07$), although the FSD interface achieves the best results.
+Post hoc analysis on the total search duration (Figure~\ref{urbanMusic}.a) shows no significant differences between the three interfaces (PSD-FSD: $p=0.09$; PSD-AD: $p=1$; FSD-AD: $p=0.07$), although the FSD interface achieves the best results.
 
-Concerning the numbers of heard sounds (Figure~\ref{fig1urban}.b), FSD significantly outperforms the AD interface (FSD-AD: $p<0.01$),  whereas PSD and FSD as well as PSD and AD show similar outcomes (PSD-FSD: $p=0.06$; PSD-AD: $p=1$).
+Concerning the numbers of heard sounds (Figure~\ref{urbanMusic}.b), FSD significantly outperforms the AD interface (FSD-AD: $p<0.01$),  whereas PSD and FSD as well as PSD and AD show similar outcomes (PSD-FSD: $p=0.06$; PSD-AD: $p=1$).
 
-Figure~\ref{fig1urban}.c shows the results for the number of unique heard sounds.
-This time the results of AD are significantly lower than those of both PSD and FSD (PSD-AD: $p<0.01$; FSD-AD: $p<0.01$).
+Figure~\ref{urbanMusic}.c shows the results for the number of unique heard sounds.
+This time the results of AD are significantly lower \nm{higher ?} than those of both PSD and FSD (PSD-AD: $p<0.01$; FSD-AD: $p<0.01$).
 For AD, 75\% of the subjects heard more than 140 sounds, and 25\% heard at least 148 sounds, that is almost the entire database.
 Considering PSD, 75\% of subjects heard less than 134 sounds, versus 143 for FSD.
-This time PSD and FSD perform equivalently (PD-FD: $p=1$).
+This time PSD and FSD perform equivalently (PSD-FSD: $p=1$).
 
@@
 As shown by a linear regression of the data, there is an overall increase of efficiency with respect to the index of the search.
-For the duration and the number of heard sounds (Figure~\ref{fig2urban}.b), the AD and PSD perform equivalently, whereas the FSD exhibit better performance.
-In the case of the number of unique heard sounds (Figure~\ref{fig2urban}.c), the two semantic displays are equivalent.
-This latter result suggest that for both semantic displays, users have parsed the same range of the dataset, although users of FSD have done so by listening to fewer sounds and more quickly.
+For the duration and the number \nm{"the number" ?} of heard sounds (Figure~\ref{fig2urban}.b), the AD and PSD perform equivalently \nm{mais pt-etre faudrait-il mentionner ici que la variation autour de la droite de regression de PSD est beaucoup plus importante ...
+?}, whereas the FSD exhibit better performance.
+In the case of the number of unique heard sounds (Figure~\ref{fig2urban}.c), the two semantic displays are equivalent \nm{idem}.
+This latter result suggests that for both semantic displays, users have parsed the same range of the dataset, although users of FSD have done so by listening to fewer sounds and more quickly.
 
 \section{Results on Musical Sounds} \label{resultsm}
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[scale=0.4]{gfx/analyse_music1.eps} 
-\end{center}
-\caption{\label{fig1Music} Average and standard deviation for (a) the total search durations, (b) the number of heard sounds and (c) the number of unique heard sounds.}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[scale=0.4]{gfx/analyse_music2.eps} 
-\end{center}
-\caption{\label{fig2Music} evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search.
-Lines are linear regression fits.}
-\end{figure}
-
-%\gl{merge fig. 5 and 7}
-% if space is need fig 1 could be merged
 
@@
 
-The averages and standard deviations for the three metrics are presented in Figure~\ref{fig1Music}.
-The type of display has a significant effect on the the \textit{total search duration} ($F[2,38]=10.25; p<0.01$), the number of heard sounds ($F[2,38]=13.15; p<0.01$) and the number of unique heard sounds ($F[2,38]=6.39; p<0.01$).
+The averages and standard deviations for the three metrics are presented in Figure~\ref{urbanMusic}.
+The type of display has a significant effect on the the \textit{total search duration} ($F[2,43]=13; p<0.01$), the number of heard sounds ($F[2,43]=14; p<0.01$) and the number of unique heard sounds ($F[2,43]=5; p<0.05$).
 
-Considering the \textit{total search duration} (Figure~\ref{fig1Music}.a), post hoc analyses shows that PSD and FSD perform better than the AD interfaces (PSD-AD: $p<0.05$; FSD-AD: $p<0.01$;).
-No significant difference is found between PSD and FSD (PSD-FSD: $p=0.32$).
+Considering the \textit{total search duration} (Figure~\ref{urbanMusic}.d), post hoc analyses shows that PSD and FSD perform better than the AD interfaces (PSD-AD: $p<0.01$; FSD-AD: $p<0.01$;).
+No significant difference is found between PSD and FSD (PSD-FSD: $p=0.20$).
 
-Figure~\ref{fig1Music}.b shows the results for the numbers of heard sounds.
-This time AD and PSD exhibit similar results (PSD-AD: $p<0.43$), while being both outperformed by FSD (FSD-PSD: $p<0.01$; FSD-AD: $p<0.01$).
+Figure~\ref{urbanMusic}.e shows the results for the numbers of heard sounds.
+This time AD and PSD exhibit similar results (PSD-AD: $p=0.19$), while being both outperformed by FSD (FSD-PSD: $p<0.01$; FSD-AD: $p<0.01$).
 
-Considering the numbers of unique heard sounds (Figure~\ref{fig1Music}.c), FSD significantly outperforms the AD interface (FSD-AD: $p<0.01$), while both PSD \emph{vs.} FSD and PSD \emph{vs.} AD show similar outcomes (PSD-FSD: $p=0.6$; PSD-AD: $p=0.9$).
+Considering the numbers of unique heard sounds (Figure~\ref{urbanMusic}.f), FSD significantly outperforms the AD interface (FSD-AD: $p<0.05$), while both PSD \emph{vs.} FSD and PSD \emph{vs.} AD show similar outcomes (PSD-FSD: $p=0.06$; PSD-AD: $p=1$).
 Those results confirms the trends observed for the urban sounds (see Section~\ref{resultsu}):
@@
 
-Considering that the musical corpus have perhaps a obvious semantic organization, PSD and FSD significantly reduces the search durations compared to an acoustic organization (AD), which was not the case for the urban sound dataset.
+Considering that the musical corpus have perhaps an obvious semantic organization, PSD and FSD significantly reduces the search durations compared to an acoustic organization (AD), which was not the case for the urban sound dataset.
 
@@
 \section{Conclusion}
+
+\nm{Cela manque pt-etre d'une partie Discussion / Synthese ? Et la conclusion manque pt-etre un peu de perspectives : vers simScene, par exemple ?}
 

### 2016-06-08 lafaySsf2016.tex
Quelques corrections, quelques reformulations, quelques "raccourcissements".
Je me suis permis d'enlever les commentaires "visibles" (y compris ceux de NM, dont j'ai appliqué la plupart des suggestions mais pas toutes) pour qu'on puisse mieux voir la longueur.

Quelques commentaires restants, mis en commentaire dans le .tex :

L. 185 :

% MR : S'il y a besoin de gagner de la place, retirer le paragraphe suivant, ça n'apporte pas grand chose

During the experiment, two visual cues are shown to the subject:
\begin{itemize}
\item \textit{The search state}: which can be "pause" if the subject is n

@@
 For the sake of comparison, we also consider a display whose spatial organization is based on acoustic cues.
-Those three displays are evaluated in terms of search speed in a crowdsourcing experiment using two different corpora: the first being composed of sounds of urban environment \nm{je mettrais plutot: "environmental sounds from urban"} and the second of sounds produced by musical instruments.
-Coherent results achieved by considering the two different corpora demonstrate the usefulness of using an implicit semantic organization to display sounds, both in terms of search speed and of \gl{learning efficiency}.
+Those three displays are evaluated in terms of search speed in a crowdsourcing experiment using two different corpora: the first is composed of environmental sounds from urban environments and the second of sounds produced by musical instruments.
+Coherent results achieved by considering the two different corpora demonstrate the usefulness of using an implicit semantic organization to display sounds, both in terms of search speed and of learning efficiency.
 \end{abstract}
@@
 
-With the growing capability of recording and storage devices, the problem of indexing large databases of audio \nm{"audio databases" ?} has recently been the object of much attention \cite{Wold1996}.
+With the growing capability of recording and storage devices, the problem of indexing large audio databases has recently been the object of much attention \cite{Wold1996}.
 Most of that effort is dedicated to the automatic inference of indexing metadata from the actual audio recording \cite{Zhang1999, tzanetakis2002musical}; in contrast, the ability to browse such databases in an effective manner has been less considered.
 
-Most media assets management are based on keyword-driven queries.
+Most media asset management systems are based on keyword-driven queries.
 The user enters a word which best characterizes the desired item, and the interface presents him with items related to this word.
 The effectiveness of this principle is primarily based on the typological structure and nomenclature of the database.
-However, for databases and more specifically for databases of sounds, several issues arise:
+However, for sound databases, several issues arise:
 
 \begin{enumerate}
-\item Sounds, as many others things, can be described in many ways.
-\nm{ceci est valable pour des sons de nature "causale", i.e.
-dont la source est identifiable; pour les sons abstraits, la semantique de description est davanage de l'ordre de l'acoustique (briallnt, aigu, sourd, rond, et.) voir p.
-ex.
-Lemaitre, G., Houix, O., Misdariis, N., Susini, P., « Listener expertise and sound identification influence the categorization of environmental sounds », Journal of Experimental Psychology: Applied, 2010, vol.
-16, n° 1, pp.
-16-32} Sound may be designated by their
-sources (a car door), as well as by the action of those sources (the slamming of a car door) or their environments (slamming a car door in a garage) \cite{houix2012lexical, niessen2010categories, brown2011towards}.
+\item Sounds, as many others things, can be described in many ways. Environmental sounds in particular may be designated by their
+source (a car door),  the action of that source (the slamming of a car door) or their environment (slamming a car door in a garage) \cite{houix2012lexical, niessen2010categories, brown2011towards}.
 Designing an effective keyword-based search system requires an accurate description of each sound, which has to be tunable to the sound representation of each user.
@@
 Their effectiveness is studied through a search-based task whose aim is to find a given target sound by browsing the database using the display under evaluation.
-Two corpora are considered for this study.
-The first one is composed of sounds recorded in a urban environment and the second one is composed of sounds of musical \nm{"produced by musical ..." ?} instruments.
+Two corpora are considered for this study: the first one is composed of sounds recorded in a urban environment, and the second of sounds produced by musical instruments.
 
 The paper is organized as follows: in Section~\ref{previous}, previous work on the topic of sound database browsing is reviewed.
-Then, the two corpora used in this study are described in Section~\ref{dataset}.
+The two corpora used in this study are then introduced in Section~\ref{dataset}.
 The three displays under evaluation are next described in Section~\ref{display}, and the crowdsourcing test used to compare them using the two different corpora is presented in Section~\ref{test}.
@@
 Typically, the items (sound effects or pieces of music) are modeled by processing the digital audio waveform through some signal processing pipeline in order to get a compact description of each item
-\cite{coleman2007mused} with sometime additional textual features.
+\cite{coleman2007mused}, with sometimes additional textual features.
 Then, statistical projections or embedding techniques, like Principal Components Analysis (PCA) \cite{MusicBox},  Multi Dimensional Scaling (MDS) \cite{schwarz2009sound,Cano2002}, Self Organizing Maps (SOM) \cite{pampalk2004exploring,pampalk2006musicrainbow,knees2006innovative} and the like are used to project the items in a two or three dimensional space while preserving as much as possible the distances among them.
@@
 
-The datasets considered in this study are respectively composed of 149 urban environmental sound events and 137 musical sounds produced by musical instruments.
+The datasets considered in this study are respectively composed of 149 urban environmental sound events and 137 sounds produced by musical instruments.
 
-The former has a semantic organization as follows: a sound is characterized by a tag describing the physical source of the sound (\textit{man-yelling}, \textit{car-passing}).
+In the first corpus, a sound is characterized by a tag describing the physical source of the sound (\textit{man-yelling}, \textit{car-passing}).
 Sounds are then hierarchically grouped into classes according to their tags (\textit{car} $>$ \textit{car-passing}; \textit{car} $>$ \textit{car-starting}).
 Those classes are in turn packed into classes until high level classes describing broader concepts are reached (\textit{traffic} $>$ \textit{car} $>$ \textit{car-passing}).
-The sound dataset is thus organized into a hierarchical structure of semantic classes as described in Figure~\ref{figdataset}.
-Strictly speaking, the sound samples of the dataset are the leaf semantic levels.
+The sound dataset is thus organized into a hierarchical structure of semantic classes as described in Figure~\ref{figdataset}, where the sound samples of the dataset are the leaf semantic levels.
 All the other classes are represented by a \textit{prototype sound} that best characterizes the sounds belonging to the class.
@@
 
-Similarly, the latter \nm{"the latter corpus" / "the second corpus" ?} also has a semantic organization: high and mid level classes categorize musical instruments according to the means by which sound is produced, while leaf level classes are related to specific instruments (\textit{string} $>$ \textit{plucked-string} $>$ \textit{viola}; \gl{\textit{brass} $>$ \textit{ordinario} $>$ \textit{trumpet}}).
+Similarly, the second corpus also has a semantic organization: high and mid level classes categorize musical instruments according to the means by which sound is produced, while leaf level classes are related to specific instruments (\textit{string} $>$ \textit{plucked-string} $>$ \textit{viola}; \gl{\textit{brass} $>$ \textit{ordinario} $>$ \textit{trumpet}}).
 
@@
 
-The aim of the displays described in this section is to allow the user to efficiently browse the above presented dataset of sounds without any written textual help.
+The aim of the displays described in this section is to allow the user to efficiently browse the sound datasets without any written textual help.
 In all displays, each sound is graphically represented by a filled circle which, once clicked, plays the actual sound.
@@
 The Euclidean distance between time averaged features is then computed for each sound pair.
-A non metric multidimensional scaling (MDS) with Kruskal's normalized stress \cite{kruskal1964multidimensional} is then employed to project the data into two dimensions according to this distance, see Figures \ref{figXP3urban} \ref{figXP3music}.
+A non metric multidimensional scaling (MDS) with Kruskal's normalized stress \cite{kruskal1964multidimensional} is then employed to project the data into two dimensions according to this distance, as shown on Figures~\ref{figXP3urban} and~\ref{figXP3music}.
 The latter exhibit a clear distinction between percussive (left) and sustained sounds (right).
 
-Among the several alternative that exists, MFCCs features are considered as they are widely used in the audio processing community.
-The MDS is selected as it has lower sensitivity to outliers than the PCA and have the advantage of not relying on any parameter setting unlike the SOM approach.
+Among the several alternatives that exists, MFCCs features are considered as they are widely used in the audio processing community.
+MDS is selected for 2D projection as it has lower sensitivity to outliers than the PCA and has the advantage of not relying on any parameter setting unlike the SOM approach.
 
@@
 Circle packing functions of the D3.js (Data-Driven Documents) javascript library \cite{2011-d3} are used to distribute the sound classes in the space.
-Depending on the display, the user can either access each leaf class directly, or has to click through intermediate levels of the hierarchy.
+Depending on the display, the user can either access each leaf class directly (FSD), or has to click through intermediate levels of the hierarchy (PSD).
 
@@
 When users click on a circle, they hear the sound prototype of the class and the subclasses are progressively revealed, represented by white circles contained in the circle that has just been clicked.
-The same action repeats itself until the user reaches the leaf classes of the hierarchy.
+The same action is repeated until the user reaches the leaf classes of the hierarchy.
 The leaf classes are represented with small gray circles, indicating that there is no subclass to uncover.
@@
 In this display, the leaf classes are distributed  in the same manner as FSD.
-Thus, the spatial configuration of the unfolded version of PSD, which may be obtained after discovering all the classes and subclasses, is equivalent to the one of FSD.
+Thus, the spatial configuration of the unfolded version of PSD, which may be obtained after discovering all the classes and subclasses, is identical to the one of FSD.
 
-The latter interface is introduced in order to study if a progressive top-down display of the hierarchical structure helps the user explore and understand the dataset hierarchy.
+\mr{À enlever si vous êtes d'accord, c'est répété juste en-dessous : The PSD interface is introduced in order to study if a progressive top-down display of the hierarchical structure helps the user explore and understand the dataset hierarchy.}
 
@@
 
-We evaluate and compare those three displays with a crowdsourcing web-based experiment \footnote{The tests are available at \url{http://soundthings.org/research/speedSoundFinding/index.html} and \url{http://soundthings.org/research/speedSoundFindingMusic/index.html}, each of the 3 displays is also available by appending a digit from 1 to 3 to \textit{index} in the above cited url.} for which we adopt a between subject approach, \textit{i.e.} a subject can only test one display.
+We evaluate and compare those three displays with a crowdsourcing web-based experiment\footnote{The tests are available at \url{http://soundthings.org/research/speedSoundFinding/index.html} and \url{http://soundthings.org/research/speedSoundFindingMusic/index.html}, each of the 3 displays is also available by appending a digit from 1 to 3 to \textit{index} in the above cited url.} for which we adopt a between subject approach, \textit{i.e.} a subject can only test one display.
 In this experiment, subjects are asked to successively retrieve 13 target sounds among the whole dataset.
 The target sounds are selected such that there are at least two target sounds in each top-level class of the semantic hierarchical structure of the dataset.
-To reduce order effects, target sounds are presented to each subject in random order.
+To reduce order effects, target sounds are presented to each subject in a random order.
 
-First, the subject clicks a \textit{Play target sound} button to listen to the target sound.
+First, the subject clicks a button to listen to the target sound.
 They may then replay the target sound as many times as they like, even while browsing.
@@
 This action 1) stops the timer and 2) loads a new target sound.
-If the subject designates and incorrect sound, an error message appears, and the experiment goes on with the same target sound.
+If the subject designates an incorrect sound, an error message appears, and the experiment goes on with the same target sound.
+
+% MR : S'il y a besoin de gagner de la place, retirer le paragraphe suivant, ça n'apporte pas grand chose
 
@@
 The experiment ends when all the target sounds have been found.
-One shall notice that the PSD display is not reintialized at each change of the target sound.
-Thus, when a circle is unfolded, it remains visible during the whole experiment.
+An important feature of the PSD display is that it is not reinitialized at each change of target sound.
+Thus, when a circle is unfolded, its content remains visible during the whole experiment.
 
@@
 \begin{itemize}
-\item the total duration of the entire experiment. It includes breaks between two target sound searches and it is called the \textit{absolute duration}.
+\item the \textit{absolute duration}, which is the total duration of the entire experiment, including breaks between two target sound searches;
 \item the duration of each search.
-The sum of the 13 search durations, which is the \textit{absolute duration} minus the break times between two target sound searches, is called the \textit{total search duration}.
-\item the name of each sound which has been heard. \nm{... and so the number of heard sound (to make the link with sect. 4.6 below)}
+The sum of the 13 search durations, which is the \textit{absolute duration} minus the break times between two target sound searches, is called the \textit{total search duration};
+\item the name of each sound which has been listened to (from which the number of heard sounds can be deduced);
 \item the time at which each sound has been heard.  
@@
 The experiment is designed to be performed using the \textit{Chrome} web browser.
-The link to the experiment has been sent to three  mailing lists, namely \textit{Music-Ir, Auditory} and an internal IRCAM mailing list.
+The link to the experiment has been sent to three  mailing lists, namely \textit{Music-Ir}, \textit{Auditory} and an internal IRCAM mailing list.
 Subjects are only allowed to perform the experiment once, and on one interface only.
-\nm{may be worth specifying how this is done : IP id tracking ?} Data were automatically collected server-side at the end of the experiments.
+Data were automatically collected server-side at the end of the experiments.
 Subjects were asked to use headphones.
 All sounds of the dataset are normalized to the same root mean square (RMS) level.
+
+% Je ne pense pas utile de préciser la technique utilisée pour empêcher les gens de faire l'expé deux fois, parce que (1) ce n'est pas fondamental, et (2) la technique est loin d'être parfaite.
 
@@
 60 subjects have completed the experiment for the urban sound dataset, 20 for each interface.
-\gl{48 subjects have completed the experiment for the urban sound dataset: 17 for the PSD interface, 16 for the FSD interface and 15 for the AD interface.}
+48 subjects have completed the experiment for the urban sound dataset: 17 for the PSD interface, 16 for the FSD interface and 15 for the AD interface.
 
@@
 \begin{enumerate}
-\item the \textit{absolute duration}: to detect subjects who took abnormally long breaks between searches.
-\item the \textit{total search duration}: to detect subjects who spent an abnormally long time to complete the experiment.
-\item number of heard sounds: to detect subjects who had to hear an abnormally high number of sounds to complete the experiment.
+\item the \textit{absolute duration}: to detect subjects who took abnormally long breaks between searches;
+\item the \textit{total search duration}: to detect subjects who spent an abnormally long time to complete the experiment;
+\item number of heard sounds: to detect subjects who had to hear an abnormally high number of sounds to complete the experiment;
 \item the maximum number of times a target sound had to be heard before being found: to detect subjects who had difficulty to recognize the target sounds.
@@
 
-Using the $IQR$ method, 5 subjects \nm{sauf erreur, le total des sujets outliers fait 7 (3+1+1+2) alors qu'il est annonce 5 en debut de paragraphe} \gl{non il y a bien 5 outliers} are detected as outliers for the urban experiment, and removed from the analysis.
-2 of those subjects used PSD and 3 subjects AD.
-3 subjects are detected by considering the \textit{absolute duration}.
-They spent respectively 50 minutes, 5 hours and 17.5 hours to complete the experiment.
-1 subject is detected by considering the \textit{total search duration} (46 minutes), and another by observing the \textit{total number of heard sounds} (1800 heard sounds, roughly 12 times the total size of the corpus).
-And lastly 2 subjects are detected by observing the maximum number of times they had to hear a target sound before finding it (4 and 8 times).
+Using the $IQR$ method, 5 subjects are detected as outliers for the urban experiment, and removed from the analysis.
+2 of those subjects used PSD and 3 used AD.
+3 subjects are detected by considering the \textit{absolute duration}; they spent respectively 50 minutes, 5 hours and 17.5 hours to complete the experiment.
+One subject is detected by considering the \textit{total search duration} (46 minutes), and another by observing the \textit{total number of heard sounds} (1800 heard sounds, roughly 12 times the total size of the corpus).
+Lastly, 2 subjects are detected by observing the maximum number of times they had to hear a target sound before finding it (4 and 8 times).
+It should be noted that some subjects met several of the criteria.
 
@@
 \begin{itemize}
-\item the \textit{total search duration}
-\item the number of heard sounds during the whole experiment
+\item the \textit{total search duration},
+\item the number of heard sounds during the whole experiment,
 \item the number of unique sounds heard during the whole experiment. By "unique" we mean that, if a same sound is heard 10 times during the 13 searches of the  experiment, it counts only for one. 
@@
 Figure~\ref{urbanMusic}.c shows the results for the number of unique heard sounds.
-This time the results of AD are significantly lower \nm{higher ?} than those of both PSD and FSD (PSD-AD: $p<0.01$; FSD-AD: $p<0.01$).
+This time the results of AD are significantly higher (indicating an inferior performance) than those of both PSD and FSD (PSD-AD: $p<0.01$; FSD-AD: $p<0.01$).
 For AD, 75\% of the subjects heard more than 140 sounds, and 25\% heard at least 148 sounds, that is almost the entire database.
-Considering PSD, 75\% of subjects heard less than 134 sounds, versus 143 for FSD.
+Considering PSD, 75\% of the subjects heard less than 134 sounds, versus 143 for FSD.
 This time PSD and FSD perform equivalently (PSD-FSD: $p=1$).
@@
 As shown by a linear regression of the data, there is an overall increase of efficiency with respect to the index of the search.
-For the duration and the number \nm{"the number" ?} of heard sounds (Figure~\ref{fig2urban}.b), the AD and PSD perform equivalently \nm{mais pt-etre faudrait-il mentionner ici que la variation autour de la droite de regression de PSD est beaucoup plus importante ...
-?}, whereas the FSD exhibit better performance.
-In the case of the number of unique heard sounds (Figure~\ref{fig2urban}.c), the two semantic displays are equivalent \nm{idem}.
+For the duration and the number of heard sounds (Figure~\ref{fig2urban}.b), AD and PSD perform equivalently, whereas FSD exhibits better performance.
+In the case of the number of unique heard sounds (Figure~\ref{fig2urban}.c), the two semantic displays are equivalent.
 This latter result suggests that for both semantic displays, users have parsed the same range of the dataset, although users of FSD have done so by listening to fewer sounds and more quickly.
@@
 
-Considering that the musical corpus have perhaps an obvious semantic organization, PSD and FSD significantly reduces the search durations compared to an acoustic organization (AD), which was not the case for the urban sound dataset.
+PSD and FSD significantly reduce the search durations compared to an acoustic organization (AD), which was not the case for the urban sound dataset.
+This may be explained by the fact that the musical corpus has perhaps a more obvious semantic organization.
 
@@
 Considering the number of unique heard sounds (Figure~\ref{fig2Music}.c), the learning of the dataset range seems to be quicker for the FSD interface than for the PSD interface.
-Again, results show that enforcing the parsing of the hierarchy is not beneficial.
+Once again, results show that enforcing the parsing of the hierarchy is not beneficial.
 
@@
 
-\nm{Cela manque pt-etre d'une partie Discussion / Synthese ? Et la conclusion manque pt-etre un peu de perspectives : vers simScene, par exemple ?}
+% \nm{Cela manque pt-etre d'une partie Discussion / Synthese ? Et la conclusion manque pt-etre un peu de perspectives : vers simScene, par exemple ?}
+% MR : Cela manque peut-être, mais bon, 8 pages...
 
@@
 Those two semantic displays are assessed and compared to a third display in which the spatial configuration depends upon acoustic features.
-The tests consist in data retrieval tasks done using two different corpora.
+The tests consists in data retrieval tasks done using two different corpora.
 The Full Semantic Display (FSD), that allows users to directly visualize the leaf classes of the semantic hierarchical structure, proves to be the most effective interface for the task.

### 2016-06-08 responseToReviewers.tex
revision version

@@
 
-\ml{TODO: put analysis data online}
+%\ml{TODO: put analysis data online}
 
@@
 
-This is an excellent suggestion that we decided to do by considering an alternative dataset of sounds produced by musical instruments.
+This is an excellent suggestion.
+We consider an new dataset of sounds produced by musical instruments.
 As shown on Section 6 of the updated paper, the findings are consistent with the ones obtained using the urban sound dataset.
 
-We thus keep the original title.
+We thus choose to stay with the original title.
 
@@
 
-Even if this complementary study would have increase the level of knowledge, we decided to focus the complementary experiments to the use of an alternative dataset.
+Even if this complementary study would have increase the level of knowledge, we again decided to focus the complementary experiments to the use of an alternative dataset.
 

### 2016-06-08 lafaySsf2016.tex
revision version

@@
 \newcommand{\nm}[1]{\textcolor{magenta}{nm: #1}}
-%\renewcommand{\baselinestretch}{1.5}
+%\renewcommand{\baselinestretch}{2}
 
@@
 We present and study the efficiency of two sound presentation strategies that organize sounds spatially so as to reflect an underlying semantic hierarchy.
-For the sake of comparison, we also consider a display whose spatial organization is based on acoustic cues.
+For the sake of comparison, we also consider a display whose spatial organization is only based on acoustic cues.
 Those three displays are evaluated in terms of search speed in a crowdsourcing experiment using two different corpora: the first is composed of environmental sounds from urban environments and the second of sounds produced by musical instruments.
@@
 
-Most media asset management systems are based on keyword-driven queries.
+Thus, most media asset management systems are based on keyword-driven queries.
 The user enters a word which best characterizes the desired item, and the interface presents him with items related to this word.
@@
 source (a car door),  the action of that source (the slamming of a car door) or their environment (slamming a car door in a garage) \cite{houix2012lexical, niessen2010categories, brown2011towards}.
-Designing an effective keyword-based search system requires an accurate description of each sound, which has to be tunable to the sound representation of each user.
+Designing an effective keyword-based search system requires an accurate description of each sound, which has to be suited to the sound representation of the user to be effective.
 \item Pre-defined verbal descriptions of the sounds made available to the users may potentially bias their browsing and final selection.
@@
 
-Similarly, the second corpus also has a semantic organization: high and mid level classes categorize musical instruments according to the means by which sound is produced, while leaf level classes are related to specific instruments (\textit{string} $>$ \textit{plucked-string} $>$ \textit{viola}; \gl{\textit{brass} $>$ \textit{ordinario} $>$ \textit{trumpet}}).
+Similarly, the second corpus also has a semantic organization: high and mid level classes categorize musical instruments according to the means by which sound is produced, while leaf level classes are related to specific instruments (\textit{string} $>$ \textit{plucked-string} $>$ \textit{viola}; \textit{brass} $>$ \textit{ordinario} $>$ \textit{trumpet}).
 
@@
 
-Among the several alternatives that exists, MFCCs features are considered as they are widely used in the audio processing community.
-MDS is selected for 2D projection as it has lower sensitivity to outliers than the PCA and has the advantage of not relying on any parameter setting unlike the SOM approach.
+Among the several alternatives that exists, MFCCs features are chosen as they are widely used in the audio processing community.
+MDS is selected for 2D projection as it has a lower sensitivity to outliers than the PCA and do not relyi on any parameter optimization, unlike the SOM approach.
 
@@
 \end{center}
-\caption{\label{figXP3music} Acoustical Display for the musical sound dataset (AD). Colors and legend are removed during the experiment. \gl{remove ?}}
+\caption{\label{figXP3music} Acoustical Display for the musical sound dataset (AD). Colors and legend are removed during the experiment.}
 \end{figure}
@@
 
-\mr{À enlever si vous êtes d'accord, c'est répété juste en-dessous : The PSD interface is introduced in order to study if a progressive top-down display of the hierarchical structure helps the user explore and understand the dataset hierarchy.}
+%\mr{À enlever si vous êtes d'accord, c'est répété juste en-dessous : The PSD interface is introduced in order to study if a progressive top-down display of the hierarchical structure helps the user explore and understand the dataset hierarchy.}
 
@@
 
-During the experiment, two visual cues are shown to the subject:
-\begin{itemize}
-\item \textit{The search state}: which can be "pause" if the subject is not currently looking for a target sound (at the beginning of the experiment, or between two target sounds) or  "in progress"  if they are.
-\item \textit{Remaining target sounds}: the number of target sounds that remain to be found.
-\end{itemize}
+%During the experiment, two visual cues are shown to the subject:
+%\begin{itemize}
+%\item \textit{The search state}: which can be "pause" if the subject is not currently looking for a target sound (at the beginning of the experiment, or between two target sounds) or  "in progress"  if they are.
+%\item \textit{Remaining target sounds}: the number of target sounds that remain to be found.
+%\end{itemize}
 
@@
 A crowdsourcing approach is adopted. 
-The experiment is designed to be performed using the \textit{Chrome} web browser.
+%The experiment is designed to be performed using the \textit{Chrome} web browser.
 The link to the experiment has been sent to three  mailing lists, namely \textit{Music-Ir}, \textit{Auditory} and an internal IRCAM mailing list.
-Subjects are only allowed to perform the experiment once, and on one interface only.
-Data were automatically collected server-side at the end of the experiments.
+Subjects are recruited using research mailing list and are only allowed to perform the experiment once, and on one interface only.
+% Data were automatically collected server-side at the end of the experiments.
 Subjects were asked to use headphones.
@@
 \end{center}
-\caption{\label{urbanMusic} \gl{Average and standard deviation for (a,d) the total search durations, (b,e) the number of heard sounds and (c,f) the number of unique heard sounds; considering the urban dataset (a,b,c) or the music dataset (d,e,f)}}
+\caption{\label{urbanMusic} Average and standard deviation for (a,d) the total search durations, (b,e) the number of heard sounds and (c,f) the number of unique heard sounds; considering respectively the urban dataset (a,b,c) and the music dataset (d,e,f).}
 \end{figure}
@@
 \end{center}
-\caption{\label{fig2urban} \gl{evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the urban dataset.
-Lines are linear regression fits.}}
+\caption{\label{fig2urban} Evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the urban dataset.
+Lines are linear regression fits.}
 \end{figure}
@@
 \end{center}
-\caption{\label{fig2Music} \gl{evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the music dataset.
-Lines are linear regression fits.}}
+\caption{\label{fig2Music} Evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the music dataset.
+Lines are linear regression fits.}
 \end{figure}
@@
 According to those results, a spatial display following a hierarchical organization of the dataset based on semantic values (PSD and FSD) is more efficient than an organization based on acoustic descriptors (AD) as it allows the users to quickly apprehend the dataset range and thus to specify their search.
-Furthermore the FSD interface allows the users to retrieve the 13 target sounds by listening to a smaller amount of sounds, than the AD.
+Furthermore the FSD interface allows the users to retrieve the 13 target sounds by listening to a smaller amount of sounds than the AD.
 However, those two effects are significantly compromised when users have to parse the entire hierarchy to reach the first target sound, as in the PSD display.
@@
 The Full Semantic Display (FSD), that allows users to directly visualize the leaf classes of the semantic hierarchical structure, proves to be the most effective interface for the task.
+This display has been integrated as a sound selector into a acoustic scene simulator that benefit from the advantages of such an approach: no need of localization and reduction of the potential bias of using description tags\footnote{\url{http://soundthings.org/research/urbanSoundScape/XP2014/simScene}}.
 
@@
 
-In the datasets considered in this study, there are as many leaf classes as sounds in the dataset, which would be unrealistic for large datasets.
-The organization can easily be adapted by considering the leaf classes as \textit{collections} of semantically similar sound samples.
+In the datasets considered in this study, there are as many leaf classes as sounds in the dataset, which would be unrealistic for larger datasets.
+In that case, the organization can easily be adapted by considering the leaf classes as \textit{collections} of semantically similar sound samples.
 Thus, two sounds of \textit{male-yelling} would be grouped into a single leaf class with the tag \textit{male-yelling}.

### 2016-07-05 lafaySsf2016.tex
photos and try to bio ;)

@@
 
+\begin{figure}[t]
+\begin{center}
+\includegraphics[scale=0.24]{gfx/dataset.pdf} 
+\end{center}
+\caption{\label{figdataset} Semantic hierarchical structure of the dataset of urban environmental sounds.}
+\end{figure}
+
 Most of the research effort in the sound design and Music Information Retrieval (MIR) communities is focused on acoustical based indexing and browsing \cite{tzanetakis2003musescape, streich2008music,goto2005musicream,pampalk2007musicsun}.
@@
 \section{Datasets} \label{dataset}
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[scale=0.24]{gfx/dataset.pdf} 
-\end{center}
-\caption{\label{figdataset} Semantic hierarchical structure of the dataset of urban environmental sounds.}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[scale=0.24]{gfx/SSF.pdf} 
-\end{center}
-\caption{\label{figSSF} Semantic displays are  based on the semantic hierarchical structure of the dataset.}
-\end{figure}
 
@@
 
+\begin{figure}[t]
+\begin{center}
+\includegraphics[scale=0.24]{gfx/SSF.pdf} 
+\end{center}
+\caption{\label{figSSF} Semantic displays are  based on the semantic hierarchical structure of the dataset.}
+\end{figure}
+
 %We choose to represent each sound leaf class of our sound data set with a circle.
@@
 The Euclidean distance between time averaged features is then computed for each sound pair.
-A non metric multidimensional scaling (MDS) with Kruskal's normalized stress \cite{kruskal1964multidimensional} is then employed to project the data into two dimensions according to this distance, as shown on Figures~\ref{figXP3urban} and~\ref{figXP3music}.
+A non metric multidimensional scaling (MDS) with Kruskal's normalized stress \cite{kruskal1964multidimensional} is then employed to project the data into two dimensions according to this distance, as shown on Figure~\ref{figXP3music}.
 The latter exhibit a clear distinction between percussive (left) and sustained sounds (right).
+
+% s~\ref{figXP3urban} and
 
@@
 
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=\columnwidth]{gfx/urban_mds.png} 
-\end{center}
-\caption{\label{figXP3urban} Acoustical Display for the urban sound dataset (AD). Colors and legend are removed during the experiment.}
-\end{figure}
+%\begin{figure}[t]
+%\begin{center}
+%\includegraphics[width=\columnwidth]{gfx/urban_mds.png} 
+%\end{center}
+%\caption{\label{figXP3urban} Acoustical Display for the urban sound dataset (AD). Colors and legend are removed during the experiment.}
+%\end{figure}
 
@@
 
+\biography{\textbf{Gr\'egoire Lafay}}{gfx/lafay.jpg}{IRCCYN, Ecole Centrale de Nantes, France}
+
+\biography{\textbf{Nicolas Misdariis}}{gfx/lafay.jpg}{IRCAM, France}
+
+\biography{\textbf{Mathieu Lagrange}}{gfx/lagrange.jpg}{IRCCYN, Ecole Centrale de Nantes, France}
+
+\biography{\textbf{Mathias Rossignol}}{gfx/rossignol.jpg}{IRCAM, France}
+
+%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lafay}}]{Gr\'egoire Lafay} was born in Paris, France, in 1990.
+He received in 2011 the B.S.
+degree in Acoustic from the University Pierre and Marie Curie (UPMC), Paris, France, and the B.S.
+degree in Musicology from the Sorbonne University, Paris, France.
+He received his M.S.
+degree in acoustics, signal processing and musical informatics (ATIAM) from the University Pierre and Marie Curie and the IRCAM laboratory, Paris, France, in 2013.
+Since 2013, he is a Ph.D student at Irccyn, a French laboratory dedicated to cybernetics.
+His current research interests include acoustic scene similarity and classification, as well as acoustic scene perception.
+%\end{biography}
+%
+%\begin{biography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at Irccyn, a French laboratory dedicated to cybernetics.
+He obtained his PhD in computer science at the University of Bordeaux in 2004, and visited several institutions, in Canada (University of Victoria, McGill University) and in France (Orange Labs, Telecom ParisTech, Ircam).
+His research focuses on machine listening algorithms applied to the analysis of musical and environmental audio.
+%\end{biography}
+%
+%\begin{biography}[{\includegraphics[width=1.05in,height=1.5in,clip,keepaspectratio]{gfx/rossignol}}]{Mathias Rossignol} obtained his PhD from Rennes University (France) in 2005, with a specialty in Natural Language Processing.
+He then focused his research on speech analysis and recognition at the International Research Center MICA (Hanoi, Vietnam), before moving on to Auditory Scene Analysis at IRCAM (Institute for Acoustic / Music Research and Coordination, Paris, France).
+%\end{biography}
+%
+%\begin{biography}[{\includegraphics[width=0.9in,height=1.4in,clip,keepaspectratio]{gfx/lagrange.png}}]
+%{Emmanouil Benetos} (S'09-M'12) received the B.Sc. and M.Sc. degrees in 
+%informatics from the Aristotle University of Thessaloniki, Greece, in 
+%2005 and 2007, respectively, and the Ph.D. degree in electronic 
+%engineering from Queen Mary University of London, U.K., in 2012. From 
+%2013 to 2015, he was a University Research Fellow with the Department of 
+%Computer Science, City University London, London, U.K. He is currently a Royal 
+%Academy of Engineering Research Fellow with the Centre for Digital Music, 
+%Queen Mary University of London, U.K. His research focuses on signal 
+%processing and machine learning for music and audio analysis, as well as 
+%applications to music information retrieval, acoustic scene analysis, 
+%and computational musicology.
+%\end{biography}
+
 %In particular, much consideration has been given in order to avoid any semantic bias.

### 2016-07-05 lafaySsf2016.tex
Derniere relecture, Mathias RossignolL

@@
 The first one, considered as a reference baseline, does not rely on any semantic information and positions sounds according to their acoustical properties (time averaged spectral features).
-The second and third displays exhibit a spatial organization based  on a predefined hierarchical semantic organization; those last two displays differs by how semantics are used.
+The second and third displays exhibit a spatial organization based  on a predefined hierarchical semantic organization; those last two displays differ by how semantics are used.
 
@@
 A non metric multidimensional scaling (MDS) with Kruskal's normalized stress \cite{kruskal1964multidimensional} is then employed to project the data into two dimensions according to this distance, as shown on Figure~\ref{figXP3music}.
-The latter exhibit a clear distinction between percussive (left) and sustained sounds (right).
+In the case of instruments sounds, it exhibit a clear distinction between percussive (left) and sustained sounds (right).
 
@@
 Among the several alternatives that exists, MFCCs features are chosen as they are widely used in the audio processing community.
-MDS is selected for 2D projection as it has a lower sensitivity to outliers than the PCA and do not relyi on any parameter optimization, unlike the SOM approach.
+MDS is selected for 2D projection as it has a lower sensitivity to outliers than the PCA and does not rely on any parameter optimization, unlike the SOM approach.
 
@@
 \end{center}
-\caption{\label{figXP3music} Acoustical Display for the musical sound dataset (AD). Colors and legend are removed during the experiment.}
+\caption{\label{figXP3music} Acoustical Display for the instrument sound dataset (AD). Colors and legend are removed during the experiment.}
 \end{figure}
@@
 Two semantically oriented displays are then proposed, called Full Semantic Display (FSD) and Progressive Semantic Display (PSD).
-Both display  consider the hierarchical structure of the dataset to organize sounds.
+Both displays consider the hierarchical structure of the dataset to organize sounds.
 Each sound class is represented by a filled circle.
@@
 Thus, subclasses belonging to the same class are close to each others.
-Circle packing functions of the D3.js (Data-Driven Documents) javascript library \cite{2011-d3} are used to distribute the sound classes in the space.
+Circle packing functions of the \emph{D3.js} (Data-Driven Documents) javascript library \cite{2011-d3} are used to distribute the sound classes in the space.
 Depending on the display, the user can either access each leaf class directly (FSD), or has to click through intermediate levels of the hierarchy (PSD).
@@
 
-During this experiment, the three displays are compared using two corpora: the urban sound corpus and the musical sound corpus.
+During this experiment, the three displays are compared using two corpora: the urban sound corpus and the instrument sound corpus.
 By comparing AD and xSD (the two semantic displays), the goal is to evaluate the  relative efficiencies of semantic based and acoustic based spatial configurations.
@@
 In this experiment, subjects are asked to successively retrieve 13 target sounds among the whole dataset.
-The target sounds are selected such that there are at least two target sounds in each top-level class of the semantic hierarchical structure of the dataset.
+The target sounds are selected so that there are at least two target sounds in each top-level class of the semantic hierarchical structure of the dataset.
 To reduce order effects, target sounds are presented to each subject in a random order.
@@
 
-60 subjects have completed the experiment for the urban sound dataset, 20 for each interface.
-48 subjects have completed the experiment for the urban sound dataset: 17 for the PSD interface, 16 for the FSD interface and 15 for the AD interface.
+60 subjects have completed the experiment for the urban sounds dataset, 20 for each interface.
+48 subjects have completed the experiment for the instrument sounds dataset: 17 for the PSD interface, 16 for the FSD interface and 15 for the AD interface.
 
@@
 Outlier detection is an important step of any crowdsourcing experiment as experimenters do not control the environment in which the subjects perform the experiment \cite{komarov2013crowdsourcing,buchholz2011crowdsourcing}.
-A commonly used method to detect outliers in  human-computer interaction studies is to consider as outlier an observation which deviates of at least $\pm2$ standard deviation from the average \cite{komarov2013crowdsourcing}.
+A commonly used method to detect outliers in  human-computer interaction studies is to consider as outlier an observation which deviates of at least $\pm2$ standard deviations from the average \cite{komarov2013crowdsourcing}.
 However, as this latter method is not robust to the presence of isolated extreme observations (as it is often the case for crowdsourcing experiment), we follow the  method proposed by \cite{komarov2013crowdsourcing} and use the Inter-Quartile Range ($IQR$).
@@
 \end{center}
-\caption{\label{urbanMusic} Average and standard deviation for (a,d) the total search durations, (b,e) the number of heard sounds and (c,f) the number of unique heard sounds; considering respectively the urban dataset (a,b,c) and the music dataset (d,e,f).}
+\caption{\label{urbanMusic} Average and standard deviation for (a,d) the total search durations, (b,e) the number of heard sounds and (c,f) the number of unique heard sounds; considering respectively the urban dataset (a,b,c) and the instrument dataset (d,e,f).}
 \end{figure}
@@
 \end{center}
-\caption{\label{fig2Music} Evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the music dataset.
+\caption{\label{fig2Music} Evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the instrument dataset.
 Lines are linear regression fits.}
@@
 \item the number of heard sounds during the whole experiment,
-\item the number of unique sounds heard during the whole experiment. By "unique" we mean that, if a same sound is heard 10 times during the 13 searches of the  experiment, it counts only for one. 
+\item the number of unique sounds heard during the whole experiment.
+By ``,unique\,'' we mean that, if a same sound is heard 10 times during the 13 searches of the  experiment, it is only counted once.
 \end{itemize}
@@
 However, those two effects are significantly compromised when users have to parse the entire hierarchy to reach the first target sound, as in the PSD display.
-It seems that enforcing an explicit exploration of the hierarchy disturbs or confuses the user instead of allowing him to learn the semantically motivated spatial organization of the classes.
+It seems that enforcing an explicit exploration of the hierarchy disturbs or confuses the user instead of allowing them to learn the semantically motivated spatial organization of the classes.
 
@@
 
-\section{Results on Musical Sounds} \label{resultsm}
+\section{Results on Instrumental Sounds} \label{resultsm}
 
@@
 
-2 subjects are detected as outliers for the music experiment and removed from the analysis.
+2 subjects are detected as outliers for the instrument sounds experiment and removed from the analysis.
 1 subject used PSD and 1 subject used FSD.
@@
 PSD and FSD significantly reduce the search durations compared to an acoustic organization (AD), which was not the case for the urban sound dataset.
-This may be explained by the fact that the musical corpus has perhaps a more obvious semantic organization.
+This may be explained by the fact that the instrumental corpus has perhaps a more obvious semantic organization.
 
@@
 Figure~\ref{fig2Music} shows the evolution of the metrics with respect to the target sound index.
-Results for the music dataset are similar to those observed for the urban dataset.
+Results for the instrument dataset are similar to those observed for the urban dataset.
 An overall increase of efficiency with respect to the index of the search is observed for the tree interfaces.

### 2016-07-05 final/lafaySsf2016.tex
creating final version

--- 2016-07-05 lafaySsf2016.tex
@@
 %\renewcommand{\baselinestretch}{2}
+
+\graphicspath{{../}}
 
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/music_mds.png} 
+\includegraphics[width=\columnwidth]{gfx/music_mds_bw.png} 
 \end{center}
@@
 
-\biography{\textbf{Nicolas Misdariis}}{gfx/lafay.jpg}{IRCAM, France}
+\biography{\textbf{Nicolas Misdariis}}{gfx/lafay.jpg}{STMS Ircam-CNRS-UPMC, France}
 
@@
 
-\biography{\textbf{Mathias Rossignol}}{gfx/rossignol.jpg}{IRCAM, France}
+\biography{\textbf{Mathias Rossignol}}{gfx/rossignol.jpg}{STMS Ircam-CNRS-UPMC, France}
 

### 2016-07-05 final/lafaySsf2016.tex
photos

@@
 
-\biography{\textbf{Gr\'egoire Lafay}}{gfx/lafay.jpg}{IRCCYN, Ecole Centrale de Nantes, France}
+\biography{\textbf{Gr\'egoire Lafay}}{../gfx/lafay.jpg}{IRCCYN, Ecole Centrale de Nantes, France}
 
-\biography{\textbf{Nicolas Misdariis}}{gfx/lafay.jpg}{STMS Ircam-CNRS-UPMC, France}
+\biography{\textbf{Nicolas Misdariis}}{../gfx/misdariisBw.jpg}{STMS Ircam-CNRS-UPMC, France}
 
-\biography{\textbf{Mathieu Lagrange}}{gfx/lagrange.jpg}{IRCCYN, Ecole Centrale de Nantes, France}
+\biography{\textbf{Mathieu Lagrange}}{../gfx/lagrange.jpg}{IRCCYN, Ecole Centrale de Nantes, France}
 
-\biography{\textbf{Mathias Rossignol}}{gfx/rossignol.jpg}{STMS Ircam-CNRS-UPMC, France}
+\biography{\textbf{Mathias Rossignol}}{../gfx/rossignol.jpg}{STMS Ircam-CNRS-UPMC, France}
 

### 2016-07-08 final/lafaySsf2016.tex
raw intro of section 2
finl version ready to ship

--- 2016-07-05 lafaySsf2016.tex
@@
 \begin{center}
-\includegraphics[scale=0.24]{gfx/dataset.pdf} 
+\includegraphics[scale=0.24]{dataset.pdf} 
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.24]{gfx/SSF.pdf} 
+\includegraphics[scale=0.24]{SSF_bw.pdf} 
 \end{center}
@@
 %\begin{center}
-%\includegraphics[width=\columnwidth]{gfx/urban_mds.png} 
+%\includegraphics[width=\columnwidth]{urban_mds.png} 
 %\end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/music_mds.png} 
+\includegraphics[width=\columnwidth]{music_mds_bw.png} 
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/urbanMusic1.eps} 
+\includegraphics[width=\columnwidth]{urbanMusic1_bw} 
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/analyse_urban2.eps} 
+\includegraphics[width=\columnwidth]{analyse_urban_bw2.pdf} 
 \end{center}
 \caption{\label{fig2urban} Evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the urban dataset.
-Lines are linear regression fits.}
+Lines are linear regression fits, represented as dashed line for AD, dash-dotted line for FSD and solid line for PSD.}
 \end{figure}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/analyse_music2.eps} 
+\includegraphics[width=\columnwidth]{analyse_music_bw2.pdf} 
 \end{center}
 \caption{\label{fig2Music} Evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the instrument dataset.
-Lines are linear regression fits.}
+Lines are linear regression fits, represented as dashed line for AD, dash-dotted line for FSD and solid line for PSD.}
 \end{figure}
@@
 
-\biography{\textbf{Gr\'egoire Lafay}}{gfx/lafay.jpg}{IRCCYN, Ecole Centrale de Nantes, France}
+\biography{\textbf{Gr\'egoire Lafay}}{lafay.jpg}{IRCCYN, Ecole Centrale de Nantes, France}
 
-\biography{\textbf{Nicolas Misdariis}}{gfx/lafay.jpg}{IRCAM, France}
+\biography{\textbf{Nicolas Misdariis}}{misdariis.jpg}{STMS Ircam-CNRS-UPMC, France}
 
-\biography{\textbf{Mathieu Lagrange}}{gfx/lagrange.jpg}{IRCCYN, Ecole Centrale de Nantes, France}
+\biography{\textbf{Mathieu Lagrange}}{lagrange.jpg}{IRCCYN, Ecole Centrale de Nantes, France}
 
-\biography{\textbf{Mathias Rossignol}}{gfx/rossignol.jpg}{IRCAM, France}
+\biography{\textbf{Mathias Rossignol}}{rossignol.jpg}{STMS Ircam-CNRS-UPMC, France}
 
-%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lafay}}]{Gr\'egoire Lafay} was born in Paris, France, in 1990.
+%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{lafay}}]{Gr\'egoire Lafay} was born in Paris, France, in 1990.
 He received in 2011 the B.S.
@@
 %
-%\begin{biography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at Irccyn, a French laboratory dedicated to cybernetics.
+%\begin{biography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at Irccyn, a French laboratory dedicated to cybernetics.
 He obtained his PhD in computer science at the University of Bordeaux in 2004, and visited several institutions, in Canada (University of Victoria, McGill University) and in France (Orange Labs, Telecom ParisTech, Ircam).
@@
 %
-%\begin{biography}[{\includegraphics[width=1.05in,height=1.5in,clip,keepaspectratio]{gfx/rossignol}}]{Mathias Rossignol} obtained his PhD from Rennes University (France) in 2005, with a specialty in Natural Language Processing.
+%\begin{biography}[{\includegraphics[width=1.05in,height=1.5in,clip,keepaspectratio]{rossignol}}]{Mathias Rossignol} obtained his PhD from Rennes University (France) in 2005, with a specialty in Natural Language Processing.
 He then focused his research on speech analysis and recognition at the International Research Center MICA (Hanoi, Vietnam), before moving on to Auditory Scene Analysis at IRCAM (Institute for Acoustic / Music Research and Coordination, Paris, France).
@@
 %
-%\begin{biography}[{\includegraphics[width=0.9in,height=1.4in,clip,keepaspectratio]{gfx/lagrange.png}}]
+%\begin{biography}[{\includegraphics[width=0.9in,height=1.4in,clip,keepaspectratio]{lagrange.png}}]
 %{Emmanouil Benetos} (S'09-M'12) received the B.Sc. and M.Sc. degrees in 

### 2016-09-08 code/analyse.m
addition of processing code

@@ -1,0 +1,223 @@
+clear all
+clc
+
+
+addpath(genpath('functions'))
+
+labels={'AD','FSD','PSD'};
+markersize_1=10;
+markersize_2=6;
+linewidth_1=3;
+linewidth_2=1.8;
+linewidthFig=1.75;
+fontsize=10;
+
+colormap(parula(10))
+cmap = colormap;
+cmap1=1;
+cmap2=6;
+cmap3=8;
+
+%% getData
+
+load('extractedData/data');
+load('extractedData/outliers')
+
+data(outliers)=[];
+
+xp=[data.xp];
+
+searchDuration=dataCell2dataMat({data.searchDuration});
+totalSearchDuration=sum(searchDuration,2);
+absoluteDuration=dataCell2dataMat({data.xpDuration});
+
+clic=dataCell2dataMat(cellfun(@(x) getClic(x,'all'),{data.sound},'UniformOutput',false));
+clicST=dataCell2dataMat(cellfun(@(x) getClic(x,'all'),{data.soundST},'UniformOutput',false));
+clicSD=dataCell2dataMat(cellfun(@(x) getClic(x,'SD'),{data.sound},'UniformOutput',false));
+clicSTSD=dataCell2dataMat(cellfun(@(x) getClic(x,'SD'),{data.soundST},'UniformOutput',false));
+
+clicT=dataCell2dataMat(cellfun(@(x) getClic(x,'T'),{data.soundST},'UniformOutput',false));
+
+totalClic=sum(clic,2);
+totalClicT=max(clicT,[],2);
+totalClicST=sum(clicST,2);
+totalClicSD=dataCell2dataMat(cellfun(@(x) getClic(x,'totalSD'),{data.sound},'UniformOutput',false));
+totalClicSTSD=dataCell2dataMat(cellfun(@(x) getClic(x,'totalSD'),{data.soundST},'UniformOutput',false));
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+figure(1)
+subplot 311
+errorbar((1:3),[mean(totalSearchDuration(xp==3,:)) mean(totalSearchDuration(xp==2,:)) mean(totalSearchDuration(xp==1,:))], ...
[…]

### 2016-09-08 code/analyse_music/analyse.m

@@ -1,0 +1,272 @@
+clear all
+clc
+
+
+addpath(genpath('../analyse_urban/functions'))
+addpath(genpath('functions'))
+
+labels={'AD','FSD','PSD'};
+markersize_1=10;
+markersize_2=6;
+linewidth_1=3;
+linewidth_2=1.8;
+linewidthFig=1.75;
+fontsize=12;
+
+colormap(parula(10))
+cmap = colormap;
+bw=1;
+
+if bw==1
+    cmap1=1;
+    cmap2=5;
+    cmap3=7;
+else
+    cmap1=1;
+    cmap2=5;
+    cmap3=8;
+end
+
+
+%% getData
+
+load('extractedData/data');
+load('extractedData/outliers')
+
+data(outliers)=[];
+
+xp=[data.xp];
+
+searchDuration=dataCell2dataMat({data.searchDuration});
+totalSearchDuration=sum(searchDuration,2);
+absoluteDuration=dataCell2dataMat({data.xpDuration});
+
+clic=dataCell2dataMat(cellfun(@(x) getClic(x,'all'),{data.sound},'UniformOutput',false));
+clicST=dataCell2dataMat(cellfun(@(x) getClic(x,'all'),{data.soundST},'UniformOutput',false));
+clicSD=dataCell2dataMat(cellfun(@(x) getClic(x,'SD'),{data.sound},'UniformOutput',false));
+clicSTSD=dataCell2dataMat(cellfun(@(x) getClic(x,'SD'),{data.soundST},'UniformOutput',false));
+
+clicT=dataCell2dataMat(cellfun(@(x) getClic(x,'T'),{data.soundST},'UniformOutput',false));
+
+totalClic=sum(clic,2);
+totalClicT=max(clicT,[],2);
+totalClicST=sum(clicST,2);
+totalClicSD=dataCell2dataMat(cellfun(@(x) getClic(x,'totalSD'),{data.sound},'UniformOutput',false));
+totalClicSTSD=dataCell2dataMat(cellfun(@(x) getClic(x,'totalSD'),{data.soundST},'UniformOutput',false));
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
[…]

### 2016-09-08 code/analyse_music/extractData.m

@@ -1,0 +1,85 @@
+clear all
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% INDICATION VARIABLE
+
+% Total : Donnees valable pour l'ensemble des sequences du sujet
+% All : Donnees valable pour l'ensemble des sujets
+% NbrTotal : somme sur l'ensemble des sequences
+% NbrAll : somme sur l'ensemble des sujets
+% SD : sans repetition
+% SC : sans target sound
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+addpath(genpath('./../analyse_urban/functions'));
+json.startup
+
+dataPath='data_test/';
+
+listing = dir([dataPath '*json']);
+
+for mm = 1:length(listing)
+    
+    navigation =  json.read([dataPath listing(mm).name]);
+    
+    timeDebut = round(navigation{end}{end}.timeDebut/1000);
+    ind_ = strfind(listing(mm).name,'_');
+    timeFin = str2double(listing(mm).name(ind_(end)+1:end-5));
+    
+    data(mm).xpDuration = timeFin-timeDebut;
+    data(mm).xp = str2double(listing(mm).name(ind_(1)+1));
+    
+    for ii = 1:length(navigation)
+        xx=1;
+        yy=1;
+        zz=1;
+        
+        %% searchTime
+        time=strsplit(navigation{ii}{end}.time,':');
+        data(mm).searchDuration(ii)= str2double(time{1})*60 + str2double(time{2});
+        
+        for jj =1:length(navigation{ii})-1
+            
+            if ~isempty(strfind(navigation{ii}{jj}.oggFile,'pause'));
+                
+                data(mm).pause{ii}{xx,1}=navigation{ii}{jj}.oggFile;
+                data(mm).pause{ii}{xx,2}=navigation{ii}{jj}.time;
+                
+                xx=xx+1;
+                
+            elseif ~isempty(strfind(navigation{ii}{jj}.oggFile,'target_sound'));
+                
+                ind=strfind(navigation{ii}{jj}.oggFile,'/');
+                targetName=navigation{ii}{jj}.oggFile(ind+1:end);
+                data(mm).sound{ii}{zz,1}=['tragetSnd_' targetName];
[…]

### 2016-09-08 code/analyse_music/functions/getSounds.m

@@ -1,0 +1,8 @@
+addpath(genpath('./../functions'))
+json.startup;
+
+
+a=loadjson('./../interface/json_csv/typologie_music.json');
+
+
+disp('')

### 2016-09-08 code/analyse_music/outlierDetection.m

@@ -1,0 +1,134 @@
+clear all
+clc
+
+%% setting
+
+addpath(genpath('functions'))
+load('extractedData/data');
+labels={'PD','FD','AD'};
+tresh=3;
+rmOutlier=0;
+
+vec2check={'absoluteDuration','totalSearchDuration','totalClicST','totalClicT'};
+
+%% get data
+
+xp=[data.xp];
+
+nb_1=sum(xp==1);
+nb_2=sum(xp==2);
+nb_3=sum(xp==3);
+
+searchDuration=dataCell2dataMat({data.searchDuration});
+totalSearchDuration=sum(searchDuration,2);
+absoluteDuration=dataCell2dataMat({data.xpDuration});
+
+clic=dataCell2dataMat(cellfun(@(x) getClic(x,'all'),{data.sound},'UniformOutput',false));
+clicST=dataCell2dataMat(cellfun(@(x) getClic(x,'all'),{data.soundST},'UniformOutput',false));
+clicSD=dataCell2dataMat(cellfun(@(x) getClic(x,'SD'),{data.sound},'UniformOutput',false));
+clicSTSD=dataCell2dataMat(cellfun(@(x) getClic(x,'SD'),{data.soundST},'UniformOutput',false));
+
+clicT=dataCell2dataMat(cellfun(@(x) getClic(x,'T'),{data.soundST},'UniformOutput',false));
+
+totalClic=sum(clic,2);
+totalClicT=max(clicT,[],2);
+totalClicST=sum(clicST,2);
+meanClicST=mean(clicST,2);
+totalClicSD=dataCell2dataMat(cellfun(@(x) getClic(x,'totalSD'),{data.sound},'UniformOutput',false));
+totalClicSTSD=dataCell2dataMat(cellfun(@(x) getClic(x,'totalSD'),{data.soundST},'UniformOutput',false));
+
+%% visu
+
+figure(1)
+subplot 221
+boxplot(absoluteDuration,xp,'whisker',3)
+title('absoluteDuration')
+subplot 222
+boxplot(totalSearchDuration,xp,'whisker',3)
+title('totalSearchDuration')
+subplot 223
+boxplot(totalClicST,xp,'whisker',3)
+title('totalSearchDuration')
+subplot 224
+boxplot(totalClicT,xp,'whisker',3)
+title('totalClicT')
+disp('')
+
+
+%% outliers
+
+outliers=[];
+
+for jj=1:length(vec2check)
[…]

### 2016-09-08 code/analyse_urban/analyse.m

@@ -1,0 +1,243 @@
+clear all
+clc
+
+addpath(genpath('functions'))
+
+labels={'AD','FSD','PSD'};
+markersize_1=10;
+markersize_2=6;
+linewidth_1=3;
+linewidth_2=1.8;
+linewidthFig=1.75;
+fontsize=12;
+
+colormap(parula(10))
+cmap = colormap;
+bw=0;
+
+if bw==1
+    cmap1=1;
+    cmap2=5;
+    cmap3=7;
+else
+    cmap1=1;
+    cmap2=5;
+    cmap3=8;
+end
+
+%% getData
+
+load('extractedData/data');
+load('extractedData/outliers')
+
+data(outliers)=[];
+
+xp=[data.xp];
+
+searchDuration=dataCell2dataMat({data.searchDuration});
+totalSearchDuration=sum(searchDuration,2);
+absoluteDuration=dataCell2dataMat({data.xpDuration});
+
+clic=dataCell2dataMat(cellfun(@(x) getClic(x,'all'),{data.sound},'UniformOutput',false));
+clicST=dataCell2dataMat(cellfun(@(x) getClic(x,'all'),{data.soundST},'UniformOutput',false));
+clicSD=dataCell2dataMat(cellfun(@(x) getClic(x,'SD'),{data.sound},'UniformOutput',false));
+clicSTSD=dataCell2dataMat(cellfun(@(x) getClic(x,'SD'),{data.soundST},'UniformOutput',false));
+
+clicT=dataCell2dataMat(cellfun(@(x) getClic(x,'T'),{data.soundST},'UniformOutput',false));
+
+totalClic=sum(clic,2);
+totalClicT=max(clicT,[],2);
+totalClicST=sum(clicST,2);
+totalClicSD=dataCell2dataMat(cellfun(@(x) getClic(x,'totalSD'),{data.sound},'UniformOutput',false));
+totalClicSTSD=dataCell2dataMat(cellfun(@(x) getClic(x,'totalSD'),{data.soundST},'UniformOutput',false));
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
[…]

### 2016-09-08 code/analyse_urban/functions/getClic.m

@@ -1,0 +1,35 @@
+function [ clic ] = getClic(data,type)
+
+switch type
+    
+    case 'all'
+        
+        clic=cellfun(@(x) size(x,1),data);
+        
+    case 'SD'
+        
+        clic=cellfun(@(x) length(unique(x(:,1))),data);
+        
+    case 'totalSD'
+        xx=1;
+        
+        for ii=1:length(data)
+            for jj=1:size(data{ii},1)
+                dataTmp{xx}=data{ii}{jj,1};
+                xx=xx+1;
+            end
+        end
+        
+        clic=length(unique(dataTmp));
+        
+    case 'T'
+        clic=cellfun(@(x) sum([x{:,3}]),data);
+        
+    otherwise
+        error('wrong input ')
+        
+end
+
+
+end
+

### 2016-09-08 code/analyse_urban/outlierDetection.m

@@ -1,0 +1,113 @@
+clear all
+clc
+
+addpath(genpath('functions'))
+load('extractedData/data');
+labels={'PD','FD','AD'};
+tresh=3;
+
+vec2check={'absoluteDuration','totalSearchDuration','totalClicST','totalClicT'};
+
+xp=[data.xp];
+
+searchDuration=dataCell2dataMat({data.searchDuration});
+totalSearchDuration=sum(searchDuration,2);
+absoluteDuration=dataCell2dataMat({data.xpDuration});
+
+clic=dataCell2dataMat(cellfun(@(x) getClic(x,'all'),{data.sound},'UniformOutput',false));
+clicST=dataCell2dataMat(cellfun(@(x) getClic(x,'all'),{data.soundST},'UniformOutput',false));
+clicSD=dataCell2dataMat(cellfun(@(x) getClic(x,'SD'),{data.sound},'UniformOutput',false));
+clicSTSD=dataCell2dataMat(cellfun(@(x) getClic(x,'SD'),{data.soundST},'UniformOutput',false));
+
+clicT=dataCell2dataMat(cellfun(@(x) getClic(x,'T'),{data.soundST},'UniformOutput',false));
+
+totalClic=sum(clic,2);
+totalClicT=max(clicT,[],2);
+totalClicST=sum(clicST,2);
+meanClicST=mean(clicST,2);
+totalClicSD=dataCell2dataMat(cellfun(@(x) getClic(x,'totalSD'),{data.sound},'UniformOutput',false));
+totalClicSTSD=dataCell2dataMat(cellfun(@(x) getClic(x,'totalSD'),{data.soundST},'UniformOutput',false));
+
+
+outliers=[];
+
+for jj=1:length(vec2check)
+    for ii=1:3
+        eval(['outliersTmp=getOutliers(' vec2check{jj} '(xp==' num2str(ii) '),' num2str(tresh) ');']);
+        
+        if ~isempty(outliersTmp)
+            
+            switch ii    
+                case 1
+                    outliers=[outliers outliersTmp];
+                    indTmp=outliersTmp;
+                    eval(['valTmp=' vec2check{jj} '(outliersTmp);'])
+                case 2
+                    outliers=[outliers 20+outliersTmp];
+                    indTmp=20+outliersTmp;
+                    eval(['valTmp=' vec2check{jj} '(20+outliersTmp);'])
+                case 3
[…]

### 2016-09-08 code/backup/extractData.m

@@ -1,0 +1,85 @@
+clear all
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% INDICATION VARIABLE
+
+% Total : Donnees valable pour l'ensemble des sequences du sujet
+% All : Donnees valable pour l'ensemble des sujets
+% NbrTotal : somme sur l'ensemble des sequences
+% NbrAll : somme sur l'ensemble des sujets
+% SD : sans repetition
+% SC : sans target sound
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+addpath(genpath('functions'));
+json.startup
+
+dataPath='data/';
+
+listing = dir([dataPath '*json']);
+
+for mm = 1:length(listing)
+    
+    navigation =  json.read([dataPath listing(mm).name]);
+    
+    timeDebut = round(navigation{end}{end}.timeDebut/1000);
+    ind_ = strfind(listing(mm).name,'_');
+    timeFin = str2double(listing(mm).name(ind_(end)+1:end-5));
+    
+    data(mm).xpDuration = timeFin-timeDebut;
+    data(mm).xp = str2double(listing(mm).name(ind_(1)+1));
+    
+    for ii = 1:length(navigation)
+        xx=1;
+        yy=1;
+        zz=1;
+        
+        %% searchTime
+        time=strsplit(navigation{ii}{end}.time,':');
+        data(mm).searchDuration(ii)= str2double(time{1})*60 + str2double(time{2});
+        
+        for jj =1:length(navigation{ii})-1
+            
+            if ~isempty(strfind(navigation{ii}{jj}.oggFile,'pause'));
+                
+                data(mm).pause{ii}{xx,1}=navigation{ii}{jj}.oggFile;
+                data(mm).pause{ii}{xx,2}=navigation{ii}{jj}.time;
+                
+                xx=xx+1;
+                
+            elseif ~isempty(strfind(navigation{ii}{jj}.oggFile,'target_Sound'));
+                
+                ind=strfind(navigation{ii}{jj}.oggFile,'/');
+                targetName=navigation{ii}{jj}.oggFile(ind+1:end);
+                data(mm).sound{ii}{zz,1}=['tragetSnd_' targetName];
[…]

### 2016-09-08 code/backup/retrieveData.sh

@@ -1,0 +1,1 @@
+scp soundthings.org:~/web/research/speedSoundFindingMusic/saveXP/* dataMusic

### 2016-09-08 code/functions/dataCell2dataMat.m

@@ -1,0 +1,8 @@
+function [ dataMat ] = dataCell2dataMat( dataCell )
+
+dataMat=zeros(length(dataCell),length(dataCell{1}));
+
+for jj=1:length(dataCell)
+    dataMat(jj,:)=dataCell{jj};
+end
+

### 2016-09-08 code/functions/getOutliers.m

@@ -1,0 +1,7 @@
+function [ outliers ] = getOutliers(data,tresh)
+
+outliers=find(data<=(quantile(data,0.25)-tresh*iqr(data)) | data>=(quantile(data,0.75)+tresh*iqr(data)));
+
+outliers=outliers(:)';
+end
+

### 2016-09-08 code/getFeatures/generateProjection.m

@@ -1,0 +1,112 @@
+function generateProjection(type, method, data, distance)
+
+if nargin==0, generateProjection('music', {'mds'}); return; end
+if ~exist('type', 'var'), type = 'music'; end
+if ~exist('method', 'var'), method = {'mds'}; end
+if ~exist('features', 'var'), features = 'mfcc'; end
+if ~exist('distance', 'var'), distance = 'euclidean'; end
+
+rng(0)
+
+%% path
+addpath(genpath('functions'));
+dataPath='data/';
+soundPath = ['sound_' type '/'];
+outputPath='./../interface/json_csv/';
+
+%% get file informations
+
+load ([dataPath 'groundtruth_' type]);
+groundTruth = cell2mat(wavFile(:,2));
+fileNames = wavFile(:,1);
+
+if exist(['data_' type '.mat'], 'file')
+    load(['data_' type]);
+else
+    %% compute features
+    data=[];
+    for k=1:length(fileNames)
+        
+        [a, fs] = audioread([soundPath fileNames{k} '.wav']);
+        a = a(:, 1);
+        
+        % features
+        switch features
+            case 'spectrum'
+                s = log10(abs(spectrogram(a, 2048)));
+            case 'mfcc'
+                s = mfcc(a);
+        end
+        % remove bad frames
+        s(:, isinf(mean(s)) | isnan(mean(s))) = [];
+        % perform average
+        data(k, :) = mean(s, 2);
+    end
+    data(isnan(data)) = 0;
+    data(isinf(data)) = 0;
+    
+    save(['data_' type '.mat'], 'data');
+end
+
+d = squareform(pdist(data,distance));
+rm = rankingMetrics(d, groundTruth);
+map1 = rm.meanAveragePrecision;
+prec1 = rm.precisionAtRank;
+
+
+%% dimensionality reduction
+for k=1:length(method)
+    switch method{k}
+        case 'mds'           
+            vd = squareform(d);
+            vd=vd./max(vd);
+            th=.16;
+            vd(vd<th)=vd(vd<th)+th;
+            [p, stress] = mdscale(vd, 2);
+            p = (p-repmat(min(p), length(fileNames), 1))./repmat(max(p)-min(p), length(fileNames), 1);
+                       
+        case 'pca'
+            [~, p] = princomp(data');
+            p = p(1:2, :)';
+        case 'som'  
+            som1=8;
+            som2=8;
[…]

### 2016-09-08 code/getFeatures/sound_music/norm_export.sh

@@ -1,0 +1,8 @@
+#!/bin/bash
+
+for i in $(ls *.wav | sed -e s/\.wav$//); 
+	do 
+		echo $i;
+		sox $i.wav ~/Dropbox/projets/ssf/interface/sound_music/$i.ogg; 
+	done 
+

### 2016-09-08 code/interface/js/XP.js

@@ -1,0 +1,247 @@
+var timercount = 0;
+var timestart  = null;
+ 
+function showtimer() {
+    if (timercount) {
+        clearTimeout(timercount);
+        clockID = 0;
+    }
+    if (!timestart) {
+        timestart = new Date();
+    }
+    var timeend = new Date();
+    var timedifference = timeend.getTime() - timestart.getTime();
+    timeend.setTime(timedifference);
+    var minutes_passed = timeend.getMinutes();
+    if (minutes_passed < 10) {
+        minutes_passed = "0" + minutes_passed;
+    }
+    var seconds_passed = timeend.getSeconds();
+    if (seconds_passed < 10) {
+        seconds_passed = "0" + seconds_passed;
+    }
+    var milliseconds_passed = timeend.getMilliseconds();
+    if (milliseconds_passed < 10) {
+        milliseconds_passed = "00" + milliseconds_passed;
+    }
+    else if (milliseconds_passed < 100) {
+        milliseconds_passed = "0" + milliseconds_passed;
+    }
+    
+    document.timeform.timetextarea.value = minutes_passed + ":" + seconds_passed + "." + milliseconds_passed;
+    timercount = setTimeout("showtimer()", 1);
+}
+ 
+function sw_start() {
+    if (!timercount) {
+	timestart   = new Date();
+	document.timeform.timetextarea.value = "00:00.000";
+	document.timeform.laptime.value = "";
+	timercount  = setTimeout("showtimer()", 1);
+    }
+    else{
+	var timeend = new Date();
+        var timedifference = timeend.getTime() - timestart.getTime();
+        timeend.setTime(timedifference);
+        var minutes_passed = timeend.getMinutes();
+        if(minutes_passed < 10){
+            minutes_passed = "0" + minutes_passed;
+        }
+        var seconds_passed = timeend.getSeconds();
+        if(seconds_passed < 10){
+            seconds_passed = "0" + seconds_passed;
+        }
+        var milliseconds_passed = timeend.getMilliseconds();
+        if(milliseconds_passed < 10){
+            milliseconds_passed = "00" + milliseconds_passed;
+        }
+        else if(milliseconds_passed < 100){
[…]

### 2016-09-08 code/interface/js/ajax.js

@@ -1,0 +1,40 @@
+function createXHR() 
+{
+    var request = false;
+        try {
+            request = new ActiveXObject('Msxml2.XMLHTTP');
+        }
+        catch (err2) {
+            try {
+                request = new ActiveXObject('Microsoft.XMLHTTP');
+            }
+            catch (err3) {
+		try {
+			request = new XMLHttpRequest();
+		}
+		catch (err1) 
+		{
+			request = false;
+		}
+            }
+        }
+    return request;
+}
+
+
+
+function Write(url, content)	// url is the script and data is a string of parameters
+	{ 
+		var xhr = createXHR();
+
+		xhr.onreadystatechange=function()
+		{ 
+			if(xhr.readyState == 4)
+			{
+				//alert("Envoyé " + content);
+			}
+		}; 
+		xhr.open("POST", url, true);		
+		xhr.setRequestHeader("Content-Type", "application/x-www-form-urlencoded");
+		xhr.send(content); 
+	} 

### 2016-09-08 code/interface/js/main.js

@@ -1,0 +1,129 @@
+
+// Ecouteurs
+// https://developer.mozilla.org/en-US/docs/DOM/EventTarget.addEventListener
+
+var sndIndex;
+
+function Construction_ecouteurs(_sndCollection,nodeTemp,dataTemp,last,parentName,cercleNoir){
+    // NB : _sndCollection garde sa taille max .....
+    if(cercleNoir)
+	sndIndex = d3.selectAll("circle").filter(function(d){return d.name == parentName}).attr('sndIndex');
+    for(ii=0;ii<dataTemp.length-1;ii++)	{
+	_sndCollection[ii].addEventListener("ended", function(_event) {
+	    focusSound(nodeTemp,sndIndex,dataTemp,last);
+    	    _sndCollection[sndIndex].play()
+	    
+	    // relevé temps 
+	    saveDataXP(dataTemp[sndIndex].oggFile)				
+	    sndIndex++;
+	    //console.log(sndIndex)
+	    d3.selectAll("circle").filter(function(d){return d.name == parentName}).attr('sndIndex',function(d){return sndIndex});
+	    //console.log(parentName + d3.selectAll("circle").filter(function(d){return d.name == parentName}).attr('sndIndex'));
+    	},false);	
+    }
+    
+    _sndCollection[dataTemp.length-1].addEventListener("ended", function(_event) {
+    	sndIndex=1;
+	d3.selectAll("circle").filter(function(d){return d.name == parentName}).attr('sndIndex', function(d){return sndIndex});
+	//console.log(parentName + d3.selectAll("circle").filter(function(d){return d.name == parentName}).attr('sndIndex'));
+	focusSound(nodeTemp,0,dataTemp,last);
+    	_sndCollection[0].play()
+	
+	// relevé temps 
+	saveDataXP(dataTemp[0].oggFile)	
+	
+    },false);
+	    					
+    focusSound(nodeTemp,sndIndex-1,dataTemp,last);
+    _sndCollection[sndIndex-1].play();
+    //console.log(parentName + d3.selectAll("circle").filter(function(d){return d.name == parentName}).attr('sndIndex'));	
+}	
+
+// focus (trait gras)
+function focusSound(nodeTemp,index,dataTemp,last) {
+    // test index du sample
+    if(index==0) {
+	var index1 = dataTemp.length-1;
+	var index2 = index;
+    } else {
+	var index1 = index-1;
+	var index2 = index;
+    }
+    // suppression old focus node
+    circleTemp=dataTemp[index1].name;
[…]

### 2016-09-08 code/paperFigure/generateFig.m

@@ -1,0 +1,73 @@
+clear all
+clc
+
+addpath(genpath('../analyse_urban/functions'))
+
+labels={'AD','FSD','PSD'};
+markersize_1=10;
+markersize_2=6;
+linewidth_1=3;
+linewidth_2=1.8;
+linewidthFig=1.75;
+fontsize=10;
+
+load('data/urban.mat');
+load('data/music.mat');
+
+figure(1)
+subplot 321
+errorbar((1:3),stats_totalSearchDuration_urban(1,:), ...
+    stats_totalSearchDuration_urban(2,:),'d','markersize',markersize_1,'linewidth',linewidth_1)
+set(gca,'xtick',1:3,'xticklabel',labels,'box','off','ytick',200:200:1200,'linewidth',linewidthFig,'fontsize',fontsize)
+title('a')
+ylabel('Duration (sec)')
+ylim([100 1300])
+
+subplot 323
+errorbar((1:3),stats_totalClicST_urban(1,:), ...
+    stats_totalClicST_urban(2,:),'d','markersize',markersize_1,'linewidth',linewidth_1)
+set(gca,'xtick',1:3,'xticklabel',labels,'box','off','ytick',200:100:800,'linewidth',linewidthFig,'fontsize',fontsize)
+ylabel('Heard sounds')
+title('b')
+ylim([150 850])
+
+subplot 325
+errorbar((1:3),stats_totalClicSTSD_urban(1,:), ...
+    stats_totalClicSTSD_urban(2,:),'d','markersize',markersize_1,'linewidth',linewidth_1)
+set(gca,'xtick',1:3,'xticklabel',labels,'box','off','ytick',100:10:150,'linewidth',linewidthFig,'fontsize',fontsize)
+title('c')
+ylabel('Unique heard sounds')
+ylim([95 155])
+
+subplot 322
+errorbar((1:3),stats_totalSearchDuration_music(1,:), ...
+    stats_totalSearchDuration_music(2,:),'d','markersize',markersize_1,'linewidth',linewidth_1)
+set(gca,'xtick',1:3,'xticklabel',labels,'box','off','ytick',200:200:1200,'linewidth',linewidthFig,'fontsize',fontsize)
+title('d')
+ylabel('Duration (sec)')
+ylim([100 1300])
+
+subplot 324
+errorbar((1:3),stats_totalClicST_music(1,:), ...
[…]

### 2016-09-08 code/soundProjection/generateGroundTruth.m

@@ -1,0 +1,32 @@
+function generateGroundTruth(type, targetLevel)
+
+if ~exist('type', 'var'), type = 'urban'; end
+if ~exist('targetLevel', 'var'), targetLevel = 2; end
+
+a=loadjson(['../interface/json_csv/typologie_' type '.json']);
+
+[wavFile, ~, ~, classNames] = parseTree(a, {}, 1, 1, {}, targetLevel);
+
+gt = cell2mat(wavFile(:,2))';
+classNames
+save (['../getFeatures/data/groundtruth_' type], 'wavFile', 'classNames');
+
+
+function [gt, id, level, names, targetLevel] = parseTree(tree, gt, id, level, names, targetLevel)
+
+if ~tree.last
+    for k=1:length(tree.children)
+        [gt, id, ~, names, targetLevel] = parseTree(tree.children{k}, gt, id, level+1, names, targetLevel);
+    end
+else
+    for k=1:length(tree.children)
+        [~, name] = fileparts(tree.children{k}.oggFile);
+        gt{end+1, 1} = name;
+        gt{end, 2} = id;
+    end
+end
+
+if level==targetLevel
+    id=id+1;
+    names{end+1} = tree.name;
+end

### 2016-09-08 code/soundProjection/sopr1features.m

@@ -1,0 +1,48 @@
+function [config, store, obs] = sopr1features(config, setting, data)
+% sopr1features FEATURES step of the expLanes experiment soundProjection
+%    [config, store, obs] = sopr1features(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: Mathieu Lagrange
+% Date: 12-May-2016
+
+% Set behavior for debug mode
+if nargin==0, soundProjection('do', 1, 'mask', {2}); return; else store=[]; obs=[]; end
+
+
+
+
+%% get file informations
+load ([config.inputPath 'data/groundtruth_' setting.database]);
+groundTruth = cell2mat(wavFile(:,2));
+fileNames = wavFile(:,1);
+
+
+%% compute features
+features=[];
+for k=1:length(fileNames)   
+    [a, fs] = audioread([config.inputPath 'sound' '_' setting.database '/' fileNames{k} '.wav']);
+    a = a(:, 1);
+    
+    % features
+    switch setting.features
+        case 'spectrum'
+            s = log10(abs(spectrogram(a, 2048)));
+        case 'mfcc'
+            s = mfcc(a);
+    end
+    % remove bad frames
+    s(:, isinf(mean(s)) | isnan(mean(s))) = [];
+    % perform average
+    features(k, :) = mean(s, 2);
+end
+features(isnan(features)) = 0;
+features(isinf(features)) = 0;
+
+store.features = features;
+store.groundTruth = groundTruth;
+store.classNames = classNames;

### 2016-09-08 code/soundProjection/sopr2projection.m

@@ -1,0 +1,67 @@
+function [config, store, obs] = sopr2projection(config, setting, data)
+% sopr2projection PROJECTION step of the expLanes experiment soundProjection
+%    [config, store, obs] = sopr2projection(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: Mathieu Lagrange
+% Date: 12-May-2016
+
+% Set behavior for debug mode
+if nargin==0, soundProjection('do', 2, 'mask', {0 1 3}); return; else store=[]; obs=[]; end
+
+load ([config.inputPath 'data/groundtruth_' setting.database]);
+groundTruth = cell2mat(wavFile(:,2));
+
+switch setting.projection
+    case 'id'
+        p = data.features;
+    case 'pca'
+        f = data.features';
+        [~, p] = princomp(f);
+        p = p(1:2, :)';
+    case 'mds'
+        vd = pdist(data.features,setting.distance);
+        vd=vd./max(vd);
+        th=.16;
+        vd(vd<th)=vd(vd<th)+th;
+        [p, stress] = mdscale(vd, 2);
+        %  p = (p-repmat(min(p), length(fileNames), 1))./repmat(max(p)-min(p), length(fileNames), 1);
+    case 'som'
+        net = selforgmap([setting.neurons setting.neurons]);
+        net.trainParam.showWindow=0;
+        net = train(net, data.features');
+        pos = gridtop(setting.neurons, setting.neurons);
+        y = net(data.features');
+        for m=1:size(data.features, 1)
+            p(m, :) = pos(:, find(y(:, m)))+setting.spread*rand(2, 1);
+        end
+end
+
+
+switch setting.database
+    case 'urban'
+        classNames = {'Traffic', 'Sirens', 'Nature', 'Leisure', 'Construction work', 'Human moves', 'Human Voice'};
+    case 'music'
+        classNames = {'Brass', 'Percussion', 'Piano', 'Voice', 'String', 'Woodwinds'};
+end
+for k=1:length(data.groundTruth)
+    gt{k} = classNames{groundTruth(k)};
+end
+h=gscatter(p(:, 1), p(:, 2), gt');
+set(h, 'Markersize', 25);
+axis off
[…]

### 2016-09-08 code/soundProjection/soprInit.m

@@ -1,0 +1,10 @@
+function [config, store] = soprInit(config)                         
+% soprInit INITIALIZATION of the expLanes experiment soundProjection
+%    [config, store] = soprInit(config)                             
+%      - config : expLanes configuration state                      
+%      -- store  : processing data to be saved for the other steps  
+                                                                    
+% Copyright: Mathieu Lagrange                                       
+% Date: 12-May-2016                                                 
+                                                                    
+if nargin==0, soundProjection(); return; else store=[];  end        

### 2016-09-08 code/soundProjection/soprReport.m

@@ -1,0 +1,17 @@
+function config = soprReport(config)                             
+% soprReport REPORTING of the expLanes experiment soundProjection
+%    config = soprInitReport(config)                             
+%       config : expLanes configuration state                    
+                                                                 
+% Copyright: Mathieu Lagrange                                    
+% Date: 12-May-2016                                              
+                                                                 
+if nargin==0, soundProjection('report', 'rcv'); return; end        
+                                                                 
+                                
+
+config = expExpose(config, 't', 'obs', [2 5], 'mask', {1 1 [1 2 3 4] 1 9 6 1}); 
+
+config = expExpose(config, 't', 'obs', [2 5], 'mask', {2 1 [1 2 3 4] 1 9 6 1}); 
+
+config = expExpose(config, 'p', 'obs', 2, 'mask', {1, 4, 1, 0, 5, 0}, 'expand', 'neurons'); 

### 2016-09-08 code/soundProjection/soundProjection.m

@@ -1,0 +1,190 @@
+function soundProjection(varargin)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              
+% Welcome to the main entry point of soundProjection                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                            
+% Please DO NOT modify this file unless you have a precise intent.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              
+                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                
+shortExperimentName = 'sopr';                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                   
+[p, experimentName] = fileparts(mfilename('fullpath'));                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         
[…]
